---
title: "Initial GSS EDA and APC - health and covariates"
author: "Christine Lucille Kuryla"
date: "2024-09-27"
output: 
  html_document:
    self_contained: true
    toc: true
knit_root_dir: "docs"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(gridExtra)
library(here)
```

Take a look at GSS dataset, what's available, some trends, etc. 

The first variable of interest is "health", which will be the main subject of our analysis. 

https://gssdataexplorer.norc.org/variables/437/vshow

Question on survey: "Would you say your own health, in general, is excellent, good, fair, or  poor?"

# Fetch GSS data 

```{r fetch_data, eval = FALSE}

# Feel free to modify to play with more covariates and variables.

#install.packages('gssr', repos =  c('https://kjhealy.r-universe.dev', 'https://cloud.r-project.org'))
# install.packages('gssrdoc', repos = c('https://kjhealy.r-universe.dev', 'https://cloud.r-project.org'))

library(gssr)
library(gssrdoc)

data("gss_all") # this file is big! 

# It's a bit excessive to download the entire GSS dataset every time we knit, so lets just save some variables of interest and write it out for future use.

data_gss <- as.data.frame(gss_all) %>% 
  select("year",      # year of survey
         "cohort",    # birthyear
         "age",       # age at time of survey
         "health",    # self-rated health
         "sex",       # sex
         "happy",     # self-rated happiness
         "life",      # is life exciting or dull
         "educ",      # years of education
         "polviews",  # 1 extremely liberal, 4 moderate, 7 extremely conservative
         "class",    # 1 lower, 2 middle, 3 working, 4 upper, 5 no class
         "satfin"    # 1 pretty well satisfied, 2 more or less satisfied, 3 not satisfied at all
         )

write_csv(data_gss, here("data/extracted_gss_variables.csv"))

```


# Variables of interest

<details>
  <summary>**Click to expand and explore the variables of interest, including the coding and the way the question is phrased.**</summary>

### "health"

https://gssdataexplorer.norc.org/variables/437/vshow?back=variableList

Variable: Health
Module: Core
Gss Tags: Health

"Would you say your own health, in general, is excellent, good, fair, or  poor?"

-100	.i: Inapplicable	
-99	.n: No answer	
-98	.d: Do not Know/Cannot Choose	
-97	.s: Skipped on Web	
1	Excellent	
2	Good	
3	Fair	
4	Poor

*Note* we will recode this data to reverse it to be more intuitive, such that for our analyses, higher numbers will indicate better self-rated health:

4	Excellent	
3	Good	
2	Fair	
1	Poor

### "year"

https://gssdataexplorer.norc.org/variables/1/vshow?back=variableList

Year of survey

Variable: Year

GSS year for this respondent

### "cohort"

https://gssdataexplorer.norc.org/variables/5507/vshow?back=variableList

Year of birth

Birth cohort of respondent.

### "happy"

general happiness

"Taken all together, how would you say things are these days--would you say that you are very happy, pretty happy, or not too happy?"

-100	.i: Inapplicable	
-99	.n: No answer	
-98	.d: Do not Know/Cannot Choose	
-97	.s: Skipped on Web	
1	Very happy	
2	Pretty happy	
3	Not too happy	

*Note* we will recode this data to reverse it to be more intuitive, such that for our analyses, higher numbers will indicate higher happiness:

3	Very happy	
2	Pretty happy	
1	Not too happy	

### "sex"

https://gssdataexplorer.norc.org/variables/81/vshow?back=variableList

respondent's sex

-100	.i: Inapplicable	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	19	0
-99	.n: No answer	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	71	9
-98	.d: Do not Know/Cannot Choose	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	8
-97	.s: Skipped on Web	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	2	3
1	MALE	
2	FEMALE	

### "educ"

https://gssdataexplorer.norc.org/variables/55/vshow

highest year of school completed

Questions associated with this variable:
ASK ALL PARTS OF QUESTION ABOUT RESPONDENT BEFORE GOING ON TO ASK ABOUT R'S FATHER; AND THEN R'S MOTHER; THEN R'S SPOUSE, IF R IS CURRENTLY MARRIED. 
A. What is the highest grade in elementary school or high school that (you/your father/ your mother/your [husband/wife]) finished and got credit for? CODE EXACT GRADE. 
B. IF FINISHED 9th-12th GRADE OR DK*: Did (you/he/she) ever get a high school diploma or a GED certificate? [SEE D BELOW.] [See REMARKS] 
C. Did (you/he/she) complete one or more years of college for credit--not including schooling such as business college, technical or vocational school? IF YES: How many years did (you/he/she) complete? 
 Do you (Does [he/she]) have any college degrees? (IF YES: What degree or degrees?) CODE HIGHEST DEGREE EARNED.

-99	.n: No answer	5	5	3	2	4	3	3	1	4	0	3	0	1	6	0	5	0	7	0	4	1	4	9	7	2	9	2	4	1	0	8	1	48	20
-98	.d: Do not Know/Cannot Choose	0	0	0	1	2	7	3	4	4	2	0	0	0	4	3	2	2	0	4	3	8	8	0	5	0	2	3	1	1	1	1	2	18	0
0	No formal schooling	14	7	6	7	4	4	6	1	7	5	3	3	6	9	1	3	4	2	3	4	4	2	1	5	4	22	6	5	3	8	2	4	9	3
1	1st grade	4	4	1	2	0	0	4	1	2	1	3	2	0	2	1	2	1	0	0	1	0	0	0	2	1	4	0	1	2	1	3	2	1	1
2	2nd grade	13	7	1	3	4	1	1
...
19 7 or more years of college
20 8 or nore years of college

### "life"

is life exciting or dull

https://gssdataexplorer.norc.org/variables/438/vshow

"In general, do you find life exciting, pretty routine, or dull?"

100	.i: Inapplicable	
-99	.n: No answer	
-98	.d: Do not Know/Cannot Choose	
-97	.s: Skipped on Web	
1	Exciting	
2	Routine	
3	Dull

*Note* we will recode this data to reverse it to be more intuitive, such that for our analyses, higher numbers will indicate a more exciting life:

3	Exciting	
2	Routine	
1	Dull

### "satfin"

satisfaction with financial situation

https://gssdataexplorer.norc.org/variables/572/vshow

"We are interested in how people are getting along financially these days. So far as you and your family are concerned, would you say that you are pretty well satisfied with your present financial situation, more or less satisfied, or not satisfied at all?"

-100	.i: Inapplicable	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	1393	1472	1518	0	0	0	0	0	0	0	0
-99	.n: No answer	4	2	4	7	3	6	1	2	8	2	6	4	1	6	4	1	2	6	3	11	4	3	5	1	2	8	2	5	5	3	6	3	1	4
-98	.d: Do not Know/Cannot Choose	1	1	2	4	4	3	2	4	5	5	3	5	3	2	3	4	3	5	7	5	4	5	9	2	3	4	5	1	2	3	5	7	0	7
-97	.s: Skipped on Web	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	15	7
1	Pretty well satisfied	
2	More or less satisfied	
3	Not satisfied at all	

*Note* we will recode this data to reverse it to be more intuitive, such that for our analyses, higher numbers will indicate more financial satisfaction:

3	Pretty well satisfied	
2	More or less satisfied	
1	Not satisfied at all	

### "polviews"

think of self as liberal or conservative

https://gssdataexplorer.norc.org/variables/178/vshow

"We hear a lot of talk these days about liberals and conservatives. 
I'm going to show you a seven-point scale on which the political views that people might hold are arranged from extremely liberal--point 1--to extremely conservative--point 7. Where would you place yourself on this scale?"

-100	.i: Inapplicable	0	0	0	0	0	0	0	795	0	0	0	0	0	0	0	0	0	0	0	0	0	1393	1472	0	0	0	0	0	0	0	15	1
-99	.n: No answer	4	12	5	6	27	17	17	3	11	9	2	42	9	7	4	4	9	12	6	8	24	5	5	23	13	10	13	24	30	22	3	27
-98	.d: Do not Know/Cannot Choose	70	81	93	71	70	22	104	31	52	63	67	98	56	88	53	54	49	101	155	133	149	36	26	154	77	61	87	65	81	79	10	73
-97	.s: Skipped on Web	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	0	40	17
1	Extremely liberal	
2	Liberal	
3	Slightly liberal	
4	Moderate, middle of the road	
5	Slightly conservative	
6	Conservative	
7	Extremely conservative	

</details>

# Load and clean data

Here we'll load our data, clean some unwanted values, and recode the unintuitive variables.

```{r load_and_clean_data}
data_gss <- read_csv(here("data/extracted_gss_variables.csv")) %>% 
  filter(cohort != 9999) %>% 
  na.omit() %>% 
  mutate(health = 5 - health)  %>%  # reverse the coding so it's more intuitive (higher number for excellent, lower number for poor)
  mutate(happy = 4 - happy) %>% # same
  mutate(life = 4 - life) %>% # reverse again, these variables tend to be unintuitively ordered!!!
  mutate(satfin = 4 - satfin) # %>% 
  # mutate(polviews_extreme = case_match(polviews,
  #                                      7 ~ 4,
  #                                      6 ~ 3,
  #                                      5 ~ 2,
  #                                      4 ~ 1,
  #                                      3 ~ 2,
  #                                      2 ~ 3,
  #                                      1 ~ 4
  #                                      ))
```

## Peek at GSS df
```{r head}

head(data_gss, n = 10)

glimpse(data_gss)

```


## Life expectancy data

```{r, eval = FALSE}

# https://population.un.org/wpp/Download/Standard/CSV/
data_le <- read_csv("big_data/WPP2024_Demographic_Indicators_Medium.csv")
data_le <- data_le %>% 
  filter(Location == "United States of America")

# Just get the years of interest for merging with the gss data
year_range <- seq(min(data_gss$year), max(data_gss$year))
data_le <- data_le %>% 
  filter(Time %in% year_range)

write_csv(data_le, "data/un_life_expectancy_etc.csv")

```


# Histograms

Let's make a few quick histograms to get a sense of the survey response distributions.

```{r histograms}

# Tidyverse and flexible number of histograms

library(patchwork)

# Create list to store ggplot objects
plot_list <- list()

# Loop through column names and create plots
for (var in colnames(data_gss)) {
  if (is.numeric(data_gss[[var]])) {  # Only create histograms for numeric columns
    p <- ggplot(data_gss, aes(x = .data[[var]])) +  # Use .data[[var]] for tidy evaluation
      geom_histogram(bins = 20, fill = "pink", color = "hotpink") +
      theme_minimal()
    
    plot_list[[var]] <- p
  }
}

# Combine all plots using patchwork and add a title
combined_plot <- wrap_plots(plot_list, ncol = 3) +
  plot_annotation(title = "Histograms of Survey Responses")

# Display the combined plot
print(combined_plot)



```


# Correlation Heatmap

Just to get some idea and familiarize ourselves with the data.

```{r}
correlation_auto <- cor(data_gss)
knitr::kable(correlation_auto)

# Create a heatmap

#upper_tri <- matrixcalc::upper.triangle(correlation_auto)
#melted_cormat <- reshape2::melt(upper_tri, na.rm = TRUE)

melted_cormat <- reshape2::melt(cor(data_gss), na.rm = TRUE)

ggplot(data = melted_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
     hjust = 1))+
 coord_fixed() +
  geom_text(aes(Var2, Var1, label = if_else(value != 0, as.character(round(value, digits = 2)), " ")))
```

# Mean health over Age/Period/Cohort

Now let's take a look at the trend of people's self-rated health over time, as well as by age, and by birthyear.

```{r over_apc}

# Mean health over time, age, and birth year

data_gss %>% 
  group_by(year) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = year, y = mean_health)) +
  geom_line(color = "cornflowerblue") +
  labs(title = "Period: Self-Rated Health Each Year",
       subtitle = "(all ages together)" )

data_gss %>% 
  group_by(age) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = age, y = mean_health)) +
  geom_line(color = "cornflowerblue") +
  labs(title = "Age: Self-Rated Health By Age" ,
       subtitle = "(all years together)" )

data_gss %>% 
  group_by(cohort) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = cohort, y = mean_health)) +
  geom_line(color = "cornflowerblue") +
  labs(title = "Cohort: Self-Rated Health by Birth Year",
       subtitle = "(all years together)" )

```





# Cohorts 

Let's explore how each cohort tends to rate their health when at the same age.

```{r more_cohorts}

data_gss %>% 
  mutate(cohort = cut(cohort, breaks = 6)) %>% # Create cohorts with 6 breaks
  group_by(age, cohort) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = age, y = mean_health, color = cohort)) +
  geom_line()

data_gss %>% 
  mutate(cohort = cut(cohort, breaks = 6)) %>% 
  group_by(year, cohort) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = year, y = mean_health, color = cohort)) +
  geom_line()

data_gss %>% 
  mutate(cohort = cut(cohort, breaks = 6)) %>% 
  group_by(happy, cohort) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = happy, y = mean_health, color = cohort)) +
  geom_line()

data_gss %>% 
  mutate(cohort = cut(cohort, breaks = 6)) %>% # Create cohorts with 6 breaks
  group_by(age, cohort) %>% 
  summarize(mean_happy = mean(happy)) %>% 
  ggplot(aes(x = age, y = mean_happy, color = cohort)) +
  geom_line()

data_gss %>% 
  mutate(age = cut(age, breaks = 6)) %>% # Create cohorts with 6 breaks
  group_by(age, year) %>% 
  summarize(mean_happy = mean(happy)) %>% 
  ggplot(aes(x = year, y = mean_happy, color = age)) +
  geom_line()

data_gss %>% 
  mutate(age = cut(age, breaks = 6)) %>% # Create cohorts with 6 breaks
  group_by(age, year) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = year, y = mean_health, color = age)) +
  geom_line()

data_gss %>% 
  mutate(cohort = cut(cohort, breaks = 6)) %>% # Create cohorts with 6 breaks
  group_by(cohort, year) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = year, y = mean_health, color = cohort)) +
  geom_line()

data_gss %>% 
  mutate(cohort = cut(cohort, breaks = 6)) %>% # Create cohorts with 6 breaks
  group_by(year, cohort) %>% 
  summarize(mean_happy = mean(happy)) %>% 
  ggplot(aes(x = year, y = mean_happy, color = cohort)) +
  geom_line()

```


# Interesting and seemingly robust

Well, this is probably the most interesting graph. Let's try different numbers of categories of age.

```{r multiple_cohort_breaks_health}
data_gss %>% 
  mutate(age = cut(age, breaks = 6)) %>% # Create cohorts with 6 breaks
  group_by(age, year) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = year, y = mean_health, color = age)) +
  geom_line() +
  geom_point()

data_gss %>% 
  mutate(age = cut(age, breaks = 10)) %>% # Create cohorts with 6 breaks
  group_by(age, year) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = year, y = mean_health, color = age)) +
  geom_line()

data_gss %>% 
  mutate(age = cut(age, breaks = 3)) %>% # Create cohorts with 6 breaks
  group_by(age, year) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = year, y = mean_health, color = age)) +
  geom_line()

data_gss %>% 
  mutate(age = cut(age, breaks = 4)) %>% # Create cohorts with 6 breaks
  group_by(age, year) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = year, y = mean_health, color = age)) +
  geom_line()
```

# Relationship of self-rated health to age, separated out by years

Well it seems like the spread of self-rated health among ages decreases as time goes on (later years). Let's look at that.

```{r health_v_age_per_year}
# health vs age per year
data_gss %>% 
  group_by(age, year) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = age, y = mean_health)) +
  geom_line(color = "cornflowerblue") +
  facet_wrap(~ year) +
  labs(title = "Self-Rated Health By Age (Per Year)" )
```

# Regress the age coefficient on year

```{r regress_age_coeff}

library(broom)

# Aggregate slopes

# years_of_gss <- c(data_gss %>% select(year) %>% unique() )
# lm_health_v_age_0 <- data_gss %>%
#   group_by(year) %>%
#   summarize(coef = coef(lm(health ~ age, data = cur_data()))["age"])

# Perform linear regression for each year and extract the coefficient of 'age' with confidence intervals, se, t stat, p val
lm_health_v_age_0 <- data_gss %>%
  group_by(year) %>%
  do(tidy(lm(health ~ age, data = .), conf.int = TRUE)) %>%  # Add conf.int = TRUE for CIs
  filter(term == "age") %>%
  select(year, coef = estimate, conf.low, conf.high, se = std.error, t_statistic = statistic,  p_value = p.value)

# View the results with confidence intervals, se, t statistic, and p value
# print(lm_health_v_age_0)
knitr::kable(lm_health_v_age_0)

# Plot coefficients
ggplot(lm_health_v_age_0, aes(x = year, y = coef)) +
  geom_line() +
  geom_point() +
  labs(
    title = "Change in 'Age' Coefficient Over Years",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()

# Plot coefficients with CI
ggplot(lm_health_v_age_0, aes(x = year, y = coef)) +
  geom_line() +
  geom_point() +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +  # Add shaded area for confidence intervals
  labs(
    title = "Change in 'Age' Coefficient Over Years with Confidence Intervals",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()

# Perform linear regression of 'coef' (age coefficient) vs 'year'
lm_coef_vs_year <- lm(coef ~ year, data = lm_health_v_age_0)

# View the summary of the regression
summary(lm_coef_vs_year)

ggplot(lm_health_v_age_0, aes(x = year, y = coef)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +  # Adds the regression line with standard error shading
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +  # Confidence intervals for the coefficients
  labs(
    title = "Regression of 'Age' Coefficient Over Years",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()


```

```{r multiple_cohort_breaks_happy}
data_gss %>% 
  mutate(age = cut(age, breaks = 6)) %>% # Create cohorts with 6 breaks
  group_by(age, year) %>% 
  summarize(mean_happy = mean(happy)) %>% 
  ggplot(aes(x = year, y = mean_happy, color = age)) +
  geom_line()

data_gss %>% 
  mutate(age = cut(age, breaks = 6)) %>% # Create cohorts with 6 breaks
  group_by(age, year) %>% 
  summarize(mean_happy = mean(happy)) %>% 
  ggplot(aes(x = year, y = mean_happy, color = age)) +
  geom_line()

data_gss %>% 
  mutate(cohort = cut(cohort, breaks = 6)) %>% # Create cohorts with 6 breaks
  group_by(cohort, year) %>% 
  summarize(mean_happy = mean(happy)) %>% 
  ggplot(aes(x = year, y = mean_happy, color = cohort)) +
  geom_line()
```


```{r}
data_gss %>% 
  mutate(cohort = cut(cohort, breaks = 4)) %>% # Create cohorts with 6 breaks
  group_by(cohort, year) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = year, y = mean_health, color = cohort)) +
  geom_line()

data_gss %>% 
  mutate(cohort = cut(cohort, breaks = 4)) %>% # Create cohorts with 6 breaks
  group_by(cohort, age) %>% 
  summarize(mean_health = mean(health)) %>% 
  ggplot(aes(x = age, y = mean_health, color = cohort)) +
  geom_line()

```



## Happiness and health

```{r happiness_and_health}

data_gss %>% 
  ggplot(aes(x = health, y = happy)) +
  geom_smooth(method = "lm")

data_gss %>% 
  mutate(year = cut(year, breaks = 6)) %>% 
  mutate(age = cut(age, breaks = 6)) %>% 
  ggplot(aes(x = health, y = happy, color = age)) +
  facet_wrap(~ year) +
  geom_smooth(method = "lm")

data_gss %>% 
  mutate(cohort = cut(cohort, breaks = 6)) %>% 
  mutate(year = cut(year, breaks = 6)) %>% 
  ggplot(aes(x = health, y = happy, color = cohort)) +
  facet_wrap(~ year) +
  geom_smooth(method = "lm")

######data_gss %>% 
  # group_by(year) %>% 
  # summarize(mean_health = mean(health)) %>% 
  # ggplot(aes(x = cohort, y = mean_health)) +
  # geom_line()

data_gss %>% 
  mutate(cohort = cut(cohort, breaks = 6)) %>% 
#  mutate(year = cut(year, breaks = 6)) %>% 
#  mutate(age = cut(year, breaks = 6)) %>% 
  ggplot(aes(x = age, y = health, color = cohort)) +
#  facet_wrap(~ year) +
  geom_smooth(method = "lm")

```

```{r}


library(ggpubr)

data_gss %>%
  group_by(year) %>%
  do(tidy(lm(health ~ polviews, data = .), conf.int = TRUE)) %>%  # Add conf.int = TRUE for CIs
  filter(term == "polviews") %>%
  select(year, coef = estimate, conf.low, conf.high, se = std.error, t_statistic = statistic,  p_value = p.value) %>% 
  ggplot(aes(x = year, y = coef)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +  # Adds the regression line with standard error shading
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +  # Confidence intervals for the coefficients
  labs(
    title = "Regression of 'Age' Coefficient Over Years",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  stat_cor(label.x = 2000) +#label.y = 40)+ 
                   stat_regline_equation()+#label.y = 45) +
  theme_minimal()

```

