---
title: "NHANES 3 and 4"
author: "Christine Lucille Kuryla"
date: "2024-12-13"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)

```

In this document, we will look at SRH for NHANES III and NHANES IV. 

I'm not sure how to incorporate weights for NHANES III or if it's even comparable to NHANES IV, so we will not weight.

# Import and check data

For reference:
### NHANES 3 
```{r, eval = FALSE}

library(SAScii)
# nhanes3.tf <- tempfile()
daturl <- "https://wwwn.cdc.gov/nchs/data/nhanes3/1a/adult.dat"
code_url ="https://wwwn.cdc.gov/nchs/data/nhanes3/1a/adult.sas"
# Sas_code <- url(code_url)
# writeLines ( readLines(Sas_code) , con = nhanes3.tf )
# nhanes3.fwf.parameters <- parse.SAScii( nhanes3.tf , beginline = 5 )
# str( nhanes3.fwf.parameters )
# #-----
# 'data.frame':   90 obs. of  4 variables:
#   $ varname: chr  "SEQN" "HYK1A" "HYK1B" "HYK2A" ...
# $ width  : num  5 1 1 2 2 2 2 4 4 2 ...
# $ char   : logi  FALSE FALSE FALSE FALSE FALSE FALSE ...
# $ divisor: num  1 1 1 1 1 1 1 1 1 1 ...
# #------

daturl <- "https://wwwn.cdc.gov/nchs/data/nhanes3/1a/adult.dat"
in.nhanes3 <- read.fwf(daturl, widths=nhanes3.fwf.parameters$width,
                     col.names= nhanes3.fwf.parameters$varname)

in2 <- read.SAScii( daturl, code_url)

#write_csv(in2, "big_data/NHANES/nhanes_3/nhanes3.csv")

nhanes3_data <- read_csv("big_data/NHANES/nhanes_3/nhanes3.csv")

nhanes3_selected <- nhanes3_data %>% 
  select(SEQN, 
         DMPFSEQ, 
         HSAGEIR, # age in years
         HAB1, # self-rated health: 1:excellent, very good, good, fair, 5: poor (get rid of 6 and 7)
         HSSEX, # 1 male, 2 female
         SDPPHASE, # 1 1988-1991, 2 1991-1994 
         HSDOIMO, # date of screener (month)
         HSAGEU, # age unit
         HSAITMOR # age in months at interview (screener)
         ) %>% 
  filter(HAB1 %in% 1:5) %>% 
  mutate(age = HSAGEIR,
         sex = ifelse(HSSEX == 1, "Male", "Female"),
         year = ifelse(SDPPHASE == 1, 1989.5, 1992.5),
         srh = 6 - HAB1)

glimpse(nhanes3_selected)

#write_csv(nhanes3_selected, "data/nhanes3_selected.csv")

```

### NHANES 4
```{r, eval = FALSE}

nhanes4_key <- read_csv("big_data/NHANES/nhanes_4/nhanes4_key.csv")

library(tidyverse)
library(haven)

# Assume nhanes4_key is loaded

# Separate keys by type of file (DEMO, HUQ) for simplicity
demo_key <- nhanes4_key %>% filter(str_detect(nhanes_file, "^DEMO"))
huq_key <- nhanes4_key %>% filter(str_detect(nhanes_file, "^HUQ"))

# A helper function to read and process a given domain of files
read_nhanes_domain <- function(key_table) {
  # Get unique files for this domain
  files <- key_table %>% distinct(nhanes_file)
  
  domain_data <- files %>%
    mutate(
      data = map(nhanes_file, ~ {
        vars_for_file <- key_table %>% filter(nhanes_file == .x)
        needed_vars <- c("SEQN", unique(vars_for_file$nhanes_var))
        
        file_path <- paste0("big_data/NHANES/nhanes_4/", .x, ".xpt")
        
        # Read and select needed variables
        df <- read_xpt(file_path) %>%
          select(any_of(needed_vars)) %>%
          # Rename nhanes_var to my_var
          rename_with(
            .fn = ~ vars_for_file$my_var[match(., vars_for_file$nhanes_var)],
            .cols = vars_for_file$nhanes_var
          ) %>%
          mutate(
            nhanes_yr_1 = vars_for_file$nhanes_yr_1[1],
            nhanes_yr_2 = vars_for_file$nhanes_yr_2[1]
          )
        
        df
      })
    ) %>%
    unnest(cols = data)  # Unnest after renaming done
  
  domain_data
}


# Read DEMO and HUQ data separately
demo_data <- read_nhanes_domain(demo_key)
huq_data <- read_nhanes_domain(huq_key)

# Now join demo and huq data by SEQN and cycle years.
# Note: If multiple cycles overlap, you may need to use both SEQN and nhanes_yr_1/nhanes_yr_2 as join keys.
# Typically SEQN is unique within a cycle, so joining on SEQN and year information might be prudent.
final_data <- demo_data %>%
  full_join(huq_data, by = c("SEQN", "nhanes_yr_1", "nhanes_yr_2"))

# Now select the columns you need:
final_data <- final_data %>%
  select(
    SEQN,
    age,
    srh_huq010,
    SDDSRVYR,
    nhanes_yr_1,
    nhanes_yr_2
  )

glimpse(final_data)

nhanes4_selected <- final_data %>%
  filter(srh_huq010 %in% 1:5) %>% 
  filter(age >= 18) %>% 
  mutate(srh = 6 - srh_huq010) %>% 
  mutate(year = (nhanes_yr_1 + nhanes_yr_2 ) / 2 ) %>% 
  mutate(cohort = year - age)

glimpse(nhanes4_selected)

write_csv(nhanes4_selected, "big_data/NHANES/nhanes_4/nhanes4_selected_apcsrh.csv")
write_csv(nhanes4_selected, "data/nhanes4_selected_apcsrh.csv")

```

## Import and check formatted data

```{r}

nhanes3 <- read_csv("data/nhanes3_selected.csv") %>% 
  select(SEQN, age, year, cohort, srh)

nhanes4 <- read_csv("data/nhanes4_selected_apcsrh.csv") %>% 
  select(SEQN, age, year, cohort, srh)

data_nhanes <- rbind(nhanes3, nhanes4) 

data_nhanes <- data_nhanes %>% 
  na.omit() %>% 
  filter(age >= 18)

glimpse(data_nhanes)

table(data_nhanes$srh)
table(data_nhanes$year)
table(data_nhanes$cohort)
hist(data_nhanes$age)
hist(data_nhanes$year)
hist(data_nhanes$cohort)
hist(data_nhanes$srh)

```

# NHANES III and IV

```{r}
data_nhanes %>%
    mutate(age = cut(age, breaks = 6)) %>% # Create cohorts with 6 breaks
    group_by(age, year) %>% 
    dplyr::summarize(mean_health = mean(srh, na.rm = TRUE), .groups = "drop") %>% 
    ggplot(aes(x = year, y = mean_health, color = age)) +
    geom_line() +
    geom_point() +
    theme_minimal() +
    labs(title = "Average SRH Per Year for Each Age Group",
         subtitle = "NHANES III and IV Datasets",
         x = "Year", 
         y = "Average SRH")
```



```{r}

data_nhanes %>% 
  group_by(age, year) %>% 
  summarize(mean_health = mean(srh)) %>% 
  ggplot(aes(x = age, y = mean_health)) +
  geom_line(color = "cornflowerblue") +
  facet_wrap(~ year) +
  labs(title = "Self-Rated Health By Age (Per Year)",
       subtitle = "NHANES III and IV Datasets",
       x = "Age of Respondent", 
       y = "Average SRH",
       )

```

```{r}

library(broom)

# Aggregate slopes

# Perform linear regression for each year and extract the coefficient of 'age' with confidence intervals, se, t stat, p val
lm_health_v_age_0 <- data_nhanes %>%
  group_by(year) %>%
  do(tidy(lm(srh ~ age, data = .), conf.int = TRUE)) %>%  # Add conf.int = TRUE for CIs
  filter(term == "age") %>%
  select(year, coef = estimate, conf.low, conf.high, se = std.error, t_statistic = statistic,  p_value = p.value)

# View the results with confidence intervals, se, t statistic, and p value
# print(lm_health_v_age_0)
knitr::kable(lm_health_v_age_0, 
             caption = "NHANES III and IV Datasets")

```

```{r}

# Plot coefficients
ggplot(lm_health_v_age_0, aes(x = year, y = coef)) +
  geom_point() +
  labs(
    title = "Change in 'Age' Coefficient Over Years",
    subtitle = "NHANES III and IV Datasets",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()

```


```{r}

## Regress the srh vs age coefficients from each year on the year of the survey


# Visualize
ggplot(lm_health_v_age_0, aes(x = year, y = coef)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +  # Adds the regression line with standard error shading
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +  # Confidence intervals for the coefficients
  labs(
    title = "Regression of 'Age' Coefficient Over Years",
    subtitle = "NHANES III and IV Datasets",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()

# Perform linear regression of 'coef' (age coefficient) vs 'year'
lm_coef_vs_year <- lm(coef ~ year, data = lm_health_v_age_0)

# View the summary of the regression
summary(lm_coef_vs_year)

```




```{r}
# Self-Rated Health Category Distribution
data_nhanes %>%
  mutate(health = srh) %>% 
  select(year, health) %>%
  filter(!is.na(health)) %>%
  mutate(
    srh = factor(health, 
                    levels = 1:5,
                    labels = c("Poor", "Fair", "Good", "Very Good", "Excellent"))) %>% 
  # Remove missing values
  # Calculate percentages by year
  group_by(year) %>%
  count(srh) %>%
  mutate(percent = n / sum(n) * 100) %>%
  ungroup() %>% 
  ggplot(aes(x = year, y = percent, color = srh)) +
  geom_line() +
  geom_point() +
  labs(title = "Self-Rated Health Category Distribution",
       subtitle = "GSS Dataset",
       y = "Percent",
       x = "Year") +
  theme_minimal()


# Self-Rated Health Category Distribution by Age Group
data_nhanes %>%
  mutate(age_group = cut(age, breaks = 6)) %>% # Create cohorts with 6 breaks
  mutate(health = srh) %>% 
  select(year, health, age_group) %>%
  filter(!is.na(health)) %>%
  mutate(
    srh = factor(health, 
                    levels = 1:5,
                    labels = c("Poor", "Fair", "Good", "Very Good", "Excellent"))) %>% 
  # Remove missing values
  # Calculate percentages by year
  group_by(year, age_group) %>%
  count(srh) %>%
  mutate(percent = n / sum(n) * 100) %>%
  ungroup() %>% 
  ggplot(aes(x = year, y = percent, color = srh)) +
  geom_line() +
  geom_point() +
  labs(title = "Self-Rated Health Category Distribution",
       subtitle = "GSS Dataset",
       y = "Percent",
       x = "Year") +
  facet_wrap(~age_group) +
  theme_minimal()


```


Maybe the different NHANES studies aren't comparable -- let's try only NHANES 4

# NHANES IV Only


```{r}

nhanes4 <- nhanes4 %>% 
  filter(age >= 18) %>% 
  na.omit()

nhanes4 %>%
    mutate(age = cut(age, breaks = 6)) %>% # Create cohorts with 6 breaks
    group_by(age, year) %>% 
    dplyr::summarize(mean_health = mean(srh, na.rm = TRUE), .groups = "drop") %>% 
    ggplot(aes(x = year, y = mean_health, color = age)) +
    geom_line() +
    geom_point() +
    labs(title = "Average SRH Per Year for Each Age Group",
         subtitle = "NHANES IV",
         x = "Year", 
         y = "Average SRH")

```


```{r}

nhanes4 %>% 
  group_by(age, year) %>% 
  summarize(mean_health = mean(srh)) %>% 
  ggplot(aes(x = age, y = mean_health)) +
  geom_line(color = "cornflowerblue") +
  geom_smooth() +
  facet_wrap(~ year) +
  labs(title = "Self-Rated Health By Age (Per Year)",
       subtitle = "NHANES IV Dataset",
       x = "Age of Respondent", 
       y = "Average SRH",
       )

nhanes4 %>% 
  group_by(age, year) %>% 
  summarize(mean_health = mean(srh)) %>% 
  ggplot(aes(x = age, y = mean_health)) +
  geom_line(color = "cornflowerblue") +
  geom_smooth() +
  facet_wrap(~ year) +
  labs(title = "Self-Rated Health By Age (Per Year)",
       subtitle = "NHANES IV Dataset",
       x = "Age of Respondent", 
       y = "Average SRH",
       )

nhanes4 %>% 
  group_by(age, year) %>% 
  summarize(mean_health = mean(srh)) %>% 
  ggplot(aes(x = age, y = mean_health)) +
  geom_line(color = "cornflowerblue") +
  geom_smooth(method = "lm") +
  facet_wrap(~ year) +
  labs(title = "Self-Rated Health By Age (Per Year)",
       subtitle = "NHANES IV Dataset",
       x = "Age of Respondent", 
       y = "Average SRH",
       )
```

```{r}

library(broom)

# Aggregate slopes

# Perform linear regression for each year and extract the coefficient of 'age' with confidence intervals, se, t stat, p val
lm_health_v_age_0 <- nhanes4 %>%
  group_by(year) %>%
  do(tidy(lm(srh ~ age, data = .), conf.int = TRUE)) %>%  # Add conf.int = TRUE for CIs
  filter(term == "age") %>%
  select(year, coef = estimate, conf.low, conf.high, se = std.error, t_statistic = statistic,  p_value = p.value)

# View the results with confidence intervals, se, t statistic, and p value
# print(lm_health_v_age_0)
knitr::kable(lm_health_v_age_0, 
             caption = "NHANES IV")

```

```{r}

# Plot coefficients
ggplot(lm_health_v_age_0, aes(x = year, y = coef)) +
  geom_point() +
  labs(
    title = "Change in 'Age' Coefficient Over Years",
    subtitle = "NHANES IV Dataset",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()

```


```{r}

## Regress the srh vs age coefficients from each year on the year of the survey


# Visualize
ggplot(lm_health_v_age_0, aes(x = year, y = coef)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +  # Adds the regression line with standard error shading
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +  # Confidence intervals for the coefficients
  labs(
    title = "Regression of 'Age' Coefficient Over Years",
    subtitle = "NHANES IV Dataset",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()

# Perform linear regression of 'coef' (age coefficient) vs 'year'
lm_coef_vs_year <- lm(coef ~ year, data = lm_health_v_age_0)

# View the summary of the regression
summary(lm_coef_vs_year)

```


```{r}

# Self-Rated Health Category Distribution
nhanes4 %>%
  mutate(health = srh) %>% 
  select(year, health) %>%
  filter(!is.na(health)) %>%
  mutate(
    srh = factor(health, 
                    levels = 1:5,
                    labels = c("Poor", "Fair", "Good", "Very Good", "Excellent"))) %>% 
  # Remove missing values
  # Calculate percentages by year
  group_by(year) %>%
  count(srh) %>%
  mutate(percent = n / sum(n) * 100) %>%
  ungroup() %>% 
  ggplot(aes(x = year, y = percent, color = srh)) +
  geom_line() +
  geom_point() +
  labs(title = "Self-Rated Health Category Distribution",
       subtitle = "GSS Dataset",
       y = "Percent",
       x = "Year") +
  theme_minimal()


# Self-Rated Health Category Distribution by Age Group
nhanes4 %>%
  mutate(age_group = cut(age, breaks = 6)) %>% # Create cohorts with 6 breaks
  mutate(health = srh) %>% 
  select(year, health, age_group) %>%
  filter(!is.na(health)) %>%
  mutate(
    srh = factor(health, 
                    levels = 1:5,
                    labels = c("Poor", "Fair", "Good", "Very Good", "Excellent"))) %>% 
  # Remove missing values
  # Calculate percentages by year
  group_by(year, age_group) %>%
  count(srh) %>%
  mutate(percent = n / sum(n) * 100) %>%
  ungroup() %>% 
  ggplot(aes(x = year, y = percent, color = srh)) +
  geom_line() +
  geom_point() +
  labs(title = "Self-Rated Health Category Distribution",
       subtitle = "GSS Dataset",
       y = "Percent",
       x = "Year") +
  facet_wrap(~age_group) +
  theme_minimal()

```

