---
title: "NHANES IV Weighted"
author: "Christine Lucille Kuryla"
date: "2024-12-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(broom)
library(srvyr)
library(here)
```



# Import, wrangle, and inspect data

### For reference, importing data
```{r, eval = FALSE}

nhanes4_key <- read_csv(here("big_data/NHANES/nhanes_4/nhanes4_key.csv")) %>% 
  filter(year != 2018.5)

library(tidyverse)
library(haven)

# Assume nhanes4_key is loaded

# Separate keys by type of file (DEMO, HUQ) for simplicity
demo_key <- nhanes4_key %>% filter(str_detect(nhanes_file, "^DEMO"))
huq_key <- nhanes4_key %>% filter(str_detect(nhanes_file, "^HUQ"))

# A helper function to read and process a given domain of files
read_nhanes_domain <- function(key_table) {
  # Get unique files for this domain
  files <- key_table %>% distinct(nhanes_file)
  
  domain_data <- files %>%
    mutate(
      data = map(nhanes_file, ~ {
        vars_for_file <- key_table %>% filter(nhanes_file == .x)
        needed_vars <- c("SEQN", unique(vars_for_file$nhanes_var))
        
        file_path <- here(paste0("big_data/NHANES/nhanes_4/", .x, ".xpt"))
        
        # Read and select needed variables
        df <- read_xpt(file_path) %>%
          select(any_of(needed_vars)) %>%
          # Rename nhanes_var to my_var
          rename_with(
            .fn = ~ vars_for_file$my_var[match(., vars_for_file$nhanes_var)],
            .cols = vars_for_file$nhanes_var
          ) %>%
          mutate(
            nhanes_yr_1 = vars_for_file$nhanes_yr_1[1],
            nhanes_yr_2 = vars_for_file$nhanes_yr_2[1]
          )
        
        df
      })
    ) %>%
    unnest(cols = data)  # Unnest after renaming done
  
  domain_data
}


# Read DEMO and HUQ data separately
demo_data <- read_nhanes_domain(demo_key)
huq_data <- read_nhanes_domain(huq_key)


### 2. DEMO Weight/Design data: SEQN, SDMVSTRA, SDMVPSU, WTINT2YR 

# These variables are standard survey design and weighting variables found in DEMO files.
# They are not part of `nhanes4_key` (or may not need to be), so we just extract them directly.
# We'll assume all DEMO files have these variables. If some do not, they'll be missing.

demo_weight_data <- demo_key %>% 
  distinct(nhanes_file) %>%   # One entry per DEMO file
  pull(nhanes_file) %>%
  map_df(~ {
    file_path <- here(paste0("big_data/NHANES/nhanes_4/", .x, ".xpt"))
    df <- read_xpt(file_path)
    df %>%
      select(SEQN, any_of(c("SDMVSTRA", "SDMVPSU", "WTINT2YR")))
  })

### Combine all three data frames by SEQN ###

# Start by joining DEMO age/survey data with weight/design data
demo_combined <- full_join(demo_data, demo_weight_data, by = "SEQN")

# Then join the HUQ data
final_data <- full_join(demo_combined, huq_data, by = c("SEQN","nhanes_yr_1", "nhanes_yr_2"))


glimpse(final_data)
table(final_data$nhanes_yr_1)
table(final_data$nhanes_yr_2)

nhanes4_selected <- final_data %>%
  filter(srh_huq010 %in% 1:5) %>% 
  filter(age >= 18) %>% 
  mutate(srh = 6 - srh_huq010) %>% 
  mutate(year = (nhanes_yr_1 + nhanes_yr_2 ) / 2 ) %>% 
  mutate(cohort = year - age) %>% 
  select(SEQN, srh, age, year, cohort, SDDSRVYR, SDMVSTRA, SDMVPSU, WTINT2YR)

table(nhanes4_selected$year)

glimpse(nhanes4_selected)
table(nhanes4_selected$year)

#write_csv(nhanes4_selected, "big_data/NHANES/nhanes_4/nhanes4_selected_apcsrh.csv")
#write_csv(nhanes4_selected, "data/nhanes4_selected_apcsrh.csv")

```

## Load and explore data

```{r}

data_nhanes <- read_csv(here("data/nhanes4_selected_apcsrh.csv")) %>% 
  mutate(age_group = as.factor( 
                            cut(
                            age,
                            breaks = c(17, 29, 39, 49, 59, 69, Inf),
                            labels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+"),
                            levels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+"),
                            right = TRUE
                          )),
         health_cat = factor(srh, 
                        levels = 1:5,
                        labels = c("Poor", "Fair", "Good", "Very Good", "Excellent")))

glimpse(data_nhanes)

data_nhanes_clk <- data_nhanes

# Other weights
# Gloria's Aim 2 Data
original_df <- read.csv(here("big_data/NHANES/Gloria_preprocessed/Data-3/Core_Dataset_Aim2.csv"))

gloria_wts <- original_df %>% 
  select(SEQN, WTMEC2YR)

# Using other weights
data_nhanes <- merge(data_nhanes_clk, gloria_wts, by = "SEQN")

glimpse(data_nhanes)


table(data_nhanes$year) # Survey years
table(data_nhanes$srh)
hist(data_nhanes$srh)
hist(data_nhanes$age)
table(data_nhanes$age_group)
hist(data_nhanes$year)
hist(data_nhanes$cohort)

```

### Incorporate weighting!

```{r}

library(survey)

# svy_nhanes <- svydesign(
#   id = ~SDMVPSU,
#   strata = ~SDMVSTRA,
#   weights = ~WTINT2YR,   # or WTINT2YR if you are analyzing interview-only data
#   nest = TRUE,
#   data = nhanes_data_4
# )

svy_nhanes <- data_nhanes %>%
  as_survey_design(
    ids = SDMVPSU,           # PSU identifiers (use 1 if not available)
    weights = WTINT2YR,  # original -- interview weights -- larger sample size but fewer covariates
  #  weights = WTMEC2YR, # Gloria's -- enable more covariates
    strata = SDMVSTRA,
    nest = TRUE
    )

```


# Main Analysis

# Mean SRH Over Time (by Age Group)

```{r}


svy_nhanes %>%
  mutate(srh = as.numeric(srh)) %>% 
  group_by(year) %>%
  summarize(
    mean_health = survey_mean(srh, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  ggplot(aes(x = year, y = mean_health)) +
  geom_line(size = 1, color = "cornflowerblue") +
  geom_point(size = 2, color = "darkblue") +
  labs(
    title = "Self-Rated Health",
    x = "Year",
    y = "Mean SRH",
    subtitle = "NHANES Dataset"
  ) +
  theme_minimal() 

svy_nhanes %>%
  filter(age >= 18) %>% 
  mutate(srh = as.numeric(srh)) %>% 
  group_by(age) %>%
  summarize(
    mean_health = survey_mean(srh, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  ggplot(aes(x = age, y = mean_health)) +
  geom_line(size = 1, color = "cornflowerblue") +
  geom_point(size = 2, color = "darkblue") +
  labs(
    title = "Self-Rated Health",
    x = "Age",
    y = "Mean SRH",
    subtitle = "NHANES Dataset"
  ) +
  theme_minimal() 

svy_nhanes %>%
  #mutate(age_group = cut(age, breaks = 6)) %>% 
  group_by(age_group, year) %>%
  summarise(
    mean_health = survey_mean(srh, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = year, y = mean_health, color = age_group)) +
  #  geom_smooth(alpha = 0.2) +
  geom_line() +
  geom_point() +
  labs(
    title = "Average SRH Per Year for Each Age Group",
    subtitle = "NHANES IV",
    x = "Year",
    y = "Average Self-Rated Health",
    color = "Age Group"
  ) +
  theme_minimal()


```

# Relationship of self-rated health to age, separated out by years

```{r}

svy_nhanes %>% 
  group_by(age, year) %>% 
  summarize(mean_health = survey_mean(srh)) %>% 
  ggplot(aes(x = age, y = mean_health)) +
  geom_line(color = "cornflowerblue") +
  facet_wrap(~ year) +
  labs(title = "Self-Rated Health By Age (Per Year)",
       subtitle = "NHANES IV Dataset",
       x = "Age of Respondent", 
       y = "Average SRH",
       )



```

## Regress self-rated health on age, for each year

Let's do a regression on each self-rated-health vs age, subsetted for each year (the plots on the faceted figure), look at the significance, and plot the coefficients for age with 95% CIs:

```{r}

# Perform weighted regression for each year
weighted_lm_by_year <- svy_nhanes %>%
  group_by(year) %>%
  group_map_dfr(~ {
    model <- survey::svyglm(srh ~ age, design = .x)
    tidy(model, conf.int = TRUE)
  }) %>%
  filter(term == "age") %>%
  select(year, estimate, std.error, conf.low, conf.high, statistic, p.value)

knitr::kable(weighted_lm_by_year)



```

## Regress the coefficients over years

```{r}

# Visualize
ggplot(weighted_lm_by_year, aes(x = year, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), width=.2,
                position=position_dodge(0.05)) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.3) +
  labs(
    title = "Regression of 'Age' Coefficient Over Years",
    subtitle = "NHANES IV Dataset",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()

# Perform linear regression of 'coef' (age coefficient) vs 'year'
lm_coef_vs_year <- lm(estimate ~ year, data = weighted_lm_by_year)

# View the summary of the regression
summary(lm_coef_vs_year)

```

```{r}

data_nhanes <- data_nhanes %>% 
  mutate(period_cut_6 = as.factor(cut(data_nhanes$year, 6)),
    period_cut_10 = as.factor(cut(data_nhanes$year, 10)),
    period_cut_12 = as.factor(cut(data_nhanes$year, 12)),
    period_groups = as.factor(cut(data_nhanes$year, 12)),
    period_10yr = as.factor(
                            cut(
                            year,
                            breaks = c(1973, 1982, 1990, 1998, 2006, 2014, Inf),
                            labels = c("1974-1982", "1982-1990", "1990-1998", 
                                       "1998-2006", "2006-2014", "2014-2022"),
                            right = TRUE
                            )
                          ),
    period_decade = as.factor(
                            cut(
                            year,
                            breaks = c(1973, 1979, 1989, 1999, 2009, 2019, Inf),
                            labels = c("1974-1979", "1980-1989", "1990-1999",
                                       "2000-2009", "2010-2019", "2020-2024"),
                            right = TRUE
                            )
                          ),
    age_group = as.factor( 
                            cut(
                            age,
                            breaks = c(17, 29, 39, 49, 59, 69, Inf),
                            labels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+"),
                            right = TRUE
                          )),
    age_groups = as.factor( 
                            cut(
                            age,
                            breaks = c(17, 29, 39, 49, 59, 69, Inf),
                            labels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+"),
                            right = TRUE
                          )),
    age_group_small = as.factor( 
                            cut(
                              age,
                              breaks = c(seq(15, 75, by = 5), Inf),  # Define breaks up to 75 and include Inf for the last group
                              labels = c("16-20", "21-25", "26-30", "31-35", "36-40", "41-45", "46-50", "51-55", "56-60", "61-65", "66-70", "71-75", "76+"),
                              right = FALSE  # Makes intervals left-closed, i.e., [x, y)
                            )
                          )
                          ,
    generation = factor(
      case_when(
        cohort >= 1901 & cohort <= 1927 ~ "Greatest (1901-1927)",
        cohort >= 1928 & cohort <= 1945 ~ "Silent (1928-1945)",
        cohort >= 1946 & cohort <= 1964 ~ "Boomers (1946-1964)",
        cohort >= 1965 & cohort <= 1980 ~ "Gen X (1965-1980)",
        cohort >= 1981 & cohort <= 1996 ~ "Millennials (1981-1996)",
        cohort >= 1997 & cohort <= 2012 ~ "Gen Z (1997-2012)",
        TRUE ~ "Other"
      ),
      levels = c(
        "Greatest (1901-1927)",
        "Silent (1928-1945)",
        "Boomers (1946-1964)",
        "Gen X (1965-1980)",
        "Millennials (1981-1996)",
        "Gen Z (1997-2012)"#,
     #   "Other"
      )
    ),
    generation_two_sections = factor(
      case_when(
        generation == "Greatest (1901-1927)" & cohort <= 1914 ~ "Greatest Early (1901-1914)",
        generation == "Greatest (1901-1927)" & cohort > 1914 ~ "Greatest Late (1915-1927)",
        generation == "Silent (1928-1945)" & cohort <= 1936 ~ "Silent Early (1928-1936)",
        generation == "Silent (1928-1945)" & cohort > 1936 ~ "Silent Late (1937-1945)",
        generation == "Boomers (1946-1964)" & cohort <= 1955 ~ "Boomers Early (1946-1955)",
        generation == "Boomers (1946-1964)" & cohort > 1955 ~ "Boomers Late (1956-1964)",
        generation == "Gen X (1965-1980)" & cohort <= 1972 ~ "Gen X Early (1965-1972)",
        generation == "Gen X (1965-1980)" & cohort > 1972 ~ "Gen X Late (1973-1980)",
        generation == "Millennials (1981-1996)" & cohort <= 1988 ~ "Millennials Early (1981-1988)",
        generation == "Millennials (1981-1996)" & cohort > 1988 ~ "Millennials Late (1989-1996)",
        generation == "Gen Z (1997-2012)" & cohort <= 2004 ~ "Gen Z Early (1997-2004)",
        generation == "Gen Z (1997-2012)" & cohort > 2004 ~ "Gen Z Late (2005-2012)",
        TRUE ~ "Other"
      ),
      levels = c(
        "Greatest Early (1901-1914)", "Greatest Late (1915-1927)",
        "Silent Early (1928-1936)", "Silent Late (1937-1945)",
        "Boomers Early (1946-1955)", "Boomers Late (1956-1964)",
        "Gen X Early (1965-1972)", "Gen X Late (1973-1980)",
        "Millennials Early (1981-1988)", "Millennials Late (1989-1996)",
        "Gen Z Early (1997-2004)", "Gen Z Late (2005-2012)"#,
     #   "Other"
      )
    ),
    generation_three_sections = factor(
      case_when(
        generation == "Greatest (1901-1927)" & cohort <= 1910 ~ "Greatest Early (1901-1910)",
        generation == "Greatest (1901-1927)" & cohort > 1910 & cohort <= 1918 ~ "Greatest Mid (1911-1918)",
        generation == "Greatest (1901-1927)" & cohort > 1918 ~ "Greatest Late (1919-1927)",
        generation == "Silent (1928-1945)" & cohort <= 1934 ~ "Silent Early (1928-1934)",
        generation == "Silent (1928-1945)" & cohort > 1934 & cohort <= 1940 ~ "Silent Mid (1935-1940)",
        generation == "Silent (1928-1945)" & cohort > 1940 ~ "Silent Late (1941-1945)",
        generation == "Boomers (1946-1964)" & cohort <= 1951 ~ "Boomers Early (1946-1951)",
        generation == "Boomers (1946-1964)" & cohort > 1951 & cohort <= 1958 ~ "Boomers Mid (1952-1958)",
        generation == "Boomers (1946-1964)" & cohort > 1958 ~ "Boomers Late (1959-1964)",
        generation == "Gen X (1965-1980)" & cohort <= 1970 ~ "Gen X Early (1965-1970)",
        generation == "Gen X (1965-1980)" & cohort > 1970 & cohort <= 1976 ~ "Gen X Mid (1971-1976)",
        generation == "Gen X (1965-1980)" & cohort > 1976 ~ "Gen X Late (1977-1980)",
        generation == "Millennials (1981-1996)" & cohort <= 1986 ~ "Millennials Early (1981-1986)",
        generation == "Millennials (1981-1996)" & cohort > 1986 & cohort <= 1992 ~ "Millennials Mid (1987-1992)",
        generation == "Millennials (1981-1996)" & cohort > 1992 ~ "Millennials Late / Gen Z (1993-2004)",
    #    generation == "Gen Z (1997-2012)" & cohort <= 2002 ~ "Gen Z Early (1997-2002)",
    #    generation == "Gen Z (1997-2012)" & cohort > 2002 & cohort <= 2008 ~ "Gen Z Mid (2003-2008)",
    #    generation == "Gen Z (1997-2012)" & cohort > 2008 ~ "Gen Z Late (2009-2012)",
        TRUE ~ "Other"
      ),
      levels = c(
        "Greatest Early (1901-1910)", "Greatest Mid (1911-1918)", "Greatest Late (1919-1927)",
        "Silent Early (1928-1934)", "Silent Mid (1935-1940)", "Silent Late (1941-1945)",
        "Boomers Early (1946-1951)", "Boomers Mid (1952-1958)", "Boomers Late (1959-1964)",
        "Gen X Early (1965-1970)", "Gen X Mid (1971-1976)", "Gen X Late (1977-1980)",
        "Millennials Early (1981-1986)", "Millennials Mid (1987-1992)", 
        "Millennials Late / Gen Z (1993-2004)"
        #"Millennials Late (1993-1996)",
      #  "Gen Z Early (1997-2002)", "Gen Z Mid (2003-2008)", "Gen Z Late (2009-2012)" #,
      #  "Other"
      )
    ))
  
svy_nhanes <- data_nhanes %>%
  as_survey_design(
    ids = SDMVPSU,           # PSU identifiers (use 1 if not available)
  #  weights = WTINT2YR,  # original -- interview weights -- larger sample size but fewer covariates
    weights = WTMEC2YR, # Gloria's -- enable more covariates
    strata = SDMVSTRA,
    nest = TRUE
    )



twofactor_agecohort_m =
  svyglm(srh ~ age_4yr + cohort_4yr, design = svy_nhanes)
summary(twofactor_agecohort_m)

threefactor_m =
  lmer(srh ~ age_4yr + race + gender + log(WTMEC4YR) + (1|period_4yr) + (1|cohort_4yr) + (1|SDMVSTRA/SDMVPSU), data = data_nhanes, REML = FALSE)
summary(threefactor_m)


clk_lmer =
  lmer(srh ~ age_4yr + (1|period_4yr) + (1|cohort_4yr) + log(WTMEC4YR) + (1|SDMVSTRA/SDMVPSU), data = data_nhanes, REML = FALSE)
summary(clk_lmer)

clk_lmer2 =
  lmer(srh ~ (1|age_4yr) + (1|period_4yr) + generation_two_sections + (1|SDMVSTRA/SDMVPSU), data = data_nhanes, REML = FALSE)
summary(clk_lmer2)

clk_lmer3 =
  lmer(srh ~ (1|age_4yr) + generation_two_sections + (1|SDMVSTRA/SDMVPSU), data = data_nhanes, REML = FALSE)
summary(clk_lmer3)

```


# Cohort effects figure

```{r}

library(broom)

generation_age_interaction_figure <- function(cohort_length_string) {
  
  # Prepare data for the plot
  coef_data <- broom::tidy(weighted_lm_generation) %>%
    filter(term != "(Intercept)") %>%
    mutate(
      # Clean the term to remove "cohort_group" prefix if present
      term_cleaned = ifelse(grepl("cohort_group", term), gsub("cohort_group", "", term), term),
      # Extract the first year from cohort terms using regex (assign NA for "age")
      first_year = ifelse(term_cleaned == "age", NA, as.numeric(stringr::str_extract(term_cleaned, "^\\d{4}"))),
      # Add a numeric ordering key: assign 0 to "age" and use the extracted first year for cohorts
      term_numeric = ifelse(term_cleaned == "age", 0, first_year),
      # Reorder terms based on the numeric ordering key
      term_cleaned = factor(term_cleaned, levels = term_cleaned[order(term_numeric)]),
      significant = ifelse(p.value < 0.05, "Significant", "Not Significant")
    )

  # Plot
  ggplot(coef_data, aes(x = term_cleaned, y = estimate, fill = significant)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.2) +
    labs(
      title = "Cohort Effects on SRH",
      subtitle = cohort_length_string,
      x = "Term (Age or Cohort Birthyear Group)",
      y = "Coefficient Estimate",
      fill = "Significance"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}


weighted_lm_generation <- svy_nhanes %>% 
  mutate(cohort_group = generation) %>% 
  svyglm(srh ~ age + cohort_group, design = .)
generation_age_interaction_figure("NHANES")

weighted_lm_generation <- svy_nhanes %>% 
  mutate(cohort_group = generation_two_sections) %>% 
  svyglm(srh ~ age + cohort_group, design = .)
generation_age_interaction_figure("NHANES")

weighted_lm_generation <- svy_nhanes %>% 
  mutate(cohort_group = generation_three_sections) %>% 
  svyglm(srh ~ age + cohort_group, design = .)
generation_age_interaction_figure("NHANES")

```


```{r}

data_nhanes %>%
  select(year, health_cat, age_group) %>%
  na.omit() %>% 
  mutate(srh = health_cat) %>% 
#  filter(!is.na(health)) %>%
#  mutate(
#    srh = factor(health, 
#                 levels = 1:4,
#                 labels = c("Poor", "Fair", "Good", "Excellent"))) %>% 
  # Remove missing values
  # Calculate percentages by year
  group_by(year, age_group) %>%
  count(srh) %>%
  mutate(percent = n / sum(n) * 100) %>%
  ungroup() %>% 
  ggplot(aes(x = year, y = percent, color = srh)) +
  geom_line() +
  facet_wrap(~age_group) +
  labs(title = "Self-Rated Health Category Distribution by Age Group",
       subtitle = "NHANES IV Dataset",
       y = "Percent",
       x = "Year") +
  theme_minimal()

```

