---
title: "SRH - GSS with Survey Weights"
author: "Christine Lucille Kuryla"
date: "2024-11-29"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(srvyr)
library(survey)
#library(Hmisc)
library(broom)
library(tidyverse)

library(conflicted)
conflicts_prefer(dplyr::summarise)
conflicts_prefer(dplyr::summarize)
conflicts_prefer(dplyr::select)
conflicts_prefer(dplyr::mutate)
conflicts_prefer(dplyr::rename)
conflicts_prefer(dplyr::filter)
```


Here's a summary of the interesting findings from my analysis of self-rated health in the GSS dataset so far. 

See https://github.com/clkuryla/self_rated_health/blob/main/gss_eda.md for more EDA and details.

* `health`
The first variable of interest is “health”, which will be the main subject of our analysis: https://gssdataexplorer.norc.org/variables/437/vshow

Question on survey: “Would you say your own health, in general, is excellent, good, fair, or poor?” 

Coded in this analysis (it was recoded from the raw data) as:
4 - Excellent
3 - Good
2 - Fair
1 - Poor

Other variables used are:

* `age`
  - Age of respondent at time of survey
  
* `year`
  - Year of survey
  
* `cohort`
  - Birth year of respondent
  
Additional covariates and analyses to come. 


# Fetch, load, clean, and recode data
## Fetch GSS data 

```{r fetch_data, eval = FALSE}

# Feel free to modify to play with more covariates and variables.

#install.packages('gssr', repos =  c('https://kjhealy.r-universe.dev', 'https://cloud.r-project.org'))
# install.packages('gssrdoc', repos = c('https://kjhealy.r-universe.dev', 'https://cloud.r-project.org'))

library(gssr)
library(gssrdoc)

data("gss_all") # this file is big! 

# It's a bit excessive to download the entire GSS dataset every time we knit, so lets just save some variables of interest and write it out for future use.

# Add 'wtss' and 'wtssall' to variable selection
data_gss <- as.data.frame(gss_all) %>% 
  select(
    year,      # Year of survey
    cohort,    # Birth year
    age,       # Age at time of survey
    health,    # Self-rated health
    sex,       # Sex
    happy,     # Self-rated happiness
    life,      # Is life exciting or dull (subjective wellbeing)
    educ,      # Years of education
    polviews,  # Political views
    class,     # Subjective class identification
    satfin,    # Satisfaction with financial situation
    region,    # region (south: 1 if 4<x<8, else 0)
    attend,    # attends religious services
    race,      # race
    region,
    wrkstat,
    hispanic,
    degree,
    marital,
    partyid,
    attend,    # attended college
  #  vpsu,         # not available in gssr
  #  vstrat,       # not available in gssr
    wtssps,    # Weight for single-year analysis
    wtssall,    # Weight for multi-year analysis ** stops at 2018
    wtsscomp
  )

write_csv(data_gss, "data/extracted_gss_variables.csv")

# Note: GSSR does not provide vpsu and vstrat. Other data sources may provide it, and we will look into that eventually. 

# If found, update survey object to:
# gss_svy <- data_gss %>%
#   as_survey_design(
#     ids = vpsu,           # PSU identifiers
#     strata = vstrat,      # Stratification variable
#     weights = wtssall,    # Weight variable for multi-year analysis
#     nest = TRUE           # Important when PSUs are nested within strata
#   )

# instead of:
# gss_svy <- data_gss %>%
#   as_survey_design(
#     ids = 1,            # No clustering variable available
#     weights = wtssall
#   )
```

## Load and clean data

Here we'll load our data, clean some unwanted values, and recode the unintuitive variables.

```{r load_and_clean_data}
data_gss <- read_csv("data/extracted_gss_variables.csv") %>% 
  filter(cohort != 9999) %>% 
#  select(-c(wtssps)) %>% 
  # select(-c("region", 
  #        #   "marital", 
  #        #   "partyid", 
  #           "hispanic"
  #           )) %>% 
#  select(-wtssall) %>% 
  filter(!is.na(health)) %>% 
#  na.omit() %>% 
#  mutate(south = if_else(region %in% c(4, 5, 6, 7, 8), "South", "Not South")) %>% 
#  mutate(south = as_factor(south)) %>% 
#  mutate(region = as_factor(region)) %>% 
#  mutate(wrkstat = as_factor(wrkstat)) %>% 
#  mutate(hispanic = as_factor(hispanic)) %>% 
##  mutate(marital = as_factor(marital)) %>% 
#  mutate(partyid = as_factor(partyid)) %>% 
  mutate(health = 5 - health)  %>%  # reverse the coding so it's more intuitive (higher number for excellent, lower number for poor)
  mutate(happy = 4 - happy) %>% # same
  mutate(life = 4 - life) %>% # reverse again, these variables tend to be unintuitively ordered!!!
  mutate(satfin = 4 - satfin) %>% # same again! %>% 
 mutate(
    age_group = cut(
      age,
      breaks = c(17, 29, 39, 49, 59, 69, Inf),
      labels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+"),
      right = TRUE
    )
  ) 

table(data_gss$age_group)

```


```{r}

# more categories

data_gss <- data_gss %>% 
    mutate(
    health_cat = factor(health, 
                        levels = 1:4,
                        labels = c("Poor", "Fair", "Good", "Excellent")),
    period_cut_6 = as.factor(cut(data_gss$year, 6)),
    period_cut_10 = as.factor(cut(data_gss$year, 10)),
    period_cut_12 = as.factor(cut(data_gss$year, 12)),
    period_groups = as.factor(cut(data_gss$year, 12)),
    period_10yr = as.factor(
                            cut(
                            year,
                            breaks = c(1973, 1982, 1990, 1998, 2006, 2014, Inf),
                            labels = c("1974-1982", "1982-1990", "1990-1998", 
                                       "1998-2006", "2006-2014", "2014-2022"),
                            right = TRUE
                            )
                          ),
    period_decade = as.factor(
                            cut(
                            year,
                            breaks = c(1973, 1979, 1989, 1999, 2009, 2019, Inf),
                            labels = c("1974-1979", "1980-1989", "1990-1999",
                                       "2000-2009", "2010-2019", "2020-2024"),
                            right = TRUE
                            )
                          ),
    age_group = as.factor( 
                            cut(
                            age,
                            breaks = c(17, 29, 39, 49, 59, 69, Inf),
                            labels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+"),
                            right = TRUE
                          )),
    age_groups = as.factor( 
                            cut(
                            age,
                            breaks = c(17, 29, 39, 49, 59, 69, Inf),
                            labels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+"),
                            right = TRUE
                          )),
    age_group_small = as.factor( 
                            cut(
                              age,
                              breaks = c(seq(15, 75, by = 5), Inf),  # Define breaks up to 75 and include Inf for the last group
                              labels = c("16-20", "21-25", "26-30", "31-35", "36-40", "41-45", "46-50", "51-55", "56-60", "61-65", "66-70", "71-75", "76+"),
                              right = FALSE  # Makes intervals left-closed, i.e., [x, y)
                            )
                          )
                          ,
    generation = factor(
      case_when(
        cohort >= 1901 & cohort <= 1927 ~ "Greatest (1901-1927)",
        cohort >= 1928 & cohort <= 1945 ~ "Silent (1928-1945)",
        cohort >= 1946 & cohort <= 1964 ~ "Boomers (1946-1964)",
        cohort >= 1965 & cohort <= 1980 ~ "Gen X (1965-1980)",
        cohort >= 1981 & cohort <= 1996 ~ "Millennials (1981-1996)",
        cohort >= 1997 & cohort <= 2012 ~ "Gen Z (1997-2012)",
        TRUE ~ "Other"
      ),
      levels = c(
        "Greatest (1901-1927)",
        "Silent (1928-1945)",
        "Boomers (1946-1964)",
        "Gen X (1965-1980)",
        "Millennials (1981-1996)",
        "Gen Z (1997-2012)"#,
     #   "Other"
      )
    ),
    generation_two_sections = factor(
      case_when(
        generation == "Greatest (1901-1927)" & cohort <= 1914 ~ "Greatest Early (1901-1914)",
        generation == "Greatest (1901-1927)" & cohort > 1914 ~ "Greatest Late (1915-1927)",
        generation == "Silent (1928-1945)" & cohort <= 1936 ~ "Silent Early (1928-1936)",
        generation == "Silent (1928-1945)" & cohort > 1936 ~ "Silent Late (1937-1945)",
        generation == "Boomers (1946-1964)" & cohort <= 1955 ~ "Boomers Early (1946-1955)",
        generation == "Boomers (1946-1964)" & cohort > 1955 ~ "Boomers Late (1956-1964)",
        generation == "Gen X (1965-1980)" & cohort <= 1972 ~ "Gen X Early (1965-1972)",
        generation == "Gen X (1965-1980)" & cohort > 1972 ~ "Gen X Late (1973-1980)",
        generation == "Millennials (1981-1996)" & cohort <= 1988 ~ "Millennials Early (1981-1988)",
        generation == "Millennials (1981-1996)" & cohort > 1988 ~ "Millennials Late (1989-1996)",
        generation == "Gen Z (1997-2012)" & cohort <= 2004 ~ "Gen Z Early (1997-2004)",
        generation == "Gen Z (1997-2012)" & cohort > 2004 ~ "Gen Z Late (2005-2012)",
        TRUE ~ "Other"
      ),
      levels = c(
        "Greatest Early (1901-1914)", "Greatest Late (1915-1927)",
        "Silent Early (1928-1936)", "Silent Late (1937-1945)",
        "Boomers Early (1946-1955)", "Boomers Late (1956-1964)",
        "Gen X Early (1965-1972)", "Gen X Late (1973-1980)",
        "Millennials Early (1981-1988)", "Millennials Late (1989-1996)",
        "Gen Z Early (1997-2004)", "Gen Z Late (2005-2012)"#,
     #   "Other"
      )
    ),
    generation_three_sections = factor(
      case_when(
        generation == "Greatest (1901-1927)" & cohort <= 1910 ~ "Greatest Early (1901-1910)",
        generation == "Greatest (1901-1927)" & cohort > 1910 & cohort <= 1918 ~ "Greatest Mid (1911-1918)",
        generation == "Greatest (1901-1927)" & cohort > 1918 ~ "Greatest Late (1919-1927)",
        generation == "Silent (1928-1945)" & cohort <= 1934 ~ "Silent Early (1928-1934)",
        generation == "Silent (1928-1945)" & cohort > 1934 & cohort <= 1940 ~ "Silent Mid (1935-1940)",
        generation == "Silent (1928-1945)" & cohort > 1940 ~ "Silent Late (1941-1945)",
        generation == "Boomers (1946-1964)" & cohort <= 1951 ~ "Boomers Early (1946-1951)",
        generation == "Boomers (1946-1964)" & cohort > 1951 & cohort <= 1958 ~ "Boomers Mid (1952-1958)",
        generation == "Boomers (1946-1964)" & cohort > 1958 ~ "Boomers Late (1959-1964)",
        generation == "Gen X (1965-1980)" & cohort <= 1970 ~ "Gen X Early (1965-1970)",
        generation == "Gen X (1965-1980)" & cohort > 1970 & cohort <= 1976 ~ "Gen X Mid (1971-1976)",
        generation == "Gen X (1965-1980)" & cohort > 1976 ~ "Gen X Late (1977-1980)",
        generation == "Millennials (1981-1996)" & cohort <= 1986 ~ "Millennials Early (1981-1986)",
        generation == "Millennials (1981-1996)" & cohort > 1986 & cohort <= 1992 ~ "Millennials Mid (1987-1992)",
        generation == "Millennials (1981-1996)" & cohort > 1992 ~ "Millennials Late / Gen Z (1993-2004)",
    #    generation == "Gen Z (1997-2012)" & cohort <= 2002 ~ "Gen Z Early (1997-2002)",
    #    generation == "Gen Z (1997-2012)" & cohort > 2002 & cohort <= 2008 ~ "Gen Z Mid (2003-2008)",
    #    generation == "Gen Z (1997-2012)" & cohort > 2008 ~ "Gen Z Late (2009-2012)",
        TRUE ~ "Other"
      ),
      levels = c(
        "Greatest Early (1901-1910)", "Greatest Mid (1911-1918)", "Greatest Late (1919-1927)",
        "Silent Early (1928-1934)", "Silent Mid (1935-1940)", "Silent Late (1941-1945)",
        "Boomers Early (1946-1951)", "Boomers Mid (1952-1958)", "Boomers Late (1959-1964)",
        "Gen X Early (1965-1970)", "Gen X Mid (1971-1976)", "Gen X Late (1977-1980)",
        "Millennials Early (1981-1986)", "Millennials Mid (1987-1992)", 
        "Millennials Late / Gen Z (1993-2004)"
        #"Millennials Late (1993-1996)",
      #  "Gen Z Early (1997-2002)", "Gen Z Mid (2003-2008)", "Gen Z Late (2009-2012)" #,
      #  "Other"
      )
    )
  )

```


```{r}
# Create a survey design object using wtssall for multi-year analysis
gss_svy <- data_gss %>%
  as_survey_design(
    ids = 1,           # PSU identifiers (use 1 if not available)
    weights = wtsscomp  # wtssall pre 2018, wtsscomp combined //Use 'wtss' for single-year analysis
  )
```

# Self rated health as predicted by age over the years

Let's explore the effect of different cohorts on SRH at certain ages.

## SRH vs year of survey for different ages

In the following figure, I cut the age of participants into 6 groups and plotted the mean of the group's self-rated health for each year (that's what each dot is). As you can qualitatively see, the spread seems to narrow. 

```{r}
# Compute weighted mean of self-rated health
gss_svy %>%
  summarise(
    mean_health = survey_mean(health, na.rm = TRUE)
  )

# # Create age groups
# gss_svy <- gss_svy %>%
#   mutate(
#     age_group = cut(
#       age,
#       breaks = c(17, 25, 34, 44, 55, 65, Inf),
#       labels = c("18-25", "26-34", "35-44", "45-55", "56-65", "66+"),
#       right = TRUE
#     )
#   ) 
# #  mutate(age_group = cut(age, breaks = 6))

# Compute weighted mean health by age group and year
gss_svy %>%
  group_by(age_group, year) %>%
  summarise(
    mean_health = survey_mean(health, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = year, y = mean_health, color = age_group)) +
#  geom_smooth(alpha = 0.2) +
  geom_line() +
  geom_point() +
  labs(
    title = "Average SRH Per Year for Each Age Group",
    subtitle = "GSS Dataset with Survey Weights",
    x = "Year",
    y = "Average Self-Rated Health",
    color = "Age Group"
  ) +
  theme_minimal()

library(broom)
```


## Relationship of self-rated health to age, separated out by years

Well, it seems like the spread of self-rated health among ages decreases as time goes on (later years). Let's look at that by faceting mean self-rated health vs age by year. 

Note that intuitively, we'd expect it to be a negative slope because older people intuitively should have worse health. 

Notice *the slopes seem to flatten over time.*

```{r health_v_age_per_year}
# health vs age per year
gss_svy %>% 
  group_by(age, year) %>% 
  summarize(mean_health = survey_mean(health)) %>% 
  ggplot(aes(x = age, y = mean_health)) +
  geom_line(color = "cornflowerblue") +
  facet_wrap(~ year) +
  labs(title = "Self-Rated Health By Age (Per Year)",
       subtitle = "GSS Dataset",
       x = "Age of Respondent", 
       y = "Average SRH",
       )
```



## Regress self-rated health on age, for each year

Let's do a regression on each self-rated-health vs age, subsetted for each year (the plots on the faceted figure), look at the significance, and plot the coefficients for age with 95% CIs:

```{r}
# Perform weighted regression for each year
weighted_lm_by_year <- gss_svy %>%
  group_by(year) %>%
  group_map_dfr(~ {
    model <- survey::svyglm(health ~ age, design = .x)
    tidy(model, conf.int = TRUE)
  }) %>%
  filter(term == "age") %>%
  select(year, estimate, std.error, conf.low, conf.high, statistic, p.value)

knitr::kable(weighted_lm_by_year)

summary(weighted_lm_by_year)

# # with additinal covariates
# weighted_lm_by_year <- gss_svy %>%
#   group_by(year) %>%
#   group_map_dfr(~ {
#     model <- survey::svyglm(health ~ age + sex + educ + race + happy + class, design = .x)
#     tidy(model, conf.int = TRUE)
#   }) %>%
#   filter(term == "age") %>%
#   select(year, estimate, std.error, conf.low, conf.high, statistic, p.value)
# 
# summary(weighted_lm_by_year)

```



```{r}

# Plot the coefficients with error bars
ggplot(weighted_lm_by_year, aes(x = year, y = estimate)) +
#  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), width=.2,
                 position=position_dodge(0.05)) +
 # geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +
  labs(
    title = "Change in 'Age' Coefficient Over Years",
    subtitle = "GSS Dataset",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()

# Regress the age coefficients on year
coef_model <- lm(estimate ~ year, data = weighted_lm_by_year)
summary(coef_model)

# Plot the regression
ggplot(weighted_lm_by_year, aes(x = year, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), width=.2,
                 position=position_dodge(0.05)) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.3) +
  labs(
    title = "Regression of 'Age' Coefficient Over Years",
    subtitle = "GSS Dataset",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()

weighted_lm_by_year
summary(weighted_lm_by_year)

# Perform linear regression of 'coef' (age coefficient) vs 'year'
lm_coef_vs_year <- lm(estimate ~ year, data = weighted_lm_by_year)

# View the summary of the regression
summary(lm_coef_vs_year)

```


```{r}

srh_age_groups_point_line_figure <- function(age_cuts) {
# Create age groups
gss_svy %>%
  mutate(age_group = cut(age, breaks = age_cuts)) %>% # number of age groups
  group_by(age_group, year) %>% # Compute weighted mean health by age group and year
  summarise(
    mean_health = survey_mean(health, na.rm = TRUE)
  ) %>% 
ggplot(aes(x = year, y = mean_health, color = age_group)) +
#  geom_smooth(alpha = 0.2) +
  geom_line() +
  geom_point() +
  labs(
    title = "Average SRH Per Year for Each Age Group",
    subtitle = "GSS Dataset with Survey Weights",
    x = "Year",
    y = "Average Self-Rated Health",
    color = "Age Group"
  ) +
  theme_minimal()
}

srh_age_groups_loess_figure <- function(age_cuts) {
# Create age groups
gss_svy %>%
  mutate(age_group = cut(age, breaks = age_cuts)) %>% # number of age groups
  group_by(age_group, year) %>% # Compute weighted mean health by age group and year
  summarise(
    mean_health = survey_mean(health, na.rm = TRUE)
  ) %>% 
ggplot(aes(x = year, y = mean_health, color = age_group)) +
  geom_smooth(alpha = 0.4) +
#  geom_line() +
#  geom_point() +
  labs(
    title = "Average SRH Per Year for Each Age Group",
    subtitle = "GSS Dataset with Survey Weights",
    x = "Year",
    y = "Average Self-Rated Health",
    color = "Age Group"
  ) +
  theme_minimal()
}

srh_age_groups_point_line_figure(3)
srh_age_groups_point_line_figure(4)
srh_age_groups_point_line_figure(6)
srh_age_groups_point_line_figure(10)

srh_age_groups_loess_figure(3)
srh_age_groups_loess_figure(4)
srh_age_groups_loess_figure(6)
srh_age_groups_loess_figure(10)

```

# Cohort effect on SRH

## Generation splitting

```{r}

data_gss_generations <- data_gss %>%
  filter(cohort > 1900) %>%
  mutate(
    generation = factor(
      case_when(
        cohort >= 1901 & cohort <= 1927 ~ "Greatest (1901-1927)",
        cohort >= 1928 & cohort <= 1945 ~ "Silent (1928-1945)",
        cohort >= 1946 & cohort <= 1964 ~ "Boomers (1946-1964)",
        cohort >= 1965 & cohort <= 1980 ~ "Gen X (1965-1980)",
        cohort >= 1981 & cohort <= 1996 ~ "Millennials (1981-1996)",
        cohort >= 1997 & cohort <= 2012 ~ "Gen Z (1997-2012)",
        TRUE ~ "Other"
      ),
      levels = c(
        "Greatest (1901-1927)",
        "Silent (1928-1945)",
        "Boomers (1946-1964)",
        "Gen X (1965-1980)",
        "Millennials (1981-1996)",
        "Gen Z (1997-2012)"#,
     #   "Other"
      )
    ),
    generation_two_sections = factor(
      case_when(
        generation == "Greatest (1901-1927)" & cohort <= 1914 ~ "Greatest Early (1901-1914)",
        generation == "Greatest (1901-1927)" & cohort > 1914 ~ "Greatest Late (1915-1927)",
        generation == "Silent (1928-1945)" & cohort <= 1936 ~ "Silent Early (1928-1936)",
        generation == "Silent (1928-1945)" & cohort > 1936 ~ "Silent Late (1937-1945)",
        generation == "Boomers (1946-1964)" & cohort <= 1955 ~ "Boomers Early (1946-1955)",
        generation == "Boomers (1946-1964)" & cohort > 1955 ~ "Boomers Late (1956-1964)",
        generation == "Gen X (1965-1980)" & cohort <= 1972 ~ "Gen X Early (1965-1972)",
        generation == "Gen X (1965-1980)" & cohort > 1972 ~ "Gen X Late (1973-1980)",
        generation == "Millennials (1981-1996)" & cohort <= 1988 ~ "Millennials Early (1981-1988)",
        generation == "Millennials (1981-1996)" & cohort > 1988 ~ "Millennials Late (1989-1996)",
        generation == "Gen Z (1997-2012)" & cohort <= 2004 ~ "Gen Z Early (1997-2004)",
        generation == "Gen Z (1997-2012)" & cohort > 2004 ~ "Gen Z Late (2005-2012)",
        TRUE ~ "Other"
      ),
      levels = c(
        "Greatest Early (1901-1914)", "Greatest Late (1915-1927)",
        "Silent Early (1928-1936)", "Silent Late (1937-1945)",
        "Boomers Early (1946-1955)", "Boomers Late (1956-1964)",
        "Gen X Early (1965-1972)", "Gen X Late (1973-1980)",
        "Millennials Early (1981-1988)", "Millennials Late (1989-1996)",
        "Gen Z Early (1997-2004)", "Gen Z Late (2005-2012)"#,
     #   "Other"
      )
    ),
    generation_three_sections = factor(
      case_when(
        generation == "Greatest (1901-1927)" & cohort <= 1910 ~ "Greatest Early (1901-1910)",
        generation == "Greatest (1901-1927)" & cohort > 1910 & cohort <= 1918 ~ "Greatest Mid (1911-1918)",
        generation == "Greatest (1901-1927)" & cohort > 1918 ~ "Greatest Late (1919-1927)",
        generation == "Silent (1928-1945)" & cohort <= 1934 ~ "Silent Early (1928-1934)",
        generation == "Silent (1928-1945)" & cohort > 1934 & cohort <= 1940 ~ "Silent Mid (1935-1940)",
        generation == "Silent (1928-1945)" & cohort > 1940 ~ "Silent Late (1941-1945)",
        generation == "Boomers (1946-1964)" & cohort <= 1951 ~ "Boomers Early (1946-1951)",
        generation == "Boomers (1946-1964)" & cohort > 1951 & cohort <= 1958 ~ "Boomers Mid (1952-1958)",
        generation == "Boomers (1946-1964)" & cohort > 1958 ~ "Boomers Late (1959-1964)",
        generation == "Gen X (1965-1980)" & cohort <= 1970 ~ "Gen X Early (1965-1970)",
        generation == "Gen X (1965-1980)" & cohort > 1970 & cohort <= 1976 ~ "Gen X Mid (1971-1976)",
        generation == "Gen X (1965-1980)" & cohort > 1976 ~ "Gen X Late (1977-1980)",
        generation == "Millennials (1981-1996)" & cohort <= 1986 ~ "Millennials Early (1981-1986)",
        generation == "Millennials (1981-1996)" & cohort > 1986 & cohort <= 1992 ~ "Millennials Mid (1987-1992)",
        generation == "Millennials (1981-1996)" & cohort > 1992 ~ "Millennials Late / Gen Z (1993-2004)",
    #    generation == "Gen Z (1997-2012)" & cohort <= 2002 ~ "Gen Z Early (1997-2002)",
    #    generation == "Gen Z (1997-2012)" & cohort > 2002 & cohort <= 2008 ~ "Gen Z Mid (2003-2008)",
    #    generation == "Gen Z (1997-2012)" & cohort > 2008 ~ "Gen Z Late (2009-2012)",
        TRUE ~ "Other"
      ),
      levels = c(
        "Greatest Early (1901-1910)", "Greatest Mid (1911-1918)", "Greatest Late (1919-1927)",
        "Silent Early (1928-1934)", "Silent Mid (1935-1940)", "Silent Late (1941-1945)",
        "Boomers Early (1946-1951)", "Boomers Mid (1952-1958)", "Boomers Late (1959-1964)",
        "Gen X Early (1965-1970)", "Gen X Mid (1971-1976)", "Gen X Late (1977-1980)",
        "Millennials Early (1981-1986)", "Millennials Mid (1987-1992)", 
        "Millennials Late / Gen Z (1993-2004)"
        #"Millennials Late (1993-1996)",
      #  "Gen Z Early (1997-2002)", "Gen Z Mid (2003-2008)", "Gen Z Late (2009-2012)" #,
      #  "Other"
      )
    )
  )



table(data_gss_generations$generation)
table(data_gss_generations$generation_two_sections)
table(data_gss_generations$generation_three_sections)

gss_svy_gen <- data_gss_generations %>%
  as_survey_design(ids = 1, weights = wtsscomp)

```


## Regression with age and cohorts

```{r}

# Create cohort groups (e.g., decades)
data_gss <- data_gss %>%
  mutate(cohort_group = cut(cohort, breaks = seq(1900, 2010, by = 10), right = FALSE,
                            labels = paste(seq(1900, 2000, by = 10), seq(1909, 2009, by = 10), sep = "-")))

# Cohort groups by 15 years
data_gss_15 <- data_gss %>%
  mutate(cohort_15_yr = cut(cohort, breaks = seq(1900, 2010, by = 15), right = FALSE,
                            labels = paste(seq(1900, 2000, by = 15), 15 + seq(1900, 2000, by = 15), sep = "-")))

# Create cohort groups (e.g., decades)
data_gss_10 <- data_gss %>%
  mutate(cohort_10_yr = cut(cohort, breaks = seq(1900, 2010, by = 10), right = FALSE,
                            labels = paste(seq(1900, 2000, by = 10), seq(1909, 2009, by = 10), sep = "-")))

# # Cohort groups by 5 years
data_gss_5 <- data_gss %>%
  mutate(cohort_05_yr = cut(cohort, breaks = seq(1900, 2005, by = 5), right = FALSE,
                            labels = paste(seq(1900, 2000, by = 5), 5 + seq(1900, 2000, by = 5), sep = "-")))

# Update survey design object
gss_svy_5 <- data_gss_5 %>%
  as_survey_design(ids = 1, weights = wtsscomp)

gss_svy_10 <- data_gss_10 %>%
  as_survey_design(ids = 1, weights = wtsscomp)

gss_svy_15 <- data_gss_15 %>%
  as_survey_design(ids = 1, weights = wtsscomp)


# 15 year cohorts
# Regression including age and cohort_group
weighted_lm_cohort <- gss_svy_15 %>%
  svyglm(health ~ age + cohort_15_yr, design = .)
# Summarize the model
summary(weighted_lm_cohort)

# 5 year cohorts
# Regression including age and cohort_group
weighted_lm_cohort <- gss_svy_10 %>%
  svyglm(health ~ age + cohort_10_yr, design = .)
# Summarize the model
summary(weighted_lm_cohort)

# 10 year cohorts
# Regression including age and cohort_group
weighted_lm_cohort <- gss_svy_5 %>%
  svyglm(health ~ age + cohort_05_yr, design = .)
# Summarize the model
summary(weighted_lm_cohort)



# Function to plot this
  
cohort_age_interation_figure <- function(cohort_length_string) {
  
# Prepare data for the plot
coef_data <- broom::tidy(weighted_lm_cohort) %>%
  filter(term != "(Intercept)") %>%
  mutate(significant = ifelse(p.value < 0.05, "Significant", "Not Significant"),
         term = ifelse(grepl("cohort_group", term), gsub("cohort_group", "", term), term))

# Plot
ggplot(coef_data, aes(x = term, y = estimate, fill = significant)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.2) +
  labs(
    title = "Cohort Effects on SRH",
    subtitle = cohort_length_string,
    x = "Term (Age or Cohort Birthyear Group)",
    y = "Coefficient Estimate",
    fill = "Significance"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

}

# Plot for different cohort sizes

weighted_lm_cohort <- gss_svy_5 %>%
  mutate(cohort_group = cohort_05_yr) %>% 
  svyglm(health ~ age + cohort_group, design = .)
cohort_age_interation_figure("5-Year Cohorts")

weighted_lm_cohort <- gss_svy_15 %>%
  mutate(cohort_group = cohort_15_yr) %>% 
  svyglm(health ~ age + cohort_group, design = .)
cohort_age_interation_figure("15-Year Cohorts")

weighted_lm_cohort <- gss_svy_10 %>%
  mutate(cohort_group = cohort_10_yr) %>% 
  svyglm(health ~ age + cohort_group, design = .)
cohort_age_interation_figure("10-Year Cohorts")




### Generations

library(ggplot2)
library(dplyr)
library(broom)

generation_age_interaction_figure <- function(cohort_length_string) {
  
  # Prepare data for the plot
  coef_data <- broom::tidy(weighted_lm_generation) %>%
    filter(term != "(Intercept)") %>%
    mutate(
      # Clean the term to remove "cohort_group" prefix if present
      term_cleaned = ifelse(grepl("cohort_group", term), gsub("cohort_group", "", term), term),
      # Extract the first year (numeric) from cohort terms or assign NA for "age"
      first_year = ifelse(term_cleaned == "age", NA, as.numeric(sub("-.*", "", term_cleaned))),
      # Add a numeric ordering key: assign 0 to "age" and the extracted year to cohort groups
      term_numeric = ifelse(term_cleaned == "age", 0, first_year),
      # Reorder terms based on the numeric ordering key
      term_ordered = reorder(term_cleaned, term_numeric),
      significant = ifelse(p.value < 0.05, "Significant", "Not Significant")
    )

  # Plot
  ggplot(coef_data, aes(x = term_ordered, y = estimate, fill = significant)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.2) +
    labs(
      title = "Cohort Effects on SRH",
      subtitle = cohort_length_string,
      x = "Term (Age or Cohort Birthyear Group)",
      y = "Coefficient Estimate",
      fill = "Significance"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

# Call the function with a sample subtitle
# generation_age_interaction_figure("Cohort Length: 10 Years")


library(ggplot2)
library(dplyr)
library(broom)

generation_age_interaction_figure <- function(cohort_length_string) {
  
  # Prepare data for the plot
  coef_data <- broom::tidy(weighted_lm_generation) %>%
    filter(term != "(Intercept)") %>%
    mutate(
      # Clean the term to remove "cohort_group" prefix if present
      term_cleaned = ifelse(grepl("cohort_group", term), gsub("cohort_group", "", term), term),
      # Extract the first year from cohort terms using regex (assign NA for "age")
      first_year = ifelse(term_cleaned == "age", NA, as.numeric(stringr::str_extract(term_cleaned, "^\\d{4}"))),
      # Add a numeric ordering key: assign 0 to "age" and use the extracted first year for cohorts
      term_numeric = ifelse(term_cleaned == "age", 0, first_year),
      # Reorder terms based on the numeric ordering key
      term_cleaned = factor(term_cleaned, levels = term_cleaned[order(term_numeric)]),
      significant = ifelse(p.value < 0.05, "Significant", "Not Significant")
    )

  # Plot
  ggplot(coef_data, aes(x = term_cleaned, y = estimate, fill = significant)) +
    geom_bar(stat = "identity", position = position_dodge()) +
    geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.2) +
    labs(
      title = "Cohort Effects on SRH",
      subtitle = cohort_length_string,
      x = "Term (Age or Cohort Birthyear Group)",
      y = "Coefficient Estimate",
      fill = "Significance"
    ) +
    theme_minimal() +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))
}

# Call the function with a sample subtitle
# cohort_age_interaction_figure("Cohort Length: 10 Years")



weighted_lm_generation <- gss_svy_gen %>% 
  mutate(cohort_group = generation_two_sections) %>% 
  svyglm(health ~ age + cohort_group, design = .)
generation_age_interaction_figure("Generations")

weighted_lm_generation <- gss_svy_gen %>% 
  mutate(cohort_group = generation) %>% 
  svyglm(health ~ age + cohort_group, design = .)
generation_age_interaction_figure("Generations")

weighted_lm_generation <- gss_svy_gen %>% 
  mutate(cohort_group = generation_three_sections) %>% 
  svyglm(health ~ age + cohort_group, design = .)
generation_age_interaction_figure("Generations")

```

Positive coefficients indicate better SRH compared to the 1900-1909 cohort (reference group), adjusting for age.

- SRH improved significantly for cohorts born between 1920-1979, peaking in 1940-1949.
- Recent cohorts (1980-2009) show no significant change.

## Interaction terms between age and cohort

Does the impact of age on SRH differ across cohorts?

```{r}




# Regression with interaction between age and cohort_group
weighted_lm_interaction <- gss_svy_10 %>%
  svyglm(health ~ age * cohort_group, design = .)

# Summarize the model
summary(weighted_lm_interaction)

# Visualize the interaction
library(ggplot2)
# Create a dataset for predictions
new_data <- expand.grid(
  age = seq(min(data_gss$age), max(data_gss$age), by = 1),
  cohort_group = levels(data_gss$cohort_group)
)

# Predict SRH

predictions <- as.data.frame(predict(weighted_lm_interaction, newdata = new_data, type = "response", se = TRUE))
new_data$predicted_health <- predictions$response
new_data$se <- predictions$SE



# Plot
ggplot(new_data, aes(x = age, y = predicted_health, color = cohort_group)) +
  geom_line() +
  labs(
    title = "Predicted SRH by Age and Cohort",
    x = "Age",
    y = "Predicted Self-Rated Health",
    color = "Cohort Group"
  ) +
  theme_minimal()

```

```{r, eval = FALSE}

library(dplyr)
library(broom)
library(ggplot2)

# Extract coefficients
coef_data <- broom::tidy(weighted_lm_interaction) %>%
  filter(term %in% grep("age:cohort_group", term, value = TRUE)) %>%
  mutate(
    cohort_group = gsub("age:cohort_group", "", term),
    significant = ifelse(p.value < 0.05, "Significant", "Not Significant")
  )

# Generate predictions for each cohort group
library(ggeffects)
predictions <- ggpredict(weighted_lm_cohort, terms = c("age", "cohort_group"))

# Plot predictions
ggplot(predictions, aes(x = x, y = predicted, color = group)) +
  geom_line(size = 1) +
  labs(
    title = "Interaction of Age and Cohort Group on Self-Rated Health",
    x = "Age",
    y = "Predicted SRH",
    color = "Cohort Group"
  ) +
  theme_minimal()

# Plot interaction coefficients
ggplot(coef_data, aes(x = cohort_group, y = estimate, fill = significant)) +
  geom_bar(stat = "identity", position = position_dodge()) +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.2) +
  labs(
    title = "Interaction Effects of Age and Cohort Group on SRH",
    x = "Cohort Group",
    y = "Interaction Coefficient (Age Effect)",
    fill = "Significance"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))



```
"Age and Cohort Interaction Effects on Self-Rated Health (SRH)"

Key Findings:

Higher Baseline SRH for Later Cohorts:

Cohorts born between 1920-2009 report significantly higher SRH compared to the 1900-1909 cohort.
The 2000-2009 cohort shows the highest baseline SRH but also the most variability.
Steeper Age-Related Declines in SRH for Recent Cohorts:

The effect of age on SRH becomes progressively more negative in later cohorts.
Individuals in the 2000-2009 cohort experience the steepest decline in SRH as they age.
Mid-Century Cohorts Show Moderate Trends:

Cohorts born between 1940-1969 exhibit improved SRH at younger ages with moderate age-related declines.
Implications for Public Health:

The steeper decline in SRH with age for more recent cohorts may reflect generational challenges, such as lifestyle changes, chronic disease prevalence, or systemic health inequities.
Strategies to support health across the lifespan may need to account for these emerging disparities.

# Incorporating covariates

```{r}

gss_svy <- gss_svy %>% 
  mutate(sex = as.factor(sex),
         race = as.factor(race))

# Load libraries
library(dplyr)
library(tidyr)
library(purrr)
library(broom)
library(ggplot2)

formulas <- list(
  age_only = health ~ age,
  age_race = health ~ age + race,
  age_sex = health ~ age + sex,
  age_hap = health ~ age + happy,
  age_sf_edu = health ~ age + satfin + educ,
  age_sf_hap = health ~ age + satfin + happy,
  age_sf_hap_educ = health ~ age + happy + satfin + educ,
  age_sf_hap_educ_race_sex = health ~ age + happy + satfin + educ + race + sex
)

# Get a vector of unique years
years <- sort(unique(data_gss$year))

# Create a data frame of all combinations
formula_df <- expand.grid(formula_name = names(formulas), year = years, stringsAsFactors = FALSE)

# Run regressions and extract results
results <- formula_df %>%
  mutate(
    formula = formulas[formula_name],
    data = map(year, ~ filter(data_gss, year == .x)),
    model = map2(formula, data, ~ lm(.x, data = .y)),
    tidy = map(model, tidy),
    glance = map(model, glance)
  )

# Extract coefficients
coefficients_df <- results %>%
  unnest(tidy)

# Extract model summary statistics
model_stats_df <- results %>%
  unnest(glance)

# Focus on age coefficient
age_coefficients <- coefficients_df %>%
  filter(term == "age") %>%
  select(formula_name, year, estimate, std.error, statistic, p.value)

# Regress Age Coefficient Over Time
#For each formula, regress the age coefficient over time to analyze the trend:

age_coeff_trend <- age_coefficients %>%
  group_by(formula_name) %>%
  nest() %>%
  mutate(
    trend_model = map(data, ~ lm(estimate ~ year, data = .x)),
    trend_tidy = map(trend_model, tidy),
    trend_glance = map(trend_model, glance)
  )

# Extract trend summary statistics
trend_summary <- age_coeff_trend %>%
  unnest(trend_glance) %>%
  select(formula_name, r.squared, adj.r.squared, p.value, statistic)

# Merge age coefficients with trend summary
final_results <- age_coefficients %>%
  left_join(trend_summary, by = "formula_name") %>%
  arrange(formula_name, year)

# Summary table
# Assuming 'age_coeff_trend' is from the previous analysis

# Extract trend statistics for each formula
trend_stats <- age_coeff_trend %>%
  mutate(
    trend_summary = map(trend_model, ~ {
      model <- .x
      summary <- summary(model)
      slope <- coef(summary)["year", "Estimate"]
      std_error <- coef(summary)["year", "Std. Error"]
      p_value <- coef(summary)["year", "Pr(>|t|)"]
      conf_int <- confint(model)["year", ]
      r_squared <- summary$r.squared
      
      tibble(
        slope = slope,
        std_error = std_error,
        ci_lower = conf_int[1],
        ci_upper = conf_int[2],
        p_value = p_value,
        r_squared = r_squared
      )
    })
  ) %>%
  unnest(trend_summary) %>%
  select(formula_name, slope, std_error, ci_lower, ci_upper, p_value, r_squared)

trend_stats

knitr::kable(trend_stats)

ggplot(age_coefficients, aes(x = year, y = estimate, color = formula_name)) +
  geom_line() +
#  geom_smooth(method = "lm", alpha = 0.2) +
  geom_point() +
  labs(
    title = "Age Coefficient for Various Models Over Time",
    x = "Year",
    y = "Age Coefficient",
    color = "Formula"
  ) +
  theme_minimal()

ggplot(age_coefficients, aes(x = year, y = estimate, color = formula_name)) +
 # geom_line() +
  geom_smooth(method = "lm", alpha = 0.2) +
  geom_point() +
  labs(
    title = "Trend for Age Coefficient for Various Models Over Time",
    x = "Year",
    y = "Age Coefficient",
    color = "Formula"
  ) +
  theme_minimal()


```

# Treat it as a Categorical Variable

SRH is actually an ordinal (categorical) variable. Let's explore what the distribution actually looks like over time, and then facet it by age. 

```{r}

# Self-Rated Health Category Distribution
data_gss %>%
  select(year, health) %>%
  filter(!is.na(health)) %>%
  mutate(
    srh = factor(health, 
                    levels = 1:4,
                    labels = c("Poor", "Fair", "Good", "Excellent"))) %>% 
  # Remove missing values
  # Calculate percentages by year
  group_by(year) %>%
  count(srh) %>%
  mutate(percent = n / sum(n) * 100) %>%
  ungroup() %>% 
  ggplot(aes(x = year, y = percent, color = srh)) +
  geom_line() +
  labs(title = "Self-Rated Health Category Distribution",
       subtitle = "GSS Dataset",
       y = "Percent",
       x = "Year") +
  theme_minimal()

# Self-Rated Health Category Distribution by Age Group
data_gss %>%
  select(year, health, age_group) %>%
  filter(!is.na(health)) %>%
  mutate(
    srh = factor(health, 
                 levels = 1:4,
                 labels = c("Poor", "Fair", "Good", "Excellent"))) %>% 
  # Remove missing values
  # Calculate percentages by year
  group_by(year, age_group) %>%
  count(srh) %>%
  mutate(percent = n / sum(n) * 100) %>%
  ungroup() %>% 
  ggplot(aes(x = year, y = percent, color = srh)) +
  geom_line() +
  facet_wrap(~age_group) +
  labs(title = "Self-Rated Health Category Distribution by Age Group",
       subtitle = "GSS Dataset",
       y = "Percent",
       x = "Year") +
  theme_minimal()

```


# Cohort Effects

## Visualize age profiles by cohort

First let's visualize potential cohort effects.
```{r}

gss_svy_15 %>% 
  mutate(cohort = cohort_15_yr) %>% 
#  mutate(cohort = cut(cohort, breaks = 7)) %>% # Create cohorts with 6 breaks
  group_by(age, cohort) %>% 
  summarize(mean_health = survey_mean(health, na.rm = TRUE)) %>% 
  ggplot(aes(x = age, y = mean_health, color = cohort)) +
  labs(title = "Age Profiles by Cohort") +
  geom_smooth()

gss_svy_gen %>% 
  mutate(cohort = generation_two_sections) %>% 
#  mutate(cohort = cut(cohort, breaks = 7)) %>% # Create cohorts with 6 breaks
  group_by(age, cohort) %>% 
  summarize(mean_health = survey_mean(health, na.rm = TRUE)) %>% 
  ggplot(aes(x = age, y = mean_health, color = cohort)) +
  labs(title = "Age Profiles by Cohort") +
 # geom_smooth()
  geom_line()

```



## Model srh vs age for each cohort and plot the betas and predicted age values

Let's try to see why the oldest cohort line sort of looks like it lines up with the youngest by modeling the trend of srh vs age (seems linear) and comparing the expected mean health rating at 20, 40, and 60 years of age for the different cohorts, as well as the coefficient.


```{r}

library(dplyr)
library(purrr)
library(srvyr)
library(broom)

lm_health_v_age_cohorts <- gss_svy %>%
  filter(cohort >= 1900 & cohort <= 1996) %>% 
  mutate(cohort_cut = cut(cohort, breaks = 25)) %>%  # Create cohort categories
  group_by(cohort_cut) %>%
  summarise(
    data = list(cur_data()) # Nest data for each cohort group
  ) %>%
  mutate(
    # Fit a weighted model for each cohort using the survey design
    model = map(data, ~ svyglm(health ~ age, design = as_survey_design(.x, weights = wtsscomp))),
    # Predict SRH at specific ages
    predict_20 = map2_dbl(model, data, ~ predict(.x, newdata = tibble(age = 20))),
    predict_25 = map2_dbl(model, data, ~ predict(.x, newdata = tibble(age = 25))),
    predict_30 = map2_dbl(model, data, ~ predict(.x, newdata = tibble(age = 30))),
    predict_35 = map2_dbl(model, data, ~ predict(.x, newdata = tibble(age = 35))),
    predict_40 = map2_dbl(model, data, ~ predict(.x, newdata = tibble(age = 40))),
    predict_45 = map2_dbl(model, data, ~ predict(.x, newdata = tibble(age = 45))),
    predict_50 = map2_dbl(model, data, ~ predict(.x, newdata = tibble(age = 50))),
    predict_55 = map2_dbl(model, data, ~ predict(.x, newdata = tibble(age = 55))),
    predict_65 = map2_dbl(model, data, ~ predict(.x, newdata = tibble(age = 65))),
    predict_70 = map2_dbl(model, data, ~ predict(.x, newdata = tibble(age = 70))),
    predict_75 = map2_dbl(model, data, ~ predict(.x, newdata = tibble(age = 75))),
    predict_80 = map2_dbl(model, data, ~ predict(.x, newdata = tibble(age = 80))),
    # Extract the coefficient (beta) and standard error
    beta = map_dbl(model, ~ coef(.x)["age"]),
    se = map_dbl(model, ~ broom::tidy(.x) %>% filter(term == "age") %>% pull(std.error)),
    # Calculate the mean cohort year
    cohort_mean = map_dbl(cohort_cut, ~ mean(as.numeric(stringr::str_extract_all(.x, "\\d+")[[1]])))
  )


# View the results
print(lm_health_v_age_cohorts)

# Plot the coefficient (beta) for SRH vs Age for each cohort
lm_health_v_age_cohorts %>%
  ggplot(aes(x = cohort_mean, y = beta)) +
  geom_errorbar(aes(ymin = beta - 1.96 * se, ymax = beta + 1.96 * se), width = 0.2) +
  labs(
    title = "Estimate for Coefficient for SRH vs Age for each Cohort",
    x = "Birth Year",
    y = "Coefficient"
  ) +
  geom_point()

# Plot predicted SRH at age 20 for each cohort
lm_health_v_age_cohorts %>%
  ggplot(aes(x = cohort_mean, y = predict_20)) +
  labs(
    title = "Predicted SRH at age 20 for each Cohort",
    x = "Birth Year",
    y = "Predicted SRH"
  ) +
  geom_point()

ggplot(lm_health_v_age_cohorts, aes(x = cohort_mean, y = predict_30)) +
  labs(title = "Predicted SRH at age 30 for each Cohort") +
  geom_point()
 
ggplot(lm_health_v_age_cohorts, aes(x = cohort_mean, y = predict_40)) +
  labs(title = "Predicted SRH at age 40 for each Cohort") +
  geom_point()

  ggplot(lm_health_v_age_cohorts, aes(x = cohort_mean, y = predict_65)) +
  labs(title = "Predicted SRH at age 65 for each Cohort") +
  geom_point()

# Plot many ages
  
# Reshape the data: gather all the prediction columns into one long format
lm_health_long <- lm_health_v_age_cohorts %>%
  pivot_longer(cols = starts_with("predict"), 
               names_to = "prediction_age", 
               values_to = "predicted_srh")

# Create the plot using the long data format
p <- ggplot(lm_health_long, aes(x = cohort_mean, y = predicted_srh)) +
  geom_point() +
  facet_wrap(~prediction_age, ncol = 4) +  # Facet by the different predictions (predict_20, predict_25, etc.)
  labs(
    title = "Predicted Self-Rated Health at Various Ages for Each Cohort",
    x = "Cohort Mean",
    y = "Predicted Self-Rated Health"
  )

# Display the combined plot
print(p)


```



# Facet by cohort

```{r}


library(dplyr)
library(purrr)
library(survey)
library(broom)

weighted_lm_by_cohort <- gss_svy %>%
  filter(cohort >= 1900 & cohort <= 1996) %>% 
  mutate(cohort_cut = cut(cohort, breaks = 25)) %>% 
  group_by(cohort_cut) %>%
  group_map_dfr(~ {
    # Create a survey design for the current group
    survey_design <- .x %>%
      as_survey_design(weights = wtsscomp)
    
    # Fit the weighted regression model
    model <- svyglm(health ~ age, design = survey_design)
    
    # Extract model results
    broom::tidy(model, conf.int = TRUE) %>%
      mutate(cohort_cut = unique(.x$cohort_cut)) # Add cohort_cut for identification
  }) %>%
  filter(term == "age") %>%
  select(cohort_cut, estimate, std.error, conf.low, conf.high, statistic, p.value) %>% 
  mutate(cohort = map_dbl(cohort_cut, ~ mean(as.numeric(str_extract_all(.x, "\\d+")[[1]]))))




knitr::kable(weighted_lm_by_cohort)

summary(weighted_lm_by_cohort)

# Plot the coefficients with error bars
weighted_lm_by_cohort %>% 
ggplot(aes(x = cohort, y = estimate)) +
  # geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), width=.2,
                position=position_dodge(0.05)) +
  # geom_ribbon(aes(ymin = conf.low, ymax = conf.high), alpha = 0.2) +
  labs(
    title = "Coefficient of Age for Different Cohorts",
    subtitle = "GSS Dataset",
    x = "Cohort",
    y = "Coefficient of Age"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 

# Regress the age coefficients on year
coef_model <- lm(estimate ~ cohort, data = weighted_lm_by_cohort)
summary(coef_model)

# Plot the regression
ggplot(weighted_lm_by_cohort, aes(x = cohort, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin=conf.low, ymax=conf.high), width=.2,
                position=position_dodge(0.05)) +
  geom_smooth(method = "lm", se = TRUE, alpha = 0.3) +
  labs(
    title = "Regression of 'Age' Coefficient Over Years",
    subtitle = "GSS Dataset",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()

weighted_lm_by_cohort
knitr::kable(weighted_lm_by_cohort)


summary(weighted_lm_by_cohort)

# Perform linear regression of 'coef' (age coefficient) vs 'year'
lm_coef_vs_year_cohort <- lm(estimate ~ cohort, data = weighted_lm_by_cohort)

weighted_lm_by_cohort

# View the summary of the regression
summary(lm_coef_vs_year_cohort)


```

## Cohort facet

```{r, eval = FALSE}
# health vs year per cohort
gss_svy_gen %>% 
  group_by(age, generation_three_sections) %>% 
  summarize(mean_health = survey_mean(health)) %>% 
  ggplot(aes(x = age, y = mean_health)) +
  geom_line(color = "cornflowerblue") +
  facet_wrap(~ generation_three_sections) +
  labs(title = "Self-Rated Health Over Time (Per Cohort)",
       subtitle = "GSS Dataset",
       x = "Age", 
       y = "Average SRH",
       )
```

# Stratify

## Plot faceted

```{r, eval = FALSE}

library(dplyr)
library(broom)
library(ggplot2)


weighted_lm_by_group <- gss_svy %>%
  group_by(year, race, sex) %>%
  group_map_dfr(~ {
    model <- survey::svyglm(health ~ age, design = .x)
    tidy(model, conf.int = TRUE)
  }) %>%
  filter(term == "age") %>%
  select(year, race, sex, estimate, std.error, conf.low, conf.high, statistic, p.value)

library(dplyr)
library(broom)
library(ggplot2)
library(purrr)

# If you haven't already run it, here's the code to create weighted_lm_by_group
# weighted_lm_by_group <- gss_svy %>%
#   group_by(year, race, sex) %>%
#   group_map_dfr(~ {
#     model <- survey::svyglm(health ~ age, design = .x)
#     tidy(model, conf.int = TRUE)
#   }) %>%
#   filter(term == "age") %>%
#   select(year, race, sex, estimate, std.error, conf.low, conf.high, statistic, p.value)

# Nest data and fit models for slope calculations
coef_info <- weighted_lm_by_group %>%
  group_by(race, sex) %>%
  nest() %>%
  mutate(
    mod = map(data, ~ lm(estimate ~ year, data = .x)),
    tidied = map(mod, tidy),
    glanced = map(mod, glance)
  ) %>%
  unnest(cols = c(tidied, glanced)) %>%
  filter(term == "year") %>%
  # We now have slope (estimate), p.value, and r.squared for each race-sex group
  select(race, sex, slope = estimate, p.value, r.squared)

# Determine annotation positions for each facet
text_positions <- weighted_lm_by_group %>%
  group_by(race, sex) %>%
  summarize(
    x_text = max(year),
    y_text = max(estimate),
    .groups = "drop"
  )

# Join slope info with positions
coef_info <- left_join(coef_info, text_positions, by = c("race", "sex"))

# Plot facets
ggplot(weighted_lm_by_group, aes(x = year, y = estimate)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(race ~ sex) +
  geom_text(
    data = coef_info,
    aes(
      x = x_text,
      y = y_text,
      label = sprintf("Slope=%.3f, p=%.3f, R²=%.3f", slope, p.value, r.squared)
    ),
    hjust = 1, vjust = 1, 
    size = 3
  ) +
  labs(
    title = "Change in 'Age' Coefficient Over Years",
    subtitle = "Stratified by Race and Sex",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()

```



```{r, eval = TRUE}

library(tidyverse)

library(dplyr)
library(purrr)
library(broom)

weighted_lm_by_group <- gss_svy %>%
  group_by(year, race, sex) %>%
  group_map_dfr(~ {
    model <- survey::svyglm(health ~ age, design = .x)
    tidy(model, conf.int = TRUE)
  }) %>%
  filter(term == "age") %>%
  select(year, race, sex, estimate, std.error, conf.low, conf.high, statistic, p.value)

coef_info <- weighted_lm_by_group %>%
  group_by(race, sex) %>%
  nest() %>%
  mutate(
    mod = map(data, ~ lm(estimate ~ year, data = .x)),
    # Select only the columns you need from `tidy` and `glance`
    tidied = map(mod, ~ tidy(.x) %>% select(term, estimate, p.value)),
    glanced = map(mod, ~ glance(.x) %>% select(r.squared))
  ) %>%
  unnest(cols = tidied) %>%
  filter(term == "year") %>%
  unnest(cols = glanced) %>%
  select(race, sex, slope = estimate, p.value, r.squared)

text_positions <- weighted_lm_by_group %>%
  group_by(race, sex) %>%
  summarize(
    x_text = max(year),
    y_text = max(estimate),
    .groups = "drop"
  )

coef_info <- left_join(coef_info, text_positions, by = c("race", "sex"))

ggplot(weighted_lm_by_group, aes(x = year, y = estimate)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(race ~ sex) +
  geom_text(
    data = coef_info,
    inherit.aes = FALSE, # turn off inheriting x/y from the main plot
    aes(
      x = x_text,
      y = y_text,
      label = sprintf("Slope=%.3e, p=%.3e, R²=%.3f", slope, p.value, r.squared)
    ),
    hjust = 1, vjust = 1, size = 3
  ) +
  labs(
    title = "Change in 'Age' Coefficient Over Years",
    subtitle = "Stratified by Race and Sex",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()


```



```{r}




# Create age groups
gss_svy <- gss_svy %>%
  mutate(age_group = cut(age, breaks = 6))

# Compute weighted mean health by age group and year
weighted_health_by_age <- gss_svy %>%
  group_by(age_group, year) %>%
  summarise(
    mean_health = survey_mean(health, na.rm = TRUE)
  )

ggplot(weighted_health_by_age, aes(x = year, y = mean_health, color = age_group)) +
#  geom_smooth(alpha = 0.2) +
  geom_line() +
  geom_point() +
  labs(
    title = "Average SRH Per Year for Each Age Group",
    subtitle = "GSS Dataset with Survey Weights",
    x = "Year",
    y = "Average Self-Rated Health",
    color = "Age Group"
  ) +
  theme_minimal()


```

```{r}
library(srvyr)

# Convert to srvyr object
gss_svy <- as_survey_design(gss_svy)

# Now you can use mutate directly
gss_svy <- gss_svy %>%
  mutate(age_group = cut(age, breaks = 6))

gss_svy <- gss_svy %>%
  mutate(age_group = cut(age, breaks = 6))

# Compute weighted mean health by age group, year, race, and sex
weighted_health_by_age <- gss_svy %>%
  group_by(race, sex, age_group, year) %>%
  summarise(
    mean_health = srvyr::survey_mean(health, na.rm = TRUE),
    .groups = "drop"
  )


ggplot(weighted_health_by_age, aes(x = year, y = mean_health, color = age_group)) +
  geom_line() +
  geom_point() +
  facet_grid(race ~ sex) +
  # Add annotations for each line within each facet
  # geom_text(
  #   data = coef_info,
  #   inherit.aes = FALSE,
  #   aes(
  #     x = x_text,
  #     y = y_text,
  #     label = sprintf("Slope=%.3e, p=%.3e, R²=%.3f", slope, p.value, r.squared),
  #     color = age_group
  #   ),
  #   hjust = 0,  # left-align at x_text
  #   vjust = 1,  # top-align at y_text
  #   size = 3,
  #   show.legend = FALSE # Don't want a separate legend for text color
  # ) +
  labs(
    title = "Average SRH Per Year by Age Group, Stratified by Race and Sex",
    subtitle = "GSS Dataset with Survey Weights",
    x = "Year",
    y = "Average Self-Rated Health",
    color = "Age Group"
  ) +
  theme_minimal()

```

```{r}

gss_svy %>% 
  group_by(sex, race, year) %>% 
  summarize(n = n()) %>% 
  group_by(sex, race ) %>% 
  summarize(avg_year = mean(n))

```

We do not have enough power to do a stratification by race.

## Stratify by sex and satfin

```{r}

gss_svy %>% 
  group_by(sex, satfin, year) %>% 
  summarize(n = n()) %>% 
  group_by(sex, satfin ) %>% 
  summarize(avg_year = mean(n))

```

# Stratify!



```{r, eval = FALSE}

library(dplyr)
library(broom)
library(ggplot2)


weighted_lm_by_group <- gss_svy %>%
  group_by(year, satfin, sex) %>%
  group_map_dfr(~ {
    model <- survey::svyglm(health ~ age, design = .x)
    tidy(model, conf.int = TRUE)
  }) %>%
  filter(term == "age") %>%
  select(year, satfin, sex, estimate, std.error, conf.low, conf.high, statistic, p.value)

library(dplyr)
library(broom)
library(ggplot2)
library(purrr)
# 
# # If you haven't already run it, here's the code to create weighted_lm_by_group
# weighted_lm_by_group <- gss_svy %>%
#   group_by(year, satfin, sex) %>%
#   group_map_dfr(~ {
#     model <- survey::svyglm(health ~ age, design = .x)
#     tidy(model, conf.int = TRUE)
#   }) %>%
#   filter(term == "age") %>%
#   select(year, satfin, sex, estimate, std.error, conf.low, conf.high, statistic, p.value)

# Nest data and fit models for slope calculations
coef_info <- weighted_lm_by_group %>%
  group_by(satfin, sex) %>%
  nest() %>%
  mutate(
    mod = map(data, ~ lm(estimate ~ year, data = .x)),
    tidied = map(mod, tidy),
    glanced = map(mod, glance)
  ) %>%
  unnest(cols = c(tidied, glanced)) %>%
  filter(term == "year") %>%
  # We now have slope (estimate), p.value, and r.squared for each satfin-sex group
  select(satfin, sex, slope = estimate, p.value, r.squared)

# Determine annotation positions for each facet
text_positions <- weighted_lm_by_group %>%
  group_by(satfin, sex) %>%
  summarize(
    x_text = max(year),
    y_text = max(estimate),
    .groups = "drop"
  )

# Join slope info with positions
coef_info <- left_join(coef_info, text_positions, by = c("satfin", "sex"))

# Plot
ggplot(weighted_lm_by_group, aes(x = year, y = estimate)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(satfin ~ sex) +
  geom_text(
    data = coef_info,
    aes(
      x = x_text,
      y = y_text,
      label = sprintf("Slope=%.3f, p=%.3f, R²=%.3f", slope, p.value, r.squared)
    ),
    hjust = 1, vjust = 1, 
    size = 3
  ) +
  labs(
    title = "Change in 'Age' Coefficient Over Years",
    subtitle = "Stratified by satfin and Sex",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()

```



```{r, eval = TRUE}

library(tidyverse)

library(dplyr)
library(purrr)
library(broom)

weighted_lm_by_group <- gss_svy %>%
  group_by(year, satfin, sex) %>%
  group_map_dfr(~ {
    model <- survey::svyglm(health ~ age, design = .x)
    tidy(model, conf.int = TRUE)
  }) %>%
  filter(term == "age") %>%
  select(year, satfin, sex, estimate, std.error, conf.low, conf.high, statistic, p.value)

coef_info <- weighted_lm_by_group %>%
  group_by(satfin, sex) %>%
  nest() %>%
  mutate(
    mod = map(data, ~ lm(estimate ~ year, data = .x)),
    # Select only the columns you need from `tidy` and `glance`
    tidied = map(mod, ~ tidy(.x) %>% select(term, estimate, p.value)),
    glanced = map(mod, ~ glance(.x) %>% select(r.squared))
  ) %>%
  unnest(cols = tidied) %>%
  filter(term == "year") %>%
  unnest(cols = glanced) %>%
  select(satfin, sex, slope = estimate, p.value, r.squared)

text_positions <- weighted_lm_by_group %>%
  group_by(satfin, sex) %>%
  summarize(
    x_text = max(year),
    y_text = max(estimate),
    .groups = "drop"
  )

coef_info <- left_join(coef_info, text_positions, by = c("satfin", "sex"))

ggplot(weighted_lm_by_group, aes(x = year, y = estimate)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_grid(satfin ~ sex) +
  geom_text(
    data = coef_info,
    inherit.aes = FALSE, # turn off inheriting x/y from the main plot
    aes(
      x = x_text,
      y = y_text,
      label = sprintf("Slope=%.3e, p=%.3e, R²=%.3f", slope, p.value, r.squared)
    ),
    hjust = 1, vjust = 1, size = 3
  ) +
  labs(
    title = "Change in 'Age' Coefficient Over Years",
    subtitle = "Stratified by satfin and Sex",
    x = "Year",
    y = "Coefficient of Age"
  ) +
  theme_minimal()


```



```{r}




# Create age groups
gss_svy <- gss_svy %>%
  mutate(age_group = cut(age, breaks = 6))

# Compute weighted mean health by age group and year
weighted_health_by_age <- gss_svy %>%
  group_by(age_group, year) %>%
  summarise(
    mean_health = survey_mean(health, na.rm = TRUE)
  )

ggplot(weighted_health_by_age, aes(x = year, y = mean_health, color = age_group)) +
#  geom_smooth(alpha = 0.2) +
  geom_line() +
  geom_point() +
  labs(
    title = "Average SRH Per Year for Each Age Group",
    subtitle = "GSS Dataset with Survey Weights",
    x = "Year",
    y = "Average Self-Rated Health",
    color = "Age Group"
  ) +
  theme_minimal()



# Compute weighted mean health by age group, year, satfin, and sex
weighted_health_by_age <- gss_svy %>%
  group_by(happy, sex, age_group, year) %>%
  summarise(
    mean_health = srvyr::survey_mean(health, na.rm = TRUE),
    .groups = "drop"
  )


  # Add annotations for each line within each facet
  # geom_text(
  #   data = coef_info,
  #   inherit.aes = FALSE,
  #   aes(
  #     x = x_text,
  #     y = y_text,
  #     label = sprintf("Slope=%.3e, p=%.3e, R²=%.3f", slope, p.value, r.squared),
  #     color = age_group
  #   ),
  #   hjust = 0,  # left-align at x_text
  #   vjust = 1,  # top-align at y_text
  #   size = 3,
  #   show.legend = FALSE # Don't want a separate legend for text color
  # ) +


```

# Stratify by one variable

## Set up factor survey object with appropriate labels
```{r}

library(survey)
library(srvyr)
library(dplyr)
library(forcats)

# Assuming gss_svy is already a srvyr survey object, for example:
# gss_svy <- as_survey(gss_design)  # If you started with a svydesign called gss_design

gss_svy_labeled <- gss_svy %>%
  mutate(
    # Reverse-coded variables:
    # health: 1=Excellent, 2=Good, 3=Fair, 4=Poor
    # Reversed: Poor (was 4), Fair (was 3), Good (was 2), Excellent (was 1)
    health = recode_factor(as.factor(health),
      `1` = "Poor",
      `2` = "Fair",
      `3` = "Good",
      `4` = "Excellent"
    ),

    # happy: 1=Very happy, 2=Pretty happy, 3=Not too happy
    # Reversed: Not too happy (was 3), Pretty happy (was 2), Very happy (was 1)
    happy = recode_factor(as.factor(happy),
      `1` = "Not too happy",
      `2` = "Pretty happy",
      `3` = "Very happy"
    ),

    # life: 1=Exciting, 2=Routine, 3=Dull
    # Reversed: Dull (was 3), Routine (was 2), Exciting (was 1)
    life = recode_factor(as.factor(life),
      `1` = "Dull",
      `2` = "Routine",
      `3` = "Exciting"
    ),

    # satfin: 1=Satisfied, 2=More or less, 3=Not satisfied
    # Reversed: Not satisfied (was 3), More or less (was 2), Satisfied (was 1)
    satfin = recode_factor(as.factor(satfin),
      `1` = "Not satisfied",
      `2` = "More or less",
      `3` = "Satisfied"
    ),

    # Non-reversed variables:
    # sex: 1=Male, 2=Female
    sex = recode_factor(as.factor(sex),
      `1` = "Male",
      `2` = "Female"
    ),

    # polviews: 1=Extremely liberal, 2=Liberal, 3=Slightly liberal,
    #           4=Moderate, 5=Slightly conservative, 6=Conservative, 7=Extremely conservative
    polviews = recode_factor(as.factor(polviews),
      `1` = "Extremely liberal",
      `2` = "Liberal",
      `3` = "Slightly liberal",
      `4` = "Moderate",
      `5` = "Slightly conservative",
      `6` = "Conservative",
      `7` = "Extremely conservative"
    ),

    # class: 1=Lower class, 2=Working class, 3=Middle class, 4=Upper class
    class = recode_factor(as.factor(class),
      `1` = "Lower class",
      `2` = "Working class",
      `3` = "Middle class",
      `4` = "Upper class"
    ),

    # attend: 0=Never, 1=Less than once a year, 2=Once a year,
    #         3=Several times a year, 4=Once a month, 5=2-3 times a month,
    #         6=Nearly every week, 7=Every week, 8=More than once a week
    attend = recode_factor(as.factor(attend),
      `0` = "Never",
      `1` = "Less than once a year",
      `2` = "Once a year",
      `3` = "Several times a year",
      `4` = "Once a month",
      `5` = "2-3 times a month",
      `6` = "Nearly every week",
      `7` = "Every week",
      `8` = "More than once a week"
    ),

    # race: 1=White, 2=Black, 3=Other
    race = recode_factor(as.factor(race),
      `1` = "White",
      `2` = "Black",
      `3` = "Other"
    ),

    # wrkstat: 1=Working fulltime, 2=Working parttime,
    #          3=Temp not working, 4=Unemployed/laid off,
    #          5=Retired, 6=In school, 7=Housewife, 8=Other
    wrkstat = recode_factor(as.factor(wrkstat),
      `1` = "Working fulltime",
      `2` = "Working parttime",
      `3` = "Temporary not working",
      `4` = "Unemployed/laid off",
      `5` = "Retired",
      `6` = "In school",
      `7` = "Housewife",
      `8` = "Other"
    ),

    # degree: 0=Less than HS, 1=High school, 2=Junior college, 3=Bachelor, 4=Graduate
    degree = recode_factor(as.factor(degree),
      `0` = "Less than high school",
      `1` = "High school",
      `2` = "Junior college",
      `3` = "Bachelor",
      `4` = "Graduate"
    ),

    # marital: 1=Married, 2=Widowed, 3=Divorced, 4=Separated, 5=Never married
    marital = recode_factor(as.factor(marital),
      `1` = "Married",
      `2` = "Widowed",
      `3` = "Divorced",
      `4` = "Separated",
      `5` = "Never married"
    ),

    # partyid: 0=Strong Democrat, 1=Not strong Democrat, 2=Ind near Dem,
    #          3=Independent, 4=Ind near Rep, 5=Not strong Rep,
    #          6=Strong Republican, 7=Other
    partyid = recode_factor(as.factor(partyid),
      `0` = "Strong Democrat",
      `1` = "Not strong Democrat",
      `2` = "Independent, near Democrat",
      `3` = "Independent",
      `4` = "Independent, near Republican",
      `5` = "Not strong Republican",
      `6` = "Strong Republican",
      `7` = "Other party"
    )
    # ),
    # 
    # # region: 1=Northeast, 2=North Central, 3=South, 4=West
    # region = recode_factor(as.factor(region),
    #   `1` = "Northeast",
    #   `2` = "North Central",
    #   `3` = "South",
    #   `4` = "West"
    )
  
gss_svy_labeled <- gss_svy_labeled %>%
  mutate(age_group = as.factor( 
                            cut(
                            age,
                            breaks = c(17, 29, 39, 49, 59, 69, Inf),
                            labels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+"),
                            right = TRUE
                          ))) 

```


## Stratify plots

### Plot 1: Average SRH per year by age group, stratified by a variable

```{r}


stratify_srh_yr_plot <- function(var_name, var_char) {
  var_sym <- rlang::sym(var_name)
  
  # Compute weighted mean health by the variable of interest, age group, and year
  gss_svy_labeled %>%
    group_by(!!var_sym, age_group, year) %>%
    dplyr::summarise(
      mean_health = srvyr::survey_mean(as.numeric(health), na.rm = TRUE),
      .groups = "drop"
    ) %>%
    ggplot(aes(x = year, y = mean_health, color = age_group)) +
    geom_line() +
    geom_point() +
    facet_wrap(vars(!!var_sym)) +  # Use vars() for facet_wrap in ggplot2
    labs(
      title = paste0("Average SRH Per Year by Age Group"),
      subtitle = paste0("GSS: Stratified by ", var_char),
      x = "Year",
      y = "Average Self-Rated Health",
      color = "Age Group"
    ) +
    theme_minimal()
}


stratify_srh_yr_plot("sex", "Sex")
stratify_srh_yr_plot("race", "Race")
stratify_srh_yr_plot("happy", "Happiness")
stratify_srh_yr_plot("life", "Life Satisfaction")
stratify_srh_yr_plot("polviews", "Political Views")
stratify_srh_yr_plot("class", "Class")
stratify_srh_yr_plot("satfin", "Financial Satisfaction")
#stratify_srh_yr_plot("region", "Region")
stratify_srh_yr_plot("attend", "Attend")
stratify_srh_yr_plot("wrkstat", "wrkstat")
#stratify_srh_yr_plot("hispanic", "Hispanic")
stratify_srh_yr_plot("degree", "Degree")
stratify_srh_yr_plot("marital", "Marital Status")
stratify_srh_yr_plot("partyid", "Party ID")
#stratify_srh_yr_plot("south", "South")



```

### Plot 2: Change in Age Coefficient on SRH by year, stratified by a variable

### Faceted

```{r}

library(tidyverse)

library(dplyr)
library(purrr)
library(broom)

stratify_coeff_plot <- function(var_name, var_char) {
  var_sym <- rlang::sym(var_name)

  weighted_lm_by_group <- gss_svy_labeled %>%
    group_by(year, !!var_sym) %>%
    group_map_dfr(~ {
      model <- survey::svyglm(as.numeric(health) ~ age, design = .x)
      tidy(model, conf.int = TRUE)
    }) %>%
    filter(term == "age") %>%
    select(year, !!var_sym, estimate, std.error, conf.low, conf.high, statistic, p.value)

  coef_info <- weighted_lm_by_group %>%
    group_by(!!var_sym) %>%
    nest() %>%
    mutate(
      mod = map(data, ~ lm(estimate ~ year, data = .x)),
      tidied = map(mod, ~ tidy(.x) %>% select(term, estimate, p.value)),
      glanced = map(mod, ~ glance(.x) %>% select(r.squared))
    ) %>%
    unnest(cols = tidied) %>%
    filter(term == "year") %>%
    unnest(cols = glanced) %>%
    select(!!var_sym, slope = estimate, p.value, r.squared)

  text_positions <- weighted_lm_by_group %>%
    group_by(!!var_sym) %>%
    dplyr::summarize(
      x_text = min(year),
      y_text = -0.005, #max(estimate),
      .groups = "drop"
    )

  coef_info <- left_join(coef_info, text_positions, by = var_name)

  ggplot(weighted_lm_by_group, aes(x = year, y = estimate)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    facet_grid(cols = vars(!!var_sym)) +  # Fixed facetting
    geom_text(
        data = coef_info,
        inherit.aes = FALSE,
        aes(
          x = x_text,
          y = y_text,
          label = sprintf("Slope = %.3e\np = %.3e\nR² = %.3f", slope, p.value, r.squared)
        ),
        hjust = 0, vjust = 0, size = 3
      ) +
    labs(
      title = "Change in 'Age' Coefficient Over Years",
      subtitle = paste0("Stratified by ", var_char),
      x = "Year",
      y = "Coefficient of Age"
    ) +
    theme_minimal()
}





stratify_coeff_plot("sex", "Sex")
stratify_coeff_plot("race", "Race")
stratify_coeff_plot("happy", "Happiness")
stratify_coeff_plot("life", "Life Satisfaction")
stratify_coeff_plot("polviews", "Political Views")
stratify_coeff_plot("class", "Class")
stratify_coeff_plot("satfin", "Financial Satisfaction")
#stratify_coeff_plot("region", "Region")
stratify_coeff_plot("attend", "Attend")
stratify_coeff_plot("wrkstat", "wrkstat")
#stratify_coeff_plot("hispanic", "Hispanic")
stratify_coeff_plot("degree", "Degree")
stratify_coeff_plot("marital", "Marital Status")
stratify_coeff_plot("partyid", "Party ID")
# stratify_coeff_plot("south", "South")

```

### Combined

```{r}
library(tidyverse)
library(broom)
library(survey)

stratify_coeff_combined_plot_separate_table <- function(var_name, var_char) {
  var_sym <- rlang::sym(var_name)

  # Compute the weighted linear model results by group
  weighted_lm_by_group <- gss_svy_labeled %>%
    group_by(year, !!var_sym) %>%
    group_map_dfr(~ {
      model <- survey::svyglm(as.numeric(health) ~ age, design = .x)
      tidy(model, conf.int = TRUE)
    }) %>%
    filter(term == "age") %>%
    select(year, !!var_sym, estimate, std.error, conf.low, conf.high, statistic, p.value)

  # Compute slope, p-values, and R² for each stratification category
  coef_info <- weighted_lm_by_group %>%
    group_by(!!var_sym) %>%
    nest() %>%
    mutate(
      mod = map(data, ~ lm(estimate ~ year, data = .x)),
      tidied = map(mod, ~ tidy(.x) %>% select(term, estimate, p.value)),
      glanced = map(mod, ~ glance(.x) %>% select(r.squared))
    ) %>%
    unnest(cols = tidied) %>%
    filter(term == "year") %>%
    unnest(cols = glanced) %>%
    select(!!var_sym, slope = estimate, p.value, r.squared)

  # Create a table for output
  results_table <- coef_info %>%
    dplyr::rename(stratum = !!var_sym)

  # Create a single plot with different colors for each stratification category
  p <- ggplot(weighted_lm_by_group, aes(x = year, y = estimate, color = !!var_sym)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    labs(
      title = "Change in 'Age' Coefficient Over Years",
      subtitle = paste0("Stratified by ", var_char),
      x = "Year",
      y = "Coefficient of Age",
      color = var_char
    ) +
    theme_minimal()

  # Return both the plot and the table
  list(
    plot = p,
    table = results_table
  )
}

# Example usage:
 out <- stratify_coeff_combined_plot_separate_table("happy", "Happiness")
 out$plot   # to see the plot
 out$table  # to see the results table
 
 knitr::kable(out$table)

```

### Combined with table

```{r}

library(tidyverse)
library(broom)
library(survey)
library(cowplot)    # for get_legend and plot_grid
library(gridExtra)  # for tableGrob

stratify_coeff_plot_combined_with_table <- function(var_name, var_char) {
  var_sym <- rlang::sym(var_name)

  # Compute the weighted linear model results by group
  weighted_lm_by_group <- gss_svy_labeled %>%
    group_by(year, !!var_sym) %>%
    group_map_dfr(~ {
      model <- survey::svyglm(as.numeric(health) ~ age, design = .x)
      tidy(model, conf.int = TRUE)
    }) %>%
    filter(term == "age") %>%
    select(year, !!var_sym, estimate, std.error, conf.low, conf.high, statistic, p.value)

  # Compute slope, p-values, and R² for each stratification category
  coef_info <- weighted_lm_by_group %>%
    group_by(!!var_sym) %>%
    nest() %>%
    mutate(
      mod = map(data, ~ lm(estimate ~ year, data = .x)),
      tidied = map(mod, ~ tidy(.x) %>% select(term, estimate, p.value)),
      glanced = map(mod, ~ glance(.x) %>% select(r.squared))
    ) %>%
    unnest(cols = tidied) %>%
    filter(term == "year") %>%
    unnest(cols = glanced) %>%
    select(!!var_sym, slope = estimate, p.value, r.squared)

  # Rename for table clarity
  results_table <- coef_info %>%
    dplyr::rename(Stratum = !!var_sym,
           Slope = slope,
           `p-value` = p.value,
           `R²` = r.squared)

  # Create main plot with legend at top
  p <- ggplot(weighted_lm_by_group, aes(x = year, y = estimate, color = !!var_sym)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    labs(
      title = "Change in 'Age' Coefficient Over Years",
      subtitle = paste0("Stratified by ", var_char),
      x = "Year",
      y = "Coefficient of Age",
      color = var_char
    ) +
    theme_minimal() #+
   # theme(legend.position = "top")

  # Extract the legend
  legend <- cowplot::get_legend(p)

  # Remove the legend from the main plot
 # p_no_legend <- p + theme(legend.position = "none")

  # Convert table into a graphical object
  table_grob <- tableGrob(results_table)

  # Combine legend, table, and plot into one figure
  # Legend at top, table underneath legend, and plot below that
  final_plot <- plot_grid(
    p,
    table_grob,
   # p_no_legend,
    ncol = 1,
    rel_heights = c(1, 0.4) # Adjust heights as needed
  )

  # Return the combined figure
  final_plot
}

# Example usage:
stratify_coeff_plot_combined_with_table("happy", "Happiness")


```

### Colored table

```{r}

library(tidyverse)
library(broom)
library(survey)
library(cowplot)
library(gridExtra)
library(scales) # for alpha()

stratify_coeff_plot_with_colored_table <- function(var_name, var_char) {
  var_sym <- rlang::sym(var_name)

  # Compute the weighted linear model results by group
  weighted_lm_by_group <- gss_svy_labeled %>%
    group_by(year, !!var_sym) %>%
    group_map_dfr(~ {
      model <- survey::svyglm(as.numeric(health) ~ age, design = .x)
      tidy(model, conf.int = TRUE)
    }) %>%
    filter(term == "age") %>%
    select(year, !!var_sym, estimate, std.error, conf.low, conf.high, statistic, p.value)

  # Compute slope, p-values, and R² for each stratification category
  coef_info <- weighted_lm_by_group %>%
    group_by(!!var_sym) %>%
    nest() %>%
    mutate(
      mod = map(data, ~ lm(estimate ~ year, data = .x)),
      tidied = map(mod, ~ tidy(.x) %>% select(term, estimate, p.value)),
      glanced = map(mod, ~ glance(.x) %>% select(r.squared))
    ) %>%
    unnest(cols = tidied) %>%
    filter(term == "year") %>%
    unnest(cols = glanced) %>%
    select(!!var_sym, slope = estimate, p.value, r.squared)

  # Rename for table clarity
  results_table <- coef_info %>%
    dplyr::rename(Stratum = !!var_sym,
           Slope = slope,
           `p-value` = p.value,
           `R²` = r.squared)

  # Create main plot with legend at top
  p <- ggplot(weighted_lm_by_group, aes(x = year, y = estimate, color = !!var_sym)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    labs(
      title = "Change in 'Age' Coefficient Over Years",
      subtitle = paste0("Stratified by ", var_char),
      x = "Year",
      y = "Coefficient of Age",
      color = var_char
    ) +
    theme_minimal() #+
   # theme(legend.position = "top")

  # Extract the legend
  legend <- cowplot::get_legend(p)

  # Remove the legend from the main plot
#  p_no_legend <- p + theme(legend.position = "none")

  # Assign colors to each stratum. 
  # Here we use the same hue palette as ggplot's default for discrete colors:
  unique_strata <- results_table$Stratum
  strata_colors <- setNames(hue_pal()(length(unique_strata)), unique_strata)

  # Make lighter, more transparent versions of the colors for backgrounds
  bg_shades <- alpha(strata_colors, 0.4) # adjust alpha as desired

  # Convert table into a graphical object
#  table_grob <- tableGrob(results_table)
  
  table_grob <- tableGrob(
  results_table,
  theme = ttheme_minimal(base_size = 8) # Make the table text smaller
)

  # table_grob has headers in row 1, data from row 2 onwards.
  # We'll color each data row according to the stratum:
  n_rows <- nrow(results_table)
  for (i in seq_len(n_rows)) {
    # The table's data rows start at row index 2 in table_grob
    row_pos <- i + 1
    # Get the fill color corresponding to this row's stratum
    this_stratum <- results_table$Stratum[i]
    fill_color <- bg_shades[this_stratum]

    # Identify all 'core-bg' grobs in this row and set their fill
    row_indices <- which(table_grob$layout$t == row_pos & table_grob$layout$name == "core-bg")
    for (idx in row_indices) {
      table_grob$grobs[[idx]]$gp$fill <- fill_color
    }
  }

  # Combine legend, table, and plot into one figure
  final_plot <- plot_grid(
    p,
    table_grob,
   # p_no_legend,
    ncol = 1,
    rel_heights = c(1, 0.4) # Adjust heights as needed
  )

  final_plot
}

# Example usage:
stratify_coeff_plot_with_colored_table("happy", "Happiness")
stratify_coeff_plot_with_colored_table("life", "Life Satisfaction")
stratify_coeff_plot_with_colored_table("satfin", "Financial Satisfaction")
stratify_coeff_plot_with_colored_table("sex", "Sex")
stratify_coeff_plot_with_colored_table("race", "Race")
stratify_coeff_plot_with_colored_table("happy", "Happiness")
stratify_coeff_plot_with_colored_table("life", "Life Satisfaction")
stratify_coeff_plot_with_colored_table("polviews", "Political Views")
stratify_coeff_plot_with_colored_table("class", "Class")
stratify_coeff_plot_with_colored_table("satfin", "Financial Satisfaction")
#stratify_srh_yr_plot("region", "Region")
#stratify_coeff_plot_with_colored_table("attend", "Attend")
stratify_coeff_plot_with_colored_table("wrkstat", "wrkstat")
#stratify_srh_yr_plot("hispanic", "Hispanic")
stratify_coeff_plot_with_colored_table("degree", "Degree")
# stratify_coeff_plot_with_colored_table("marital", "Marital Status")
stratify_coeff_plot_with_colored_table("partyid", "Party ID")
#stratify_srh_yr_plot("south", "South")

```

###

```{r}

library(tidyverse)
library(broom)
library(survey)
library(cowplot)
library(gridExtra)
library(scales) # for alpha()

stratify_coeff_plot_with_colored_table_inset <- function(var_name, var_char) {
  var_sym <- rlang::sym(var_name)

  # Compute the weighted linear model results by group
  weighted_lm_by_group <- gss_svy_labeled %>%
    group_by(year, !!var_sym) %>%
    group_map_dfr(~ {
      model <- survey::svyglm(as.numeric(health) ~ age, design = .x)
      tidy(model, conf.int = TRUE)
    }) %>%
    filter(term == "age") %>%
    select(year, !!var_sym, estimate, std.error, conf.low, conf.high, statistic, p.value)

  # Compute slope, p-values, and R² for each stratification category
  coef_info <- weighted_lm_by_group %>%
    group_by(!!var_sym) %>%
    nest() %>%
    mutate(
      mod = map(data, ~ lm(estimate ~ year, data = .x)),
      tidied = map(mod, ~ tidy(.x) %>% select(term, estimate, p.value)),
      glanced = map(mod, ~ glance(.x) %>% select(r.squared))
    ) %>%
    unnest(cols = tidied) %>%
    filter(term == "year") %>%
    unnest(cols = glanced) %>%
    select(!!var_sym, slope = estimate, p.value, r.squared)

  # Rename for table clarity
  results_table <- coef_info %>%
    dplyr::rename(Stratum = !!var_sym,
                  Slope = slope,
                  `p-value` = p.value,
                  `R²` = r.squared) %>% 
    # Round numeric columns to 3 decimal places
    mutate(
      Slope = round(Slope, 6),
     `p-value` = label_scientific(digits = 3)(`p-value`),
      `R²` = round(`R²`, 3)
    )

  # Create main plot
  p <- ggplot(weighted_lm_by_group, aes(x = year, y = estimate, color = !!var_sym)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    labs(
      title = "Change in 'Age' Coefficient Over Years",
      subtitle = paste0("Stratified by ", var_char),
      x = "Year",
      y = "Coefficient of Age",
      color = var_char
    ) +
    theme_minimal() +
    theme(legend.position = "right")

  # Assign colors to each stratum
  unique_strata <- results_table$Stratum
  strata_colors <- setNames(hue_pal()(length(unique_strata)), unique_strata)

  # Make lighter, more transparent versions of the colors for backgrounds
  bg_shades <- alpha(strata_colors, 0.4)

  # Convert table into a graphical object
  table_grob <- tableGrob(
    results_table,
    theme = ttheme_minimal(base_size = 8)
  )

  # Color each data row in the table according to the stratum
  n_rows <- nrow(results_table)
  for (i in seq_len(n_rows)) {
    row_pos <- i + 1
    this_stratum <- results_table$Stratum[i]
    fill_color <- bg_shades[this_stratum]

    row_indices <- which(table_grob$layout$t == row_pos & table_grob$layout$name == "core-bg")
    for (idx in row_indices) {
      table_grob$grobs[[idx]]$gp$fill <- fill_color
    }
  }

  # Overlay the table as an inset in the top-left corner of the main plot
  final_plot <- ggdraw() +
    draw_plot(p) +
    draw_grob(
      table_grob,
      x = 0.55, #0.62  # adjust horizontal position (0 = left, 1 = right)
      y = 0.1,   # adjust vertical position (0 = bottom, 1 = top)
  #    x = 0.15,   # adjust horizontal position (0 = left, 1 = right)
  #    y = 0.82,   # adjust vertical position (0 = bottom, 1 = top)
  #    hjust = 0,  # table anchor horizontally at left
  #    vjust = 1,  # table anchor vertically at top
      width = 0.3, # adjust width
      height = 0.3 # adjust height
    )

  final_plot
}

stratify_coeff_plot_with_colored_table_inset("sex", "Sex")
stratify_coeff_plot_with_colored_table_inset("race", "Race")
stratify_coeff_plot_with_colored_table_inset("happy", "Happiness")
stratify_coeff_plot_with_colored_table_inset("life", "Life Satisfaction")
stratify_coeff_plot_with_colored_table_inset("satfin", "Financial Satisfaction")
stratify_coeff_plot_with_colored_table_inset("class", "Class")
stratify_coeff_plot_with_colored_table_inset("degree", "Degree")

#stratify_srh_yr_plot("region", "Region")
#stratify_coeff_plot_with_colored_table("attend", "Attend")
stratify_coeff_plot_with_colored_table_inset("wrkstat", "wrkstat")
#stratify_srh_yr_plot("hispanic", "Hispanic")

# stratify_coeff_plot_with_colored_table("marital", "Marital Status")
stratify_coeff_plot_with_colored_table_inset("partyid", "Party ID")
#stratify_srh_yr_plot("south", "South")


```


# a
```{r}


# Compute weighted mean health by age group, year, satfin, and sex
weighted_health_by_age <- gss_svy %>%
  group_by(happy, sex, age_group, year) %>%
  summarise(
    mean_health = srvyr::survey_mean(health, na.rm = TRUE),
    .groups = "drop"
  )


ggplot(weighted_health_by_age, aes(x = year, y = mean_health, color = age_group)) +
  geom_line() +
  geom_point() +
  facet_grid(happy ~ sex) +
  # Add annotations for each line within each facet
  # geom_text(
  #   data = coef_info,
  #   inherit.aes = FALSE,
  #   aes(
  #     x = x_text,
  #     y = y_text,
  #     label = sprintf("Slope=%.3e, p=%.3e, R²=%.3f", slope, p.value, r.squared),
  #     color = age_group
  #   ),
  #   hjust = 0,  # left-align at x_text
  #   vjust = 1,  # top-align at y_text
  #   size = 3,
  #   show.legend = FALSE # Don't want a separate legend for text color
  # ) +
  labs(
    title = "Average SRH Per Year by Age Group, Stratified by satfin and Sex",
    subtitle = "GSS Dataset with Survey Weights",
    x = "Year",
    y = "Average Self-Rated Health",
    color = "Age Group"
  ) +
  theme_minimal()

```

Check which variables have enough power for stratification

```{r}

gss_svy %>% 
  group_by(sex, satfin, year) %>% 
  summarize(n = n()) %>% 
  group_by(sex, satfin ) %>% 
  summarize(avg_year = mean(n))

library(dplyr)

# Identify all variables except 'year'
vars <- setdiff(
  names(gss_svy$variables), 
  c("year", "age", "cohort", "wtsscomp", "age_group", "health")
)

# Generate all pairs of variables
all_pairs <- combn(vars, 2, simplify = FALSE)

# For each pair, compute the desired summaries
n_per_overlapping_strata <- bind_rows(lapply(all_pairs, function(pair) {
  var1 <- rlang::sym(pair[1])
  var2 <- rlang::sym(pair[2])
  
  gss_svy %>%
    group_by(!!var1, !!var2, year) %>%
    summarize(n = n(), .groups = "drop") %>%
    group_by(!!var1, !!var2) %>%
    summarize(
      avg_per_year = mean(n),
      min_per_year = min(n),
      .groups = "drop"
    ) %>%
    mutate(var1 = pair[1], var2 = pair[2]) %>%
    select(var1, var2, avg_per_year, min_per_year)
}))

# 'result' now contains a row for each variable pair and their summaries
n_per_overlapping_strata

# Average and min over all

library(dplyr)

vars <- setdiff(names(gss_svy$variables), c("year", "age", "cohort", "wtsscomp", "age_group"))
all_pairs <- combn(vars, 2, simplify = FALSE)

n_per_overlapping_strata_aggregated <- bind_rows(lapply(all_pairs, function(pair) {
  var1 <- rlang::sym(pair[1])
  var2 <- rlang::sym(pair[2])
  
  # Summarize at the category level
  intermediate <- gss_svy %>%
    group_by(!!var1, !!var2, year) %>%
    summarize(n = n(), .groups = "drop") %>%
    group_by(!!var1, !!var2) %>%
    summarize(
      avg_per_year = mean(n),
      min_per_year = min(n),
      .groups = "drop"
    )
  
  # Now aggregate over all category combinations to get a single row per variable pair
  intermediate %>%
    summarize(
      avg_per_year = mean(avg_per_year),
      min_per_year = min(min_per_year)
    ) %>%
    mutate(var1 = pair[1], var2 = pair[2]) %>%
    select(var1, var2, avg_per_year, min_per_year)
}))

n_per_overlapping_strata_aggregated %>% 
  arrange(desc(min_per_year))



```

# Trends of other variables

## Set up data

```{r}

gss_svy <- gss_svy %>%
  mutate(
    health_cat = factor(health, 
                        levels = 1:4,
                        labels = c("Poor", "Fair", "Good", "Excellent")),
    period_cut_6 = as.factor(cut(data_gss$year, 6)),
    period_cut_10 = as.factor(cut(data_gss$year, 10)),
    period_cut_12 = as.factor(cut(data_gss$year, 12)),
    period_10yr = as.factor(
                            cut(
                            year,
                            breaks = c(1973, 1982, 1990, 1998, 2006, 2014, Inf),
                            labels = c("1974-1982", "1982-1990", "1990-1998", 
                                       "1998-2006", "2006-2014", "2014-2022"),
                            right = TRUE
                            )
                          ),
    period_decade = as.factor(
                            cut(
                            year,
                            breaks = c(1973, 1979, 1989, 1999, 2009, 2019, Inf),
                            labels = c("1974-1979", "1980-1989", "1990-1999",
                                       "2000-2009", "2010-2019", "2020-2024"),
                            right = TRUE
                            )
                          ),
    age_group = as.factor( 
                            cut(
                            age,
                            breaks = c(17, 29, 39, 49, 59, 69, Inf),
                            labels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+"),
                            right = TRUE
                          )),
    age_group_small = as.factor( 
                            cut(
                              age,
                              breaks = c(seq(15, 75, by = 5), Inf),  # Define breaks up to 75 and include Inf for the last group
                              labels = c("16-20", "21-25", "26-30", "31-35", "36-40", "41-45", "46-50", "51-55", "56-60", "61-65", "66-70", "71-75", "76+"),
                              right = FALSE  # Makes intervals left-closed, i.e., [x, y)
                            )
                          )
                          ,
    generation = factor(
      case_when(
        cohort >= 1901 & cohort <= 1927 ~ "Greatest (1901-1927)",
        cohort >= 1928 & cohort <= 1945 ~ "Silent (1928-1945)",
        cohort >= 1946 & cohort <= 1964 ~ "Boomers (1946-1964)",
        cohort >= 1965 & cohort <= 1980 ~ "Gen X (1965-1980)",
        cohort >= 1981 & cohort <= 1996 ~ "Millennials (1981-1996)",
        cohort >= 1997 & cohort <= 2012 ~ "Gen Z (1997-2012)",
        TRUE ~ "Other"
      ),
      levels = c(
        "Greatest (1901-1927)",
        "Silent (1928-1945)",
        "Boomers (1946-1964)",
        "Gen X (1965-1980)",
        "Millennials (1981-1996)",
        "Gen Z (1997-2012)"#,
     #   "Other"
      )
    ),
    generation_two_sections = factor(
      case_when(
        generation == "Greatest (1901-1927)" & cohort <= 1914 ~ "Greatest Early (1901-1914)",
        generation == "Greatest (1901-1927)" & cohort > 1914 ~ "Greatest Late (1915-1927)",
        generation == "Silent (1928-1945)" & cohort <= 1936 ~ "Silent Early (1928-1936)",
        generation == "Silent (1928-1945)" & cohort > 1936 ~ "Silent Late (1937-1945)",
        generation == "Boomers (1946-1964)" & cohort <= 1955 ~ "Boomers Early (1946-1955)",
        generation == "Boomers (1946-1964)" & cohort > 1955 ~ "Boomers Late (1956-1964)",
        generation == "Gen X (1965-1980)" & cohort <= 1972 ~ "Gen X Early (1965-1972)",
        generation == "Gen X (1965-1980)" & cohort > 1972 ~ "Gen X Late (1973-1980)",
        generation == "Millennials (1981-1996)" & cohort <= 1988 ~ "Millennials Early (1981-1988)",
        generation == "Millennials (1981-1996)" & cohort > 1988 ~ "Millennials Late (1989-1996)",
        generation == "Gen Z (1997-2012)" & cohort <= 2004 ~ "Gen Z Early (1997-2004)",
        generation == "Gen Z (1997-2012)" & cohort > 2004 ~ "Gen Z Late (2005-2012)",
        TRUE ~ "Other"
      ),
      levels = c(
        "Greatest Early (1901-1914)", "Greatest Late (1915-1927)",
        "Silent Early (1928-1936)", "Silent Late (1937-1945)",
        "Boomers Early (1946-1955)", "Boomers Late (1956-1964)",
        "Gen X Early (1965-1972)", "Gen X Late (1973-1980)",
        "Millennials Early (1981-1988)", "Millennials Late (1989-1996)",
        "Gen Z Early (1997-2004)", "Gen Z Late (2005-2012)"#,
     #   "Other"
      )
    ),
    generation_three_sections = factor(
      case_when(
        generation == "Greatest (1901-1927)" & cohort <= 1910 ~ "Greatest Early (1901-1910)",
        generation == "Greatest (1901-1927)" & cohort > 1910 & cohort <= 1918 ~ "Greatest Mid (1911-1918)",
        generation == "Greatest (1901-1927)" & cohort > 1918 ~ "Greatest Late (1919-1927)",
        generation == "Silent (1928-1945)" & cohort <= 1934 ~ "Silent Early (1928-1934)",
        generation == "Silent (1928-1945)" & cohort > 1934 & cohort <= 1940 ~ "Silent Mid (1935-1940)",
        generation == "Silent (1928-1945)" & cohort > 1940 ~ "Silent Late (1941-1945)",
        generation == "Boomers (1946-1964)" & cohort <= 1951 ~ "Boomers Early (1946-1951)",
        generation == "Boomers (1946-1964)" & cohort > 1951 & cohort <= 1958 ~ "Boomers Mid (1952-1958)",
        generation == "Boomers (1946-1964)" & cohort > 1958 ~ "Boomers Late (1959-1964)",
        generation == "Gen X (1965-1980)" & cohort <= 1970 ~ "Gen X Early (1965-1970)",
        generation == "Gen X (1965-1980)" & cohort > 1970 & cohort <= 1976 ~ "Gen X Mid (1971-1976)",
        generation == "Gen X (1965-1980)" & cohort > 1976 ~ "Gen X Late (1977-1980)",
        generation == "Millennials (1981-1996)" & cohort <= 1986 ~ "Millennials Early (1981-1986)",
        generation == "Millennials (1981-1996)" & cohort > 1986 & cohort <= 1992 ~ "Millennials Mid (1987-1992)",
        generation == "Millennials (1981-1996)" & cohort > 1992 ~ "Millennials Late / Gen Z (1993-2004)",
    #    generation == "Gen Z (1997-2012)" & cohort <= 2002 ~ "Gen Z Early (1997-2002)",
    #    generation == "Gen Z (1997-2012)" & cohort > 2002 & cohort <= 2008 ~ "Gen Z Mid (2003-2008)",
    #    generation == "Gen Z (1997-2012)" & cohort > 2008 ~ "Gen Z Late (2009-2012)",
        TRUE ~ "Other"
      ),
      levels = c(
        "Greatest Early (1901-1910)", "Greatest Mid (1911-1918)", "Greatest Late (1919-1927)",
        "Silent Early (1928-1934)", "Silent Mid (1935-1940)", "Silent Late (1941-1945)",
        "Boomers Early (1946-1951)", "Boomers Mid (1952-1958)", "Boomers Late (1959-1964)",
        "Gen X Early (1965-1970)", "Gen X Mid (1971-1976)", "Gen X Late (1977-1980)",
        "Millennials Early (1981-1986)", "Millennials Mid (1987-1992)", 
        "Millennials Late / Gen Z (1993-2004)"
        #"Millennials Late (1993-1996)",
      #  "Gen Z Early (1997-2002)", "Gen Z Mid (2003-2008)", "Gen Z Late (2009-2012)" #,
      #  "Other"
      )
    )
  )
  


```


## Average variable per year by age group

```{r}

avg_var_per_year_by_age_plot <- function(var_name, var_char) {
  var_sym <- rlang::sym(var_name)
  
  # Compute weighted mean health by the variable of interest, age group, and year
  gss_svy %>%
    group_by(age_group, year) %>%
    summarise(
      mean_var = srvyr::survey_mean(as.numeric(!!var_sym), na.rm = TRUE),
      .groups = "drop"
    ) %>%
    ggplot(aes(x = year, y = mean_var, color = age_group)) +
    geom_line() +
    geom_point() +
    labs(
      title = paste0("Mean ", var_char, " Per Year by Age Group"),
      subtitle = "GSS Dataset with Survey Weights",
      x = "Year",
      y = paste0("Mean ", var_char),
      color = "Age Group"
    ) +
    theme_minimal()
}

avg_var_per_year_by_age_plot("health", "SRH")
avg_var_per_year_by_age_plot("happy", "Happiness")
avg_var_per_year_by_age_plot("satfin", "Financial Satisfaction")
avg_var_per_year_by_age_plot("life", "Life Satisfaction")
avg_var_per_year_by_age_plot("class", "Self-Assessed Class")
avg_var_per_year_by_age_plot("polviews", "Political Views")
avg_var_per_year_by_age_plot("partyid", "Party ID")


```

# Age Coeff Over Time For Different Variables

```{r}

var_age_coeff_over_time_plot <- function(var_name, var_char) {
  var_sym <- rlang::sym(var_name)
  formula <- as.formula(paste0("as.numeric(", var_name, ") ~ age"))

  weighted_lm_by_group <- gss_svy %>%
    group_by(year) %>%
    group_map_dfr(~ {
      model <- survey::svyglm(formula, design = .x)
      tidy(model, conf.int = TRUE)
    }) %>%
    filter(term == "age") %>%
    select(year, estimate, std.error, conf.low, conf.high, statistic, p.value)

  coef_info <- weighted_lm_by_group %>%
    summarise(
      slope = lm(estimate ~ year)$coefficients["year"],
      p.value = summary(lm(estimate ~ year))$coefficients["year", "Pr(>|t|)"],
      r.squared = summary(lm(estimate ~ year))$r.squared
    )

  text_positions <- weighted_lm_by_group %>%
    summarize(
      x_text = min(year),
      y_text = min(estimate) - 0.01,
      .groups = "drop"
    )

  ggplot(weighted_lm_by_group, aes(x = year, y = estimate)) +
    geom_point() +
    geom_smooth(method = "lm", se = FALSE) +
    geom_text(
      data = coef_info,
      aes(
        x = text_positions$x_text,
        y = text_positions$y_text,
        label = sprintf("Slope = %.3e\np = %.3e\nR² = %.3f", slope, p.value, r.squared)
      ),
      hjust = 0, vjust = 0, size = 3
    ) +
    labs(
      title = "Change in 'Age' Coefficient Over Years",
      subtitle = paste0("Outcome: ", var_char),
      x = "Year",
      y = "Coefficient of Age"
    ) +
    theme_minimal()
}

var_age_coeff_over_time_plot("health", "SRH")
var_age_coeff_over_time_plot("happy", "Happiness")
var_age_coeff_over_time_plot("satfin", "Financial Satisfaction")
var_age_coeff_over_time_plot("life", "Life Satisfaction")
var_age_coeff_over_time_plot("class", "Self-Assessed Class")
var_age_coeff_over_time_plot("polviews", "Political Views")
var_age_coeff_over_time_plot("partyid", "Party ID")

```

# Both Plots

```{r}
diff_var_plots <- function(var_name, var_char){
  gridExtra::grid.arrange(avg_var_per_year_by_age_plot(var_name, var_char), 
                        var_age_coeff_over_time_plot(var_name, var_char),
                        nrow = 1, 
                        ncol = 2)
}


diff_var_plots("health", "SRH")
diff_var_plots("happy", "Happiness")
diff_var_plots("satfin", "Financial Satisfaction")
diff_var_plots("life", "Life Satisfaction")
diff_var_plots("class", "Self-Perceived Class")
diff_var_plots("polviews", "Political Views")
diff_var_plots("partyid", "Party ID")
```


```{r, eval = FALSE}
gridExtra::grid.arrange(avg_var_per_year_by_age_plot("health", "SRH"), 
                        var_age_coeff_over_time_plot("health", "SRH"),
                        avg_var_per_year_by_age_plot("happy", "Happiness"),
                        var_age_coeff_over_time_plot("happy", "Happiness"),
                        avg_var_per_year_by_age_plot("satfin", "Financial Satisfaction"),
                        var_age_coeff_over_time_plot("satfin", "Financial Satisfaction"),
                        avg_var_per_year_by_age_plot("life", "Life Satisfaction"),
                        var_age_coeff_over_time_plot("life", "Life Satisfaction"),
                        avg_var_per_year_by_age_plot("class", "Self-Assessed Class"),
                        var_age_coeff_over_time_plot("class", "Self-Assessed Class"),
                        avg_var_per_year_by_age_plot("polviews", "Political Views"),
                        var_age_coeff_over_time_plot("polviews", "Political Views"),
                        avg_var_per_year_by_age_plot("partyid", "Party ID"),
                        var_age_coeff_over_time_plot("partyid", "Party ID"),
                        nrow = 7, 
                        ncol = 2)

```

# side by side

```{r}

side_by_side_srh_coeff_stratified_plot <- function(var_name, var_char){
 # var_sym <- rlang::sym(var_name)
gridExtra::grid.arrange(stratify_srh_yr_plot(var_name, var_char), 
                        stratify_coeff_plot_with_colored_table_inset(var_name, var_char),
                        nrow = 1)}


side_by_side_srh_coeff_stratified_plot("sex", "Sex")
side_by_side_srh_coeff_stratified_plot("race", "Race")
side_by_side_srh_coeff_stratified_plot("happy", "Happiness")
side_by_side_srh_coeff_stratified_plot("life", "Life Satisfaction")
side_by_side_srh_coeff_stratified_plot("satfin", "Financial Satisfaction")
side_by_side_srh_coeff_stratified_plot("class", "Self-Perceived Class")
side_by_side_srh_coeff_stratified_plot("degree", "Degree")

gridExtra::grid.arrange(
  stratify_srh_yr_plot("sex", "Sex"),
  stratify_srh_yr_plot("race", "Race"),
  stratify_srh_yr_plot("happy", "Happiness"),
  stratify_srh_yr_plot("life", "Life Satisfaction"),
  stratify_srh_yr_plot("satfin", "Financial Satisfaction"),
  stratify_srh_yr_plot("class", "Self-Perceived Class"),
  nrow = 2)

gridExtra::grid.arrange(
  stratify_coeff_plot_with_colored_table_inset("sex", "Sex"),
  stratify_coeff_plot_with_colored_table_inset("race", "Race"),
  stratify_coeff_plot_with_colored_table_inset("happy", "Happiness"),
  stratify_coeff_plot_with_colored_table_inset("life", "Life Satisfaction"),
  stratify_coeff_plot_with_colored_table_inset("satfin", "Financial Satisfaction"),
  stratify_coeff_plot_with_colored_table_inset("class", "Self-Perceived Class"),
  nrow = 2)




```


# idk

```{r, eval = FALSE}

gss_svy %>%
    group_by(age_group, year) %>%
    summarise(
      mean_var = srvyr::survey_mean(as.numeric(partyid), na.rm = TRUE),
      .groups = "drop"
    ) %>%
    ggplot(aes(x = year, y = mean_var, color = age_group)) +
    geom_line() +
    geom_point() +
    labs(
      title = paste0("Mean ", "Party ID", " Per Year by Age Group"),
      subtitle = "GSS Dataset with Survey Weights",
      x = "Year",
      y = paste0("Mean ", "Party ID"),
      color = "Age Group"
    ) +
    theme_minimal()


gss_svy %>%
    filter(age_group != "70+") %>% 
    group_by(age_group, year) %>%
    summarise(
      mean_var = srvyr::survey_mean(as.numeric(partyid), na.rm = TRUE),
      .groups = "drop"
    ) %>%
    ggplot(aes(x = year, y = mean_var, color = age_group)) +
    geom_line() +
    geom_point() +
    labs(
      title = paste0("Mean ", "Party ID", " Per Year by Age Group"),
      subtitle = "GSS Dataset with Survey Weights",
      x = "Year",
      y = paste0("Mean ", "Party ID"),
      color = "Age Group"
    ) +
    theme_minimal()

gss_svy %>%
    filter(age_group == "70+") %>% 
    group_by(age_group, year) %>%
    summarise(
      mean_var = srvyr::survey_mean(as.numeric(partyid), na.rm = TRUE),
      .groups = "drop"
    ) %>%
    ggplot(aes(x = year, y = mean_var, color = age_group)) +
    geom_line() +
    geom_point() +
    labs(
      title = paste0("Mean ", "Party ID", " Per Year by Age Group"),
      subtitle = "GSS Dataset with Survey Weights",
      x = "Year",
      y = paste0("Mean ", "Party ID"),
      color = "Age Group"
    ) +
    theme_minimal()

```

# More Visualizations


## Age x Period bar chart (cohort highlighted)

```{r}

library(APCtools)
library(dplyr)    # general data handling
library(mgcv)     # estimation of generalized additive regression models (GAMs)
library(ggplot2)  # data visualization
library(ggpubr)   # arranging multiple ggplots in a grid with ggarrange()

# set the global theme of all plots
theme_set(theme_minimal())

plot_variable(dat = data_gss %>% 
                mutate(period = year) %>% 
                mutate(health_cat = as.factor(health)), 
              y_var = "health_cat", apc_dimension = "period")

plot_densityMatrix(dat                 = data_gss,
                   y_var               = "health",
                   age_groups          = age_groups,
                   period_groups       = period_groups,
                   highlight_diagonals = list("born 1950 - 1959" = 8,
                                              "born 1970 - 1979" = 10),
                   log_scale           = TRUE)

```


## Distribution of Health Categories by Year and Age Group

```{r}
# Calculate weighted proportions of each health category by, say, age_group and year
health_props <- gss_svy %>%
  group_by(age_group, period_decade, health_cat) %>%
  summarize(
    prop = survey_prop(na.rm = TRUE, vartype = "se")
  )


# health_props now contains weighted proportions of each health category for each (age_group, year)

# Plotting these as a stacked bar chart, for instance, by year and age_group:
ggplot(health_props %>% mutate(health_cat = fct_rev(health_cat)),
       aes(x = factor(period_decade), y = prop, fill = health_cat)) +
  geom_col(position = "fill") +
  facet_wrap(~ age_group) +
  scale_y_continuous(labels = scales::percent) +
  labs(
    title = "Distribution of Health Categories by Year and Age Group (Weighted)",
    x = "Year",
    y = "Proportion of Respondents",
    fill = "Health"
  ) +
  theme_minimal()
```


# More

```{r}

plot_variable(dat = data_gss %>% mutate(period = year), y_var = "health_cat", apc_dimension = "period")

age_groups    <- list(c(80,89),c(70,79),c(60,69),c(50,59),
                      c(40,49),c(30,39),c(20,29))
period_groups <- list(c(1970,1979),c(1980,1989),c(1990,1999),
                      c(2000,2009),c(2010,2019))

plot_densityMatrix(dat              = travel,
                   y_var            = "mainTrip_distance",
                   age_groups       = age_groups,
                   period_groups    = period_groups,
                   log_scale        = TRUE)

plot_densityMatrix(dat              = data_gss,
                   y_var            = "health",
                   age_groups       = age_groups,
                   period_groups    = period_groups,
                   log_scale        = TRUE)

plot_densityMatrix(dat                 = data_gss %>% mutate(period = year),
                   y_var               = "health",
                   age_groups          = age_groups,
                   period_groups       = period_groups,
                   highlight_diagonals = list("born 1950 - 1959" = 8,
                                              "born 1970 - 1979" = 10))

plot_APCheatmap(dat            = data_gss %>% mutate(period = year),
                y_var          = "health",
                y_var_logScale = FALSE,
              #  bin_heatmap    = FALSE,
                markLines_list = list(cohort = c(1900,1920,1939,1946,
                                                 1966,1982,1994)))

```

```{r}


library(tidyverse)

library(srvyr)
library(dplyr)

# Summarize mean SRH by sex and race
gss_svy_labeled %>%
  mutate(health = as.numeric(health)) %>% 
  mutate(year = cut(year, breaks = 10)) %>% 
  group_by(year, sex, race) %>%
  summarize(
    mean_health = survey_mean(health, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  ggplot(aes(x = year, y = mean_health, fill = sex)) +
  geom_bar(stat = "identity", position = "dodge") +
  facet_wrap(~as.factor(race)) #, scales = "free_y") +
  labs(
    title = "Self-Rated Health by Sex and Race",
    x = "Year",
    y = "Mean SRH",
    fill = "Sex"
  ) +
  theme_minimal() 

# Summarize mean SRH by financial and life satisfaction
gss_svy_labeled %>%
  mutate(health = as.numeric(health)) %>% 
  group_by(year, satfin, life) %>%
  summarize(
    mean_health = survey_mean(health, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  ggplot(aes(x = year, y = mean_health, color = as.factor(satfin))) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  facet_wrap(~life, scales = "free_y") +
  labs(
    title = "Self-Rated Health by Financial and Life Satisfaction",
    x = "Year",
    y = "Mean SRH",
    color = "Financial Satisfaction"
  ) +
  theme_minimal() +
  scale_color_brewer(palette = "Set2")

gss_svy_labeled %>%
  mutate(health = as.numeric(health)) %>% 
  group_by(year, sex, race) %>%
  summarize(
    mean_health = survey_mean(health, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  ggplot(aes(x = year, y = mean_health, color = race)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  facet_wrap(~sex, scales = "free_y") +
  labs(
    title = "Self-Rated Health by Race and Sex",
    x = "Year",
    y = "Mean SRH",
    color = "Race"
  ) +
  theme_minimal() +
  scale_color_brewer(palette = "Set2")


gss_svy_labeled %>%
  mutate(health = as.numeric(health)) %>% 
  group_by(year) %>%
  summarize(
    mean_health = survey_mean(health, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  ggplot(aes(x = year, y = mean_health)) +
  geom_line(size = 1, color = "cornflowerblue") +
  geom_point(size = 2, color = "darkblue") +
  labs(
    title = "Self-Rated Health",
    x = "Year",
    y = "Mean SRH",
    subtitle = "GSS Dataset"
  ) +
  theme_minimal() +
  scale_color_brewer(palette = "Set2")






gss_svy_labeled %>%
  mutate(health = as.numeric(health)) %>% 
  group_by(age) %>%
  summarize(
    mean_health = survey_mean(health, na.rm = TRUE),
    .groups = "drop"
  ) %>% 
  ggplot(aes(x = age, y = mean_health)) +
  geom_line(size = 1, color = "cornflowerblue") +
  geom_point(size = 2, color = "darkblue") +
  labs(
    title = "Self-Rated Health",
    x = "Age",
    y = "Mean SRH",
    subtitle = "GSS Dataset"
  ) +
  theme_minimal() +
  scale_color_brewer(palette = "Set2")



```

