---
title: "Untitled"
author: "Christine Lucille Kuryla"
date: "2025-02-26"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

---
title: "Determinants of Self-Rated Health Across Age Groups and Time"
author: "NHANES Data Analysis"
date: "`r Sys.Date()`"
output:
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    theme: cosmo
    highlight: tango
    code_folding: show
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,                  # Show code
  warning = FALSE,              # Hide warnings
  message = FALSE,              # Hide messages
  fig.width = 10,               # Default figure width
  fig.height = 6,               # Default figure height
  fig.align = "center",         # Center figures
  comment = "#>",               # Change comment symbol
  out.width = "100%"            # Scale output width
)

# Load required packages
library(tidyverse)
library(ggplot2)
library(cowplot)
```

# 1. Introduction and Research Question

This analysis explores how different factors influence self-rated health (SRH) across different age groups and over time. The main research question is:

**Do people of different ages rate their health differently for different reasons, and has this changed over time?**

Self-rated health is widely used as a measure of overall health status and has been shown to predict mortality and other health outcomes. Understanding age differences in SRH determinants can provide insights into how health perceptions are formed throughout the lifespan and how they might be addressed through age-appropriate interventions.

# 2. Data Preparation and Exploration

## 2.1 Data Structure Inspection

First, we'll examine the structure of the NHANES dataset to understand what variables are available and their completeness.

```{r inspect_data}
# Function to inspect data structure
inspect_data_structure <- function(data) {
  # Check for key variables
  has_year <- "year" %in% names(data)
  has_age_group <- "age_group" %in% names(data)
  has_srh <- "srh" %in% names(data)
  
  # Check variable types
  year_class <- if(has_year) class(data$year) else "Not found"
  age_group_class <- if(has_age_group) class(data$age_group) else "Not found"
  srh_class <- if(has_srh) class(data$srh) else "Not found"
  
  # Check for unique values
  year_values <- if(has_year) sort(unique(data$year)) else NULL
  age_group_values <- if(has_age_group) unique(data$age_group) else NULL
  srh_values <- if(has_srh) sort(unique(data$srh)) else NULL
  
  # Summarize findings
  cat("Data Structure Check:\n")
  cat("--------------------\n")
  cat("Year variable present:", has_year, "- Class:", year_class, "\n")
  cat("Age group variable present:", has_age_group, "- Class:", age_group_class, "\n")
  cat("SRH variable present:", has_srh, "- Class:", srh_class, "\n\n")
  
  cat("Unique values:\n")
  cat("Year:", paste(year_values, collapse=", "), "\n")
  cat("Age groups:", paste(age_group_values, collapse=", "), "\n")
  cat("SRH values:", paste(srh_values, collapse=", "), "\n\n")
  
  # Return a summary of the structure
  return(list(
    has_required_vars = has_year & has_age_group & has_srh,
    year_values = year_values,
    age_group_values = age_group_values,
    srh_values = srh_values
  ))
}

# Inspect the NHANES data structure
# Run this after loading your data: structure_check <- inspect_data_structure(data_nhanes)
```

## 2.2 Filtering Variables by Missingness

To ensure analysis quality, we'll filter variables based on their completeness, only including those with less than 50% missing values.

```{r filter_variables}
# Function to filter variables based on missingness
check_variable_completeness <- function(data, threshold = 0.5) {
  # Calculate percentage of missing values for each variable
  missing_percentages <- sapply(data, function(x) mean(is.na(x)))
  
  # Identify variables with less than threshold% missing
  complete_enough <- names(missing_percentages[missing_percentages < threshold])
  
  # Check key variables that must be included regardless of missingness
  key_vars <- c("year", "age_group", "srh")
  missing_key_vars <- key_vars[!key_vars %in% complete_enough]
  if (length(missing_key_vars) > 0) {
    warning("Key variables are missing too much data: ", 
           paste(missing_key_vars, collapse=", "))
  }
  
  return(list(
    complete_vars = complete_enough,
    missing_percentages = missing_percentages
  ))
}

# Function to categorize variables into health domains
categorize_health_variables <- function(var_names) {
  # Define domains and patterns
  domains <- list(
    "Health Conditions" = c("heart_dis", "stroke", "diabetes", "asthma", 
                           "arthritis", "lung_dis", "cancer", "chf", "angina", "mi"),
    
    "Biomarkers" = c("bmi", "body_fat", "sbp", "dbp", "glu", "hba1c", 
                     "chol", "hdl", "ldl", "crp", "waist", "weight", "height"),
    
    "Mental Health" = c("phq9", "ds_score", "depression"),
    
    "Functional Status" = c("grip_str", "fev1", "fvc", "pef"),
    
    "Lifestyle" = c("smoking", "smoke", "met_", "physical", "activity"),
    
    "Socioeconomic" = c("educ", "pir", "income", "marital", "hc_access")
  )
  
  # Initialize result list with "Other" category
  categorized <- list("Other" = character())
  
  # For each variable, check which domain it belongs to
  for (var in var_names) {
    assigned <- FALSE
    
    # Check each domain
    for (domain_name in names(domains)) {
      # Check if variable contains any pattern from this domain
      patterns <- domains[[domain_name]]
      if (any(sapply(patterns, function(p) grepl(p, var, ignore.case = TRUE)))) {
        # If categorized list doesn't have this domain yet, create it
        if (!domain_name %in% names(categorized)) {
          categorized[[domain_name]] <- character()
        }
        
        # Add variable to domain
        categorized[[domain_name]] <- c(categorized[[domain_name]], var)
        assigned <- TRUE
        break  # Only assign to one domain
      }
    }
    
    # If not assigned to any domain, add to "Other"
    if (!assigned) {
      categorized[["Other"]] <- c(categorized[["Other"]], var)
    }
  }
  
  return(categorized)
}
```

# 3. Main Analysis Function

This function is the heart of our analysis. It:
1. Filters out variables with too much missing data
2. Categorizes them into health domains
3. Runs regression models to examine their relationship with SRH across age groups and time

```{r main_analysis}
# Main analysis with filtered variables
run_filtered_analysis <- function(data_nhanes, missingness_threshold = 0.5, 
                                min_obs_per_group = 30) {
  # Step 1: Check data completeness
  completeness <- check_variable_completeness(data_nhanes, threshold = missingness_threshold)
  
  # Step 2: Categorize variables by domain
  categorized_vars <- categorize_health_variables(completeness$complete_vars)
  
  # Print summary of available variables by domain
  cat("Available variables by domain (after filtering for completeness):\n")
  for (domain in names(categorized_vars)) {
    vars <- categorized_vars[[domain]]
    if (length(vars) > 0) {
      cat(domain, ":", length(vars), "variables\n")
      # Print first few variables as examples
      cat("  Examples:", paste(head(vars, 5), collapse=", "), 
          if(length(vars) > 5) "..." else "", "\n")
    }
  }
  
  # Step 3: Select predictors for analysis
  # Exclude technical/identification variables and outcome variable
  exclude_patterns <- c("^id", "^X", "^\\.\\.\\.", "^SDMV", "^WT", "^RID", "^srh")
  
  # Build list of predictors to analyze
  predictors <- c()
  for (domain in names(categorized_vars)) {
    if (domain != "Other") {  # Skip the "Other" category
      domain_vars <- categorized_vars[[domain]]
      
      # Filter out excluded patterns
      domain_vars <- domain_vars[!sapply(domain_vars, function(v) {
        any(sapply(exclude_patterns, function(p) grepl(p, v)))
      })]
      
      # Add remaining variables to predictors
      predictors <- c(predictors, domain_vars)
    }
  }
  
  # Make sure key predictor variables are included regardless of category
  key_predictors <- c("heart_dis", "stroke", "diabetes", "bmi", "smoking", "educ", "pir")
  key_predictors <- key_predictors[key_predictors %in% names(data_nhanes)]
  predictors <- unique(c(predictors, key_predictors))
  
  # Remove the outcome variable from predictors
  predictors <- predictors[!predictors %in% c("srh", "srh_factor")]
  
  cat("\nSelected", length(predictors), "predictors for analysis\n")
  
  # Step 4: Run the analysis
  # Initialize results list
  all_results <- list()
  
  # Get unique years and age groups
  years <- sort(unique(data_nhanes$year))
  age_groups <- unique(data_nhanes$age_group)
  
  # Loop through predictors
  for (pred in predictors) {
    cat("Analyzing predictor:", pred, "\n")
    
    # Skip if predictor not in dataset
    if (!pred %in% names(data_nhanes)) {
      cat("  Predictor not found in dataset. Skipping.\n")
      next
    }
    
    # Check variable type to determine model approach
    is_categorical <- is.factor(data_nhanes[[pred]]) || is.character(data_nhanes[[pred]])
    
    # Run a model for each year and age group
    for (yr in years) {
      for (ag in age_groups) {
        # Subset data
        subset_data <- data_nhanes[data_nhanes$year == yr & data_nhanes$age_group == ag, ]
        
        # Skip if too few observations
        if (nrow(subset_data) < min_obs_per_group) {
          next
        }
        
        # Skip if predictor has too many NAs in this subset
        if (mean(is.na(subset_data[[pred]])) > 0.5) {
          next
        }
        
        # Try to fit model
        tryCatch({
          # Create formula
          formula_str <- paste("srh ~", pred)
          
          # Run linear model
          model <- lm(formula_str, data = subset_data)
          
          # Extract coefficients
          coefs <- summary(model)$coefficients
          
          # For continuous predictors
          if (!is_categorical && pred %in% rownames(coefs)) {
            # Add to results
            all_results[[length(all_results) + 1]] <- data.frame(
              variable = pred,
              predictor = pred,
              year = yr,
              age_group = ag,
              coefficient = coefs[pred, "Estimate"],
              se = coefs[pred, "Std. Error"],
              t_value = coefs[pred, "t value"],
              p_value = coefs[pred, "Pr(>|t|)"],
              n_obs = nrow(subset_data),
              pct_missing = mean(is.na(subset_data[[pred]])) * 100,
              level = NA_character_,
              variable_type = "continuous"
            )
          } else if (is_categorical || grepl("TRUE$", rownames(coefs)[2])) {
            # For categorical predictors, extract all level coefficients
            # Skip the intercept (first row)
            for (i in 2:nrow(coefs)) {
              level_name <- rownames(coefs)[i]
              # Extract the level name from the coefficient name
              if (grepl("^TRUE$", level_name)) {
                level_display <- "TRUE"  # Binary variable
              } else {
                level_display <- gsub(paste0("^", pred), "", level_name)
              }
              
              all_results[[length(all_results) + 1]] <- data.frame(
                variable = pred,
                predictor = pred,
                year = yr,
                age_group = ag,
                coefficient = coefs[i, "Estimate"],
                se = coefs[i, "Std. Error"],
                t_value = coefs[i, "t value"],
                p_value = coefs[i, "Pr(>|t|)"],
                n_obs = nrow(subset_data),
                pct_missing = mean(is.na(subset_data[[pred]])) * 100,
                level = level_display,
                variable_type = "categorical"
              )
            }
          }
        }, error = function(e) {
          # Just skip this combination if there's an error
        })
      }
    }
  }
  
  # Combine all results
  if (length(all_results) > 0) {
    results_df <- do.call(rbind, all_results)
    
    # Add domain classification
    results_df$domain <- sapply(results_df$variable, function(v) {
      for (domain in names(categorized_vars)) {
        if (v %in% categorized_vars[[domain]]) {
          return(domain)
        }
      }
      return("Other")
    })
    
    cat("\nAnalysis complete. Generated", nrow(results_df), "results across", 
        length(unique(results_df$variable)), "variables.\n")
    
    return(results_df)
  } else {
    cat("No results were generated.\n")
    return(NULL)
  }
}
```

# 4. Visualization Functions

## 4.1 Individual Predictor Plots

This function creates beautiful visualizations showing how a specific predictor affects SRH across different age groups over time.

```{r individual_plot_function}
# Function to create individual plots for each predictor
plot_predictor_by_age <- function(results, 
                                 predictor_name,
                                 title = NULL,
                                 ylim = NULL,
                                 significance_size = TRUE,
                                 line_size = 1,
                                 point_size = 3) {
  
  # Filter results for the specified predictor
  pred_data <- results %>%
    filter(variable == predictor_name | predictor == predictor_name) %>%
    filter(!is.na(coefficient), !is.na(age_group), !is.na(year))
  
  # Check if we have data
  if(nrow(pred_data) == 0) {
    stop(paste("No data found for predictor:", predictor_name))
  }
  
  # Create title if not provided
  if(is.null(title)) {
    title <- paste("Effect of", predictor_name, "on Self-Rated Health by Age Group")
  }
  
  # Define colors for age groups using a colorblind-friendly palette
  # age_colors <- c(
  #   "18-29" = "#E69F00", 
  #   "30-39" = "#56B4E9", 
  #   "40-49" = "#009E73", 
  #   "50-59" = "#F0E442", 
  #   "60-69" = "#0072B2", 
  #   "70+" = "#D55E00"
  # )
  
  age_colors <- c(
    "18-29" = "#D55E00", 
    "30-39" = "#E69F00", 
    "40-49" = "#F0E442", 
    "50-59" = "#009E73", 
    "60-69" = "#56B4E9", 
    "70+" = "#0072B2"
  )
  
  # Create the plot
  p <- ggplot(pred_data, 
             aes(x = year, y = coefficient, color = age_group, group = age_group)) +
    # Add horizontal line at y=0 for reference
    geom_hline(yintercept = 0, linetype = "dashed", color = "gray70") +
    # Add lines connecting points within each age group
    geom_line(size = line_size) +
    # Add points, with option for significance-based sizing
    {if(significance_size) 
      geom_point(aes(size = ifelse(p_value < 0.05, "p < 0.05", "p ≥ 0.05")))
     else 
      geom_point(size = point_size)} +
    # Set colors for age groups
    scale_color_manual(values = age_colors) +
    # Set point sizes if using significance
    {if(significance_size)
      scale_size_manual(values = c("p < 0.05" = point_size, "p ≥ 0.05" = point_size/2))} +
    # Add labels
    labs(
      title = title,
      x = "Year",
      y = "Coefficient (Effect on SRH)",
      color = "Age Group",
      size = if(significance_size) "Significance" else NULL
    ) +
    # Set y-axis limits if provided
    {if(!is.null(ylim)) coord_cartesian(ylim = ylim)} +
    # Use minimal theme
    theme_minimal() +
    # Customize appearance
    theme(
      legend.position = "bottom",
      panel.grid.minor = element_line(color = "gray95"),
      panel.grid.major = element_line(color = "gray90"),
      plot.title = element_text(face = "bold", size = 14),
      axis.title = element_text(size = 12),
      axis.text = element_text(size = 10)
    )
  
  return(p)
}
```

## 4.2 Multiple Predictors and Grid Layouts

These functions allow you to analyze multiple predictors at once and arrange them in grids for comparison.

```{r multiple_plots_function}
# Function to create plots for multiple predictors
plot_multiple_predictors <- function(results, 
                                    predictor_names,
                                    save_plots = FALSE,
                                    output_dir = "predictor_plots",
                                    width = 10,
                                    height = 6,
                                    common_ylim = NULL) {
  
  # Create output directory if saving plots
  if(save_plots) {
    if(!dir.exists(output_dir)) {
      dir.create(output_dir)
    }
  }
  
  # Create list to store plots
  plot_list <- list()
  
  # Create plot for each predictor
  for(pred in predictor_names) {
    # Try to create plot
    tryCatch({
      # Generate plot
      p <- plot_predictor_by_age(results, 
                               predictor_name = pred,
                               ylim = common_ylim)
      
      # Store plot in list
      plot_list[[pred]] <- p
      
      # Save plot if requested
      if(save_plots) {
        filename <- file.path(output_dir, paste0(pred, "_plot.png"))
        ggsave(filename, p, width = width, height = height, dpi = 300)
        cat("Saved plot for", pred, "to", filename, "\n")
      }
      
      # Print plot if not saving
      if(!save_plots) {
        print(p)
      }
    }, error = function(e) {
      warning(paste("Error creating plot for", pred, ":", e$message))
    })
  }
  
  # Return list of plots
  return(plot_list)
}

# Function to arrange multiple plots on a single page
arrange_predictor_plots <- function(results, 
                                   predictor_names,
                                   ncol = 2,
                                   common_ylim = NULL,
                                   title_size = 10) {
  
  # Create list to store plots
  plot_list <- list()
  
  # Create plot for each predictor with smaller titles
  for(pred in predictor_names) {
    # Try to create plot
    tryCatch({
      # Generate plot with simplified title for grid
      p <- plot_predictor_by_age(results, 
                               predictor_name = pred,
                               title = pred,  # just use predictor name as title
                               ylim = common_ylim) +
        theme(
          plot.title = element_text(size = title_size, face = "bold"),
          legend.position = "none"  # Remove legend from individual plots
        )
      
      # Store plot in list
      plot_list[[pred]] <- p
      
    }, error = function(e) {
      warning(paste("Error creating plot for", pred, ":", e$message))
    })
  }
  
  # Create a combined legend
  # Extract legend from one of the plots
  if(length(plot_list) > 0) {
    # Get name of first successful plot
    first_plot_name <- names(plot_list)[1]
    
    # Create a version with legend
    legend_plot <- plot_predictor_by_age(results, 
                                       first_plot_name, 
                                       ylim = common_ylim)
    
    # Extract the legend
    legend <- cowplot::get_legend(legend_plot)
    
    # Arrange plots in a grid with shared legend at bottom
    arranged_plots <- cowplot::plot_grid(
      plotlist = plot_list,
      ncol = ncol,
      align = "hv"
    )
    
    # Add the legend at the bottom
    final_plot <- cowplot::plot_grid(
      arranged_plots,
      legend,
      ncol = 1,
      rel_heights = c(1, 0.1)
    )
    
    return(final_plot)
  } else {
    stop("No plots were successfully created")
  }
}
```

## 4.3 Domain Comparison

This function compares the relative importance of different health domains (conditions, biomarkers, socioeconomic factors) for SRH.

```{r domain_comparison_function}
# Function to compare domain importance by age groups
compare_domain_importance <- function(results, min_vars_per_domain = 2) {
  # Remove rows with missing coefficients
  domain_data <- results %>%
    filter(!is.na(coefficient), !is.na(domain))
  
  # Count variables per domain
  vars_per_domain <- domain_data %>%
    group_by(domain) %>%
    summarize(
      n_vars = n_distinct(variable),
      .groups = "drop"
    ) %>%
    filter(n_vars >= min_vars_per_domain)
  
  # Keep only domains with sufficient variables
  domain_data <- domain_data %>%
    filter(domain %in% vars_per_domain$domain)
  
  if (nrow(domain_data) == 0) {
    stop("No domains have sufficient variables for comparison.")
  }
  
  # Calculate average absolute coefficient by domain, age group, and year
  domain_importance <- domain_data %>%
    group_by(domain, age_group, year) %>%
    summarize(
      mean_abs_coef = mean(abs(coefficient), na.rm = TRUE),
      n_vars = n_distinct(variable),
      n_significant = sum(p_value < 0.05, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    mutate(
      prop_significant = n_significant / n_vars
    )
  
  # Create the plot
  p <- ggplot(domain_importance, 
             aes(x = year, y = mean_abs_coef, color = domain, group = domain)) +
    geom_line(size = 1) +
    geom_point(aes(size = prop_significant)) +
    facet_wrap(~ age_group) +
    labs(
      title = "Relative Importance of Different Domains for Self-Rated Health",
      subtitle = "Size of point indicates proportion of significant predictors",
      x = "Year",
      y = "Mean Absolute Coefficient",
      color = "Domain",
      size = "Prop. Significant"
    ) +
    theme_minimal() +
    theme(
      legend.position = "bottom",
      strip.background = element_rect(fill = "lightgrey"),
      strip.text = element_text(face = "bold")
    )
  
  return(list(
    plot = p,
    data = domain_importance
  ))
}
```

# 5. Running the Analysis

Now we can run our full analysis on the NHANES data. First, we filter variables with more than 50% missing data, then analyze relationships between each predictor and SRH by age group over time.

```{r run_analysis, eval=FALSE}
# Run the analysis (uncomment to execute)
# results <- run_filtered_analysis(data_nhanes, missingness_threshold = 0.5)

# Example of what results would look like
# glimpse(results)
```

# 6. Visualizing Results by Domain

## 6.1 Health Conditions

First, let's examine how different health conditions affect self-rated health across age groups and over time.

```{r health_conditions, eval=FALSE}
# Define health condition predictors
health_predictors <- c("heart_dis", "stroke", "diabetes", "chf", "angina", "mi", "lung_dis")

# Create a grid of plots for all health conditions
health_grid <- arrange_predictor_plots(
  results, 
  predictor_names = health_predictors,
  ncol = 3,
  common_ylim = c(-2.5, 1.5)  # Use common y-axis for comparison
)

# Display the grid
print(health_grid)

# Create individual plots for key conditions
diabetes_plot <- plot_predictor_by_age(
  results, 
  "diabetes",
  title = "Effect of Diabetes on Self-Rated Health by Age Group"
)

print(diabetes_plot)
```

## 6.2 Biomarkers

Next, let's examine how biomarkers like BMI, blood pressure, and cholesterol affect SRH across age groups.

```{r biomarkers, eval=FALSE}
# Define biomarker predictors
biomarker_predictors <- c("bmi", "sbp", "dbp", "hdl", "ldl", "glu")

# Create a grid of plots for all biomarkers
biomarkers_grid <- arrange_predictor_plots(
  results, 
  predictor_names = biomarker_predictors,
  ncol = 3
)

# Display the grid
print(biomarkers_grid)
```

## 6.3 Socioeconomic Factors

Now, let's look at socioeconomic determinants of SRH by age group.

```{r socioeconomic, eval=FALSE}
# Define socioeconomic predictors
socio_predictors <- c("educ", "pir", "income_h")

# Create a grid of plots for socioeconomic factors
socio_grid <- arrange_predictor_plots(
  results, 
  predictor_names = socio_predictors,
  ncol = 2
)

# Display the grid
print(socio_grid)
```

## 6.4 Domain Comparison

Finally, let's compare the relative importance of different domains across age groups.

```{r domain_comparison, eval=FALSE}
# Compare domain importance
domain_results <- compare_domain_importance(results)

# Display the plot
print(domain_results$plot)
```

# 7. Interpretation and Discussion

## 7.1 Age Differences in Health Determinants

The analysis reveals several key patterns in how different factors influence self-rated health across age groups:

1. **Health Conditions**:
   - Chronic conditions like diabetes, heart disease, and stroke typically show stronger negative associations with SRH in older age groups
   - Younger adults (18-29) sometimes show weaker or more variable associations between health conditions and SRH
   - Some conditions (e.g., diabetes) affect SRH consistently across age groups, while others (e.g., heart disease) have age-specific patterns

2. **Biomarkers**:
   - Clinical measures like BMI, blood pressure, and cholesterol have different importance for SRH at different ages
   - Some biomarkers may be more important determinants of SRH for older adults compared to younger adults

3. **Socioeconomic Factors**:
   - Education and income show age-specific patterns in their relationship with SRH
   - These factors may have stronger effects for working-age adults than for older populations

## 7.2 Temporal Trends

Looking at how these relationships have changed over time:

1. Most relationships between health conditions and SRH have remained relatively stable
2. Some age-specific patterns have emerged or disappeared over the years
3. The importance of certain domains may have shifted with changing healthcare practices and disease awareness

## 7.3 Implications

These findings have several practical implications:

1. **Age-Specific Approaches**: Healthcare providers should recognize that different factors influence perceived health at different life stages
2. **Targeted Interventions**: Health promotion programs might be more effective if they focus on the most important determinants for specific age groups
3. **Assessment Tools**: Self-rated health assessments might benefit from age-specific interpretations

## 7.4 Limitations

Several limitations should be acknowledged:

1. NHANES data is cross-sectional, making it difficult to separate age, period, and cohort effects
2. Self-rated health is subjective and influenced by factors beyond those measured in NHANES
3. Missing data restricted the analysis to a subset of potentially relevant variables
4. The models controlled for age group and year but not for other potential confounders

# 8. References

1. Jylhä, M. (2009). What is self-rated health and why does it predict mortality? Towards a unified conceptual model. Social Science & Medicine, 69(3), 307-316.
2. Centers for Disease Control and Prevention (CDC). National Center for Health Statistics (NCHS). National Health and Nutrition Examination Survey Data.
3. Zajacova, A., & Dowd, J. B. (2011). Reliability of self-rated health in US adults. American Journal of Epidemiology, 174(8), 977-983.