---
title: "Median Polish APC Decomposition of Self-Rated Health"
subtitle: "Non-parametric Age-Period-Cohort analysis across 6 US surveys"
author: "Christine Lucille Kuryla"
date: today
format:
  gfm:
    toc: true
    toc-depth: 3
    number-sections: true
    fig-width: 12
    fig-height: 7
    fig-dpi: 300
  html:
    toc: true
    toc-depth: 3
    toc-location: left
    number-sections: true
    code-fold: true
    code-summary: "Show code"
    theme: flatly
    fig-width: 12
    fig-height: 7
    fig-dpi: 300
    embed-resources: true
    smooth-scroll: true
execute:
  echo: true
  warning: false
  message: false
---

```{r setup}
#| label: setup
#| code-fold: true

library(tidyverse)
library(here)
library(survey)
library(srvyr)
library(patchwork)
library(broom)
library(knitr)
library(kableExtra)

is_html <- knitr::is_html_output()

# Helper: apply kable_styling only for HTML output
styled_kable <- function(k, ...) {
  if (is_html) kable_styling(k, ...) else k
}

source(here::here("R", "paths.R"))
source(here::here("R", "functions", "theme_srh.R"))
source(here::here("R", "functions", "plot_utils.R"))
source(here::here("R", "functions", "medpolish_apc.R"))

mp_output_dir <- here::here("output", "apc", "medpolish")

# Survey order and labels
surveys <- c("gss", "nhanes", "meps", "nhis", "cps", "brfss")
survey_labels <- c(GSS = "GSS", NHANES = "NHANES", MEPS = "MEPS",
                   NHIS = "NHIS", CPS = "CPS", BRFSS = "BRFSS")
```

# Overview

This report presents a **median polish** decomposition of self-rated health (SRH) across six US surveys. The median polish is a non-parametric method that complements the Bayesian HAPC analysis by extracting additive Age, Period, and Cohort effects without distributional assumptions.

## Method

1. Compute **survey-weighted mean SRH** in 5-year age group $\times$ 5-year period cells
2. Construct three 2D matrices: **AP** (age $\times$ period), **AC** (age $\times$ cohort), **PC** (period $\times$ cohort)
3. Apply `medpolish()` to each matrix to extract row effects, column effects, and residuals
4. The residuals from each slice reveal the effect of the *third* dimension (e.g., AP residuals $\to$ cohort effects)

## Surveys

| Survey | SRH Scale | Years | Design |
|--------|-----------|-------|--------|
| GSS | 1--4 | 1972--2024 | Weights only |
| NHANES | 1--5 | 1999--2022 | Full (strata + PSU + weights) |
| MEPS | 1--5 | 2000--2023 | Full (strata + PSU + weights) |
| NHIS | 1--5 | 1983--2024 | Full (strata + PSU + weights) |
| CPS | 1--5 | 1996--2025 | Weights only |
| BRFSS | 1--5 | 1993--2025 | Weights only (computational) |

: SRH is coded higher = better in all surveys. {.striped}

```{r load-results}
#| label: load-results

# Load all saved results from the pipeline
all_results <- list()
all_effects <- tibble()
all_direct  <- tibble()
all_lm      <- tibble()
all_var     <- tibble()

for (sv in surveys) {
  r <- readRDS(file.path(mp_output_dir, paste0("medpolish_full_", sv, ".rds")))
  all_results[[sv]] <- r

  all_effects <- bind_rows(all_effects,
                           r$effects |> mutate(survey = toupper(sv)))
  all_direct  <- bind_rows(all_direct,
                           r$direct_effects |> mutate(survey = toupper(sv)))
  all_lm      <- bind_rows(all_lm,
                           r$lm_summary |> mutate(survey = toupper(sv)))
  all_var     <- bind_rows(all_var,
                           r$var_decomp |> mutate(survey = toupper(sv)))
}

sv_levels <- toupper(surveys)
all_effects <- all_effects |> mutate(survey = factor(survey, levels = sv_levels))
all_direct  <- all_direct  |> mutate(survey = factor(survey, levels = sv_levels))
```

# Combined Results {.tabset}

## Mean-Residual APC Effects

The combined figure shows **mean residuals** from the two-way median polish, averaged across cells for each APC dimension. Points represent the estimated effect; lines show linear trend fits.

```{r fig-combined}
#| label: fig-combined
#| fig-cap: "Median polish APC effects across all 6 surveys. Columns: Age (blue), Period (green), Cohort (pink). Each point is the mean residual for that dimension value."
#| fig-height: 14
#| fig-width: 12

plot_medpolish_combined_simple(all_effects, survey_order = sv_levels)
```

## Direct Medpolish Effects

The direct effects are the **row and column effects** extracted directly by `medpolish()`. Each dimension appears in two slices (e.g., Age is a row effect in both the AP and AC matrices), providing two independent estimates. Agreement between slices indicates a robust effect; divergence indicates sensitivity to the conditioning dimension.

```{r fig-direct}
#| label: fig-direct
#| fig-cap: "Direct row/column effects from medpolish. Each dimension has two estimates from different matrix slices (solid vs dashed lines). Close agreement = robust effect."
#| fig-height: 16
#| fig-width: 13

plot_direct_effects_all_surveys(all_direct, survey_order = sv_levels)
```

## Residual Diagnostics

Cell-level residuals from each matrix slice, plotted against the third dimension. Systematic patterns in residuals indicate effects not captured by the additive row + column decomposition.

```{r fig-residuals}
#| label: fig-residuals
#| fig-cap: "Median polish residuals across surveys. Gray dots = individual cell residuals; colored line = mean residual trend. Structured patterns indicate real effects of the third dimension."
#| fig-height: 16
#| fig-width: 13

plot_residuals_all_surveys(all_results, survey_order = surveys)
```

# Variance Decomposition

For each matrix slice, the table shows the percentage of total cell-mean variation (sum of squares) explained by row effects, column effects, and residuals.

::: {.callout-note}
Because median polish uses medians (not means), the decomposition is not orthogonal --- component percentages may not sum to exactly 100%.
:::

## AP Matrix: Age vs Period {.unnumbered}

```{r tbl-ap}
#| label: tbl-ap

all_var |>
  filter(slice == "AP", component != "Total") |>
  select(Survey = survey, Component = component, `SS` = ss, `% of Total` = pct) |>
  mutate(`SS` = round(`SS`, 3),
         `% of Total` = round(`% of Total`, 1)) |>
  pivot_wider(names_from = Component, values_from = c(`SS`, `% of Total`)) |>
  select(Survey,
         `Age SS` = `SS_Age`, `Age %` = `% of Total_Age`,
         `Period SS` = `SS_Period`, `Period %` = `% of Total_Period`,
         `Residual SS` = `SS_Residual`, `Residual %` = `% of Total_Residual`) |>
  kable(digits = 1, caption = "AP matrix: Age dominates Period in every survey.") |>
  styled_kable(bootstrap_options = c("striped", "hover", "condensed"),
               full_width = FALSE)
```

## AC Matrix: Age vs Cohort {.unnumbered}

```{r tbl-ac}
#| label: tbl-ac

all_var |>
  filter(slice == "AC", component != "Total") |>
  select(Survey = survey, Component = component, `SS` = ss, `% of Total` = pct) |>
  mutate(`SS` = round(`SS`, 3),
         `% of Total` = round(`% of Total`, 1)) |>
  pivot_wider(names_from = Component, values_from = c(`SS`, `% of Total`)) |>
  select(Survey,
         `Age SS` = `SS_Age`, `Age %` = `% of Total_Age`,
         `Cohort SS` = `SS_Cohort`, `Cohort %` = `% of Total_Cohort`,
         `Residual SS` = `SS_Residual`, `Residual %` = `% of Total_Residual`) |>
  kable(digits = 1, caption = "AC matrix: Age dominates Cohort, though Cohort contributes modestly (4-21%).") |>
  styled_kable(bootstrap_options = c("striped", "hover", "condensed"),
               full_width = FALSE)
```

## PC Matrix: Period vs Cohort {.unnumbered}

```{r tbl-pc}
#| label: tbl-pc

all_var |>
  filter(slice == "PC", component != "Total") |>
  select(Survey = survey, Component = component, `SS` = ss, `% of Total` = pct) |>
  mutate(`SS` = round(`SS`, 3),
         `% of Total` = round(`% of Total`, 1)) |>
  pivot_wider(names_from = Component, values_from = c(`SS`, `% of Total`)) |>
  select(Survey,
         `Period SS` = `SS_Period`, `Period %` = `% of Total_Period`,
         `Cohort SS` = `SS_Cohort`, `Cohort %` = `% of Total_Cohort`,
         `Residual SS` = `SS_Residual`, `Residual %` = `% of Total_Residual`) |>
  kable(digits = 1, caption = "PC matrix: Cohort dominates Period once age is removed.") |>
  styled_kable(bootstrap_options = c("striped", "hover", "condensed"),
               full_width = FALSE)
```

# Linear Trend Tests

Linear models fit to the mean-residual effects test whether each APC dimension has a significant linear trend. Slopes represent the change in residual effect (SRH units) per unit increase in the dimension.

```{r tbl-lm}
#| label: tbl-lm

all_lm |>
  filter(term == "value") |>
  mutate(
    sig = case_when(
      p.value < 0.001 ~ "***",
      p.value < 0.01  ~ "**",
      p.value < 0.05  ~ "*",
      TRUE ~ ""
    ),
    Slope = sprintf("%.4f", estimate),
    `95% CI` = sprintf("[%.4f, %.4f]", conf.low, conf.high),
    p = sprintf("%.4f", p.value)
  ) |>
  select(Survey = survey, Dimension = dimension, Slope, `95% CI`, p, sig) |>
  kable(caption = "Linear trend slopes for median polish effects. * p < 0.05, ** p < 0.01, *** p < 0.001.") |>
  styled_kable(bootstrap_options = c("striped", "hover", "condensed"),
               full_width = FALSE)
```

# Per-Survey Detail {.tabset}

```{r}
#| results: asis

for (sv in surveys) {
  cat("\n\n## ", toupper(sv), " {.unnumbered}\n\n")

  r <- all_results[[sv]]

  cat("**Cell means:**", nrow(r$cell_means), "cells |",
      "**SRH range:**", round(min(r$cell_means$mean_srh), 2), "--",
      round(max(r$cell_means$mean_srh), 2), "\n\n")

  # Variance decomposition mini-table
  vd <- r$var_decomp |>
    filter(component != "Total") |>
    mutate(pct = round(pct, 1)) |>
    select(Slice = slice, Component = component, `% of Total` = pct)

  print(kable(vd, caption = paste0(toupper(sv), " variance decomposition")) |>
          styled_kable(bootstrap_options = c("striped", "condensed"),
                       full_width = FALSE))

  cat("\n\n")
}
```

### Per-survey effects plots

::: {.panel-tabset}

```{r}
#| results: asis
#| fig-height: 5
#| fig-width: 14

for (sv in surveys) {
  cat("\n\n#### ", toupper(sv), "\n\n")
  r <- all_results[[sv]]
  p <- plot_medpolish_effects(r$effects, sv)
  print(p)
  cat("\n\n")
}
```

:::

### Per-survey direct effects plots

::: {.panel-tabset}

```{r}
#| results: asis
#| fig-height: 6
#| fig-width: 14

for (sv in surveys) {
  cat("\n\n#### ", toupper(sv), "\n\n")
  r <- all_results[[sv]]
  p <- plot_direct_effects(r$direct_effects, sv)
  print(p)
  cat("\n\n")
}
```

:::

# Summary

```{r tbl-summary}
#| label: tbl-summary

# Build a concise summary table
summary_tbl <- all_var |>
  filter(component != "Total") |>
  select(survey, slice, component, pct) |>
  mutate(pct = round(pct, 1)) |>
  unite("slice_comp", slice, component, sep = ": ") |>
  pivot_wider(names_from = slice_comp, values_from = pct)

k <- kable(summary_tbl,
           caption = "Variance decomposition summary (% of total SS) across all slices and surveys.")
if (is_html) {
  k |>
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  full_width = TRUE, font_size = 12) |>
    add_header_above(c(" " = 1, "AP (Age vs Period)" = 3,
                        "AC (Age vs Cohort)" = 3,
                        "PC (Period vs Cohort)" = 3))
} else {
  k
}
```

## Key Findings

1. **Age dominates** SRH variation in every survey and every matrix slice. In the AP matrix, Age explains 74--104% of total SS versus 0.3--14% for Period.

2. **Cohort effects are secondary but real.** In the AC matrix, Cohort explains 4--21% of variation. In the PC matrix (after removing age), Cohort dominates Period (78--134% vs 3--75%).

3. **Period effects are minimal.** Period is consistently the weakest dimension, explaining <5% in the AP matrix for most surveys (except BRFSS at 14%).

4. **Linear trends** in the residual effects are mostly non-significant. Only NHANES and BRFSS show significant positive cohort slopes, and MEPS shows a significant positive period slope.

5. **Direct effects agree across slices for Age** (robust), but **diverge for Period and Cohort** (reflecting the fundamental APC identification problem).

These results are consistent with the parametric BHAPC analysis: SRH is primarily structured by age, with modest birth-cohort variation and negligible calendar-time trends.

---

*Report generated `r Sys.time()`*
