---
title: "NHANES IV SRH and Health Covariates"
author: "Christine Lucille Kuryla"
date: "2025-01-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(survey)
library(srvyr)
library(broom)
library(Hmisc)
library(tidyverse)
library(rlang)

```

NHANES IV has a wealth of data. We are going to explore the relationship between many health variables and SRH, and see if that changes with age over time.

### Load and wrangle data

```{r}

# Load and wrangle nhanes data 

data_nh <- read_csv(here("code_examples/kamaryn_nhanes/nhanes_1999-2018_2023-11-29.csv"))

dim(data_nh)
sum(is.na(data_nh$health)) # note there are 38579 NA's for health out of 96241 subj

year_to_wave <- read_csv(here("big_data/NHANES/nhanes_4/nh4_year_to_wave.csv")) %>% 
  mutate(release_nb = wave) 

data_nhanes <- data_nh %>% 
  filter(visit == 1) %>% 
  filter(age_visit >= 18) %>% 
  filter(!(is.na(health))) %>% # remove subjects without our primary variable of interest
 # left_join(year_to_wave, by = "release_nb") %>% # year of survey (***note each survey is two years, this is aproximate and just the first year)
  mutate(srh = 6 - health) %>%  # recode SRH to be more intuitive (Excellent = 5 to Poor = 1)
  mutate(age = age_visit) # more intuitive naming for APC analyses

# variables of interest: 
# self-rated health: "health" originally, reverse recoded to "srh"
# age at survey: "age_visit"
# age at follow-up (censored/deceased): "age_last"
# vital status at follow-up (alive/decesased): "deceased"
# year/wave: "release_nb"
# SEQN: "BaseID"
# "ddem_wghts"	Full Sample 2 Year MEC Exam Weight	WTMEC2YR
# "dem_wghts_4yr"	Full Sample 4 Year MEC Exam Weight	WTMEC4YR


# Adding gloria's for weight variable: SDDSRVYR and other clock and APC things if needed 

data_nhanes_gloria <- read.csv(here("big_data/NHANES/Gloria_preprocessed/Data-3/Core_Dataset_Aim2.csv"))

data_nhanes <- data_nhanes %>% 
  mutate(SEQN = as.character(BaseID)) %>% 
  left_join(data_nhanes_gloria %>% 
              rename(race_decoded = race) %>% 
              mutate(age = as.numeric(age)), 
            by = c("SEQN", "age")) %>% 
  mutate(age = as.numeric(age))

# add more

data_nhanes <- data_nhanes %>% 
  mutate(age_group = as.factor( 
                            cut(
                            age,
                            breaks = c(17, 29, 39, 49, 59, 69, Inf),
                            labels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+"),
                            levels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+"),
                            right = TRUE
                          )),
         health_cat = factor(srh, 
                        levels = 1:5,
                        labels = c("Poor", "Fair", "Good", "Very Good", "Excellent")))

# more new columns to enable APC
data_nhanes <- data_nhanes %>% 
  mutate(period_cut_6 = as.factor(cut(data_nhanes$year, 6)),
    period_cut_10 = as.factor(cut(data_nhanes$year, 10)),
    period_cut_12 = as.factor(cut(data_nhanes$year, 12)),
    period_groups = as.factor(cut(data_nhanes$year, 12)),
    period_10yr = as.factor(
                            cut(
                            year,
                            breaks = c(1973, 1982, 1990, 1998, 2006, 2014, Inf),
                            labels = c("1974-1982", "1982-1990", "1990-1998", 
                                       "1998-2006", "2006-2014", "2014-2022"),
                            right = TRUE
                            )
                          ),
    period_decade = as.factor(
                            cut(
                            year,
                            breaks = c(1973, 1979, 1989, 1999, 2009, 2019, Inf),
                            labels = c("1974-1979", "1980-1989", "1990-1999",
                                       "2000-2009", "2010-2019", "2020-2024"),
                            right = TRUE
                            )
                          ),
    age_group = as.factor( 
                            cut(
                            age,
                            breaks = c(17, 29, 39, 49, 59, 69, Inf),
                            labels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+"),
                            right = TRUE
                          )),
    age_groups = as.factor( 
                            cut(
                            age,
                            breaks = c(17, 29, 39, 49, 59, 69, Inf),
                            labels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+"),
                            right = TRUE
                          )),
    age_group_small = as.factor( 
                            cut(
                              age,
                              breaks = c(seq(15, 75, by = 5), Inf),  # Define breaks up to 75 and include Inf for the last group
                              labels = c("16-20", "21-25", "26-30", "31-35", "36-40", "41-45", "46-50", "51-55", "56-60", "61-65", "66-70", "71-75", "76+"),
                              right = FALSE  # Makes intervals left-closed, i.e., [x, y)
                            )
                          )
                          ,
    generation = factor(
      case_when(
        cohort >= 1901 & cohort <= 1927 ~ "Greatest (1901-1927)",
        cohort >= 1928 & cohort <= 1945 ~ "Silent (1928-1945)",
        cohort >= 1946 & cohort <= 1964 ~ "Boomers (1946-1964)",
        cohort >= 1965 & cohort <= 1980 ~ "Gen X (1965-1980)",
        cohort >= 1981 & cohort <= 1996 ~ "Millennials (1981-1996)",
        cohort >= 1997 & cohort <= 2012 ~ "Gen Z (1997-2012)",
        TRUE ~ "Other"
      ),
      levels = c(
        "Greatest (1901-1927)",
        "Silent (1928-1945)",
        "Boomers (1946-1964)",
        "Gen X (1965-1980)",
        "Millennials (1981-1996)",
        "Gen Z (1997-2012)"#,
     #   "Other"
      )
    ),
    generation_two_sections = factor(
      case_when(
        generation == "Greatest (1901-1927)" & cohort <= 1914 ~ "Greatest Early (1901-1914)",
        generation == "Greatest (1901-1927)" & cohort > 1914 ~ "Greatest Late (1915-1927)",
        generation == "Silent (1928-1945)" & cohort <= 1936 ~ "Silent Early (1928-1936)",
        generation == "Silent (1928-1945)" & cohort > 1936 ~ "Silent Late (1937-1945)",
        generation == "Boomers (1946-1964)" & cohort <= 1955 ~ "Boomers Early (1946-1955)",
        generation == "Boomers (1946-1964)" & cohort > 1955 ~ "Boomers Late (1956-1964)",
        generation == "Gen X (1965-1980)" & cohort <= 1972 ~ "Gen X Early (1965-1972)",
        generation == "Gen X (1965-1980)" & cohort > 1972 ~ "Gen X Late (1973-1980)",
        generation == "Millennials (1981-1996)" & cohort <= 1988 ~ "Millennials Early (1981-1988)",
        generation == "Millennials (1981-1996)" & cohort > 1988 ~ "Millennials Late (1989-1996)",
        generation == "Gen Z (1997-2012)" & cohort <= 2004 ~ "Gen Z Early (1997-2004)",
        generation == "Gen Z (1997-2012)" & cohort > 2004 ~ "Gen Z Late (2005-2012)",
        TRUE ~ "Other"
      ),
      levels = c(
        "Greatest Early (1901-1914)", "Greatest Late (1915-1927)",
        "Silent Early (1928-1936)", "Silent Late (1937-1945)",
        "Boomers Early (1946-1955)", "Boomers Late (1956-1964)",
        "Gen X Early (1965-1972)", "Gen X Late (1973-1980)",
        "Millennials Early (1981-1988)", "Millennials Late (1989-1996)",
        "Gen Z Early (1997-2004)", "Gen Z Late (2005-2012)"#,
     #   "Other"
      )
    ),
    generation_three_sections = factor(
      case_when(
        generation == "Greatest (1901-1927)" & cohort <= 1910 ~ "Greatest Early (1901-1910)",
        generation == "Greatest (1901-1927)" & cohort > 1910 & cohort <= 1918 ~ "Greatest Mid (1911-1918)",
        generation == "Greatest (1901-1927)" & cohort > 1918 ~ "Greatest Late (1919-1927)",
        generation == "Silent (1928-1945)" & cohort <= 1934 ~ "Silent Early (1928-1934)",
        generation == "Silent (1928-1945)" & cohort > 1934 & cohort <= 1940 ~ "Silent Mid (1935-1940)",
        generation == "Silent (1928-1945)" & cohort > 1940 ~ "Silent Late (1941-1945)",
        generation == "Boomers (1946-1964)" & cohort <= 1951 ~ "Boomers Early (1946-1951)",
        generation == "Boomers (1946-1964)" & cohort > 1951 & cohort <= 1958 ~ "Boomers Mid (1952-1958)",
        generation == "Boomers (1946-1964)" & cohort > 1958 ~ "Boomers Late (1959-1964)",
        generation == "Gen X (1965-1980)" & cohort <= 1970 ~ "Gen X Early (1965-1970)",
        generation == "Gen X (1965-1980)" & cohort > 1970 & cohort <= 1976 ~ "Gen X Mid (1971-1976)",
        generation == "Gen X (1965-1980)" & cohort > 1976 ~ "Gen X Late (1977-1980)",
        generation == "Millennials (1981-1996)" & cohort <= 1986 ~ "Millennials Early (1981-1986)",
        generation == "Millennials (1981-1996)" & cohort > 1986 & cohort <= 1992 ~ "Millennials Mid (1987-1992)",
        generation == "Millennials (1981-1996)" & cohort > 1992 ~ "Millennials Late / Gen Z (1993-2004)",
    #    generation == "Gen Z (1997-2012)" & cohort <= 2002 ~ "Gen Z Early (1997-2002)",
    #    generation == "Gen Z (1997-2012)" & cohort > 2002 & cohort <= 2008 ~ "Gen Z Mid (2003-2008)",
    #    generation == "Gen Z (1997-2012)" & cohort > 2008 ~ "Gen Z Late (2009-2012)",
        TRUE ~ "Other"
      ),
      levels = c(
        "Greatest Early (1901-1910)", "Greatest Mid (1911-1918)", "Greatest Late (1919-1927)",
        "Silent Early (1928-1934)", "Silent Mid (1935-1940)", "Silent Late (1941-1945)",
        "Boomers Early (1946-1951)", "Boomers Mid (1952-1958)", "Boomers Late (1959-1964)",
        "Gen X Early (1965-1970)", "Gen X Mid (1971-1976)", "Gen X Late (1977-1980)",
        "Millennials Early (1981-1986)", "Millennials Mid (1987-1992)", 
        "Millennials Late / Gen Z (1993-2004)"
        #"Millennials Late (1993-1996)",
      #  "Gen Z Early (1997-2002)", "Gen Z Mid (2003-2008)", "Gen Z Late (2009-2012)" #,
      #  "Other"
      )
    ))
  

```

### Available variables

```{r}

colnames(data_nhanes)

```


## Create survey design object

```{r}

svy_nhanes <- data_nhanes %>%
  filter(!(is.na(SDMVPSU))) %>% 
  as_survey_design(
  #  ids = 1,
    ids = SDMVPSU,           # PSU identifiers (use 1 if not available)
  #  weights = WTINT2YR,  # original -- interview weights -- larger sample size but fewer covariates
    weights = WTMEC2YR, # Gloria's -- enable more covariates
    strata = SDMVSTRA,
    nest = TRUE
    )

```

# Relationship between SRH and time, age, and cohort (overall)

## Relationship between SRH and X over time (per wave)

```{r}

# We will need model formula, design (svy_nhanes), and family 
# family=stats::gaussian(), family=quasibinomial(), family=quasipoisson()

# This will extract the coefficient for the first term in the model summary

# coefficient <- model_summary$coefficients[2,"Estimate"]
# se <- model_summary$coefficients[2,"Std. Error"]
# t_value <- model_summary$coefficients[2,"t value"] 
# p_value <- model_summary$coefficients[2,"Pr(>|t|)"]

svyglm(formula = srh ~ heart_dis,
       design = svy_nhanes,
       family = stats::gaussian()
         )

######
# This function gives us a tibble of the glm
#####

srh_vs_X_results <- function(formula,
                                      design = svy_nhanes,
                                      family = stats::gaussian(),
                                      group_by_1) {
  
  # Convert string formula to R formula
  f <- as.formula(formula)
  
  # Extract variable names
  vars_in_formula <- all.vars(f)
  var_1 <- vars_in_formula[1]  # y-variable
  var_2 <- vars_in_formula[2]  # x-variable
  
  # Convert them to symbols for filter
  var_1_sym <- rlang::sym(var_1)
  var_2_sym <- rlang::sym(var_2)
  
  # Grouping, modeling, returning a data frame of results
  model_results <- design %>%
    # Filter out NAs for just the variables in the model
    filter(!is.na(!!var_1_sym)) %>%
    filter(!is.na(!!var_2_sym)) %>%
    group_by({{group_by_1}}) %>%  #, {{group_by_2}}) %>%
    group_map_dfr(~ {
      
      # Check if there's enough variation in the *response* variable
      if (length(unique(.x$variables[[var_1]])) < 2) {
        # Not enough variation in var_1 => return row of NAs
        return(tibble(
          group_1     = unique(.x$variables[[rlang::as_name(enquo(group_by_1))]]),
     #     group_2     = unique(.x$variables[[rlang::as_name(enquo(group_by_2))]]),
          formula     = as.character(formula),
          coefficient = NA_real_,
          se          = NA_real_,
          t_value     = NA_real_,
          p_value     = NA_real_
        ))
      }
      
      # Fit the model
      model_var <- survey::svyglm(formula = f, design = ., family = family)
      
      # Extract the summary coefficients
      coefs <- summary(model_var)$coefficients
      
      # We want the row corresponding to var_2
      # If that row doesn't exist, return NA for that group
      if (!var_2 %in% rownames(coefs)) {
        return(tibble(
          group_1     = unique(.x$variables[[rlang::as_name(enquo(group_by_1))]]),
     #     group_2     = unique(.x$variables[[rlang::as_name(enquo(group_by_2))]]),
          formula     = as.character(formula),
          coefficient = NA_real_,
          se          = NA_real_,
          t_value     = NA_real_,
          p_value     = NA_real_
        ))
      }
      
      tibble(
        group_1     = unique(.x$variables[[rlang::as_name(enquo(group_by_1))]]),
    #    group_2     = unique(.x$variables[[rlang::as_name(enquo(group_by_2))]]),
        formula     = as.character(formula),
        coefficient = coefs[var_2, "Estimate"],
        se          = coefs[var_2, "Std. Error"],
        t_value     = coefs[var_2, "t value"],
        p_value     = coefs[var_2, "Pr(>|t|)"]
      )
    }) %>%
    ungroup() %>% 
    filter(formula == var_2)
  
  model_results
}

####
#This function plots the relationship
####

srh_vs_X_plot <- function(model_results, title = NULL){
  group_by_1 <- colnames(model_results)[1]   # Grab the name of the first column as a string
  group_by_1_sym <- sym(group_by_1)   # Convert it to a symbol for tidy evaluation
  title <- model_results$formula[1]
  title <- title
  
  ggplot(model_results, aes(x = !!group_by_1_sym, y = coefficient)) +
    geom_line() +
    geom_point() +
    geom_errorbar(aes(ymin = coefficient - se, ymax = coefficient + se), 
                  width = 0.2) +
    labs(title = title)
}

#Test

srh_vs_X_results_data <- srh_vs_X_results(formula = srh ~ stroke, 
        # design = svy_nhanes, 
       #  family = stats::gaussian(), 
         group_by_1 = year)

srh_vs_X_plot(srh_vs_X_results_data)

### Apply function

## Relationship between SRH and X by year

X <- "heart_dis"
srh_vs_X_results_data <- srh_vs_X_results(formula = formula(paste0("srh ~ ", X)), group_by_1 = year)
srh_vs_X_results_data
srh_vs_X_plot(srh_vs_X_results_data)

####
# Use this function if you just want the plots:
####

plot_var_vs_srh_per_something <- function(X, group_by_1){
      group_by_1_sym <- sym(group_by_1)
      model_results <- srh_vs_X_results(formula = formula(paste0("srh ~ ", X)),
                               group_by_1 = !!group_by_1_sym)
      srh_vs_X_plot(model_results)
}

# test
plot_var_vs_srh_per_something("stroke", "age_group")
```

### Apply functions

Now we have a 

```{r}

# Functions of interest
svy_nhanes %>% 
  group_by(year) %>% 
  svyglm(formula = srh ~ phq9,
         design = .,
         family = stats::gaussian()
           ) %>% 
  summary() 

## Relationship between SRH and X over time

plot_var_vs_srh_per_something("heart_dis" , "year")

## Relationship between SRH and X for age groups

## Relationship between SRH and X for cohorts


```

# Does the relationship between SRH and X for different age groups change over time? For cohorts? 

## Define functions


We have various outcome variables, of different types (continuous, dichotomous, categorical, and count). In the future we will want to possibly adjust for other covariates. Therefore, we will write a function with the model as an input, but for now, also allow for hust the variable of interest. This can be modified as needed.

```{r}

# We will need model formula, design (svy_nhanes), and family 
# family=stats::gaussian(), family=quasibinomial(), family=quasipoisson()

# This will extract the coefficient for the first term in the model summary

# coefficient <- model_summary$coefficients[2,"Estimate"]
# se <- model_summary$coefficients[2,"Std. Error"]
# t_value <- model_summary$coefficients[2,"t value"] 
# p_value <- model_summary$coefficients[2,"Pr(>|t|)"]

svyglm(formula = srh ~ heart_dis,
       design = svy_nhanes,
       family = stats::gaussian()
         )

######
# This function gives us a tibble of the glm
#####
srh_vs_X_per_Y_results <- function(formula,
                                      design = svy_nhanes,
                                      family = stats::gaussian(),
                                      group_by_1,
                                      group_by_2) {
  
  # Convert string formula to R formula
  f <- as.formula(formula)
  
  # Extract variable names
  vars_in_formula <- all.vars(f)
  var_1 <- vars_in_formula[1]  # y-variable
  var_2 <- vars_in_formula[2]  # x-variable
  
  # Convert them to symbols for filter
  var_1_sym <- rlang::sym(var_1)
  var_2_sym <- rlang::sym(var_2)
  
  # Grouping, modeling, returning a data frame of results
  model_results <- design %>%
    # Filter out NAs for just the variables in the model
    filter(!is.na(!!var_1_sym)) %>%
    filter(!is.na(!!var_2_sym)) %>%
    group_by({{group_by_1}}, {{group_by_2}}) %>%
    group_map_dfr(~ {
      
      # Check if there's enough variation in the *response* variable
      if (length(unique(.x$variables[[var_1]])) < 2) {
        # Not enough variation in var_1 => return row of NAs
        return(tibble(
          group_1     = unique(.x$variables[[rlang::as_name(enquo(group_by_1))]]),
          group_2     = unique(.x$variables[[rlang::as_name(enquo(group_by_2))]]),
          formula     = as.character(formula),
          coefficient = NA_real_,
          se          = NA_real_,
          t_value     = NA_real_,
          p_value     = NA_real_
        ))
      }
      
      # Fit the model
      model_var <- survey::svyglm(formula = f, design = ., family = family)
      
      # Extract the summary coefficients
      coefs <- summary(model_var)$coefficients
      
      # We want the row corresponding to var_2
      # If that row doesn't exist, return NA for that group
      if (!var_2 %in% rownames(coefs)) {
        return(tibble(
          group_1     = unique(.x$variables[[rlang::as_name(enquo(group_by_1))]]),
          group_2     = unique(.x$variables[[rlang::as_name(enquo(group_by_2))]]),
          formula     = as.character(formula),
          coefficient = NA_real_,
          se          = NA_real_,
          t_value     = NA_real_,
          p_value     = NA_real_
        ))
      }
      
      tibble(
        group_1     = unique(.x$variables[[rlang::as_name(enquo(group_by_1))]]),
        group_2     = unique(.x$variables[[rlang::as_name(enquo(group_by_2))]]),
        formula     = as.character(formula),
        coefficient = coefs[var_2, "Estimate"],
        se          = coefs[var_2, "Std. Error"],
        t_value     = coefs[var_2, "t value"],
        p_value     = coefs[var_2, "Pr(>|t|)"]
      )
    }) %>%
    ungroup() %>% 
    filter(formula == var_2)
  
  model_results
}


####
#This function plots the relationship
####

srh_vs_X_per_Y_plot <- function(model_results, title = NULL){
  group_by_1 <- colnames(model_results)[1]   # Grab the name of the first column as a string
  group_by_1_sym <- sym(group_by_1)   # Convert it to a symbol for tidy evaluation
  title <- model_results$formula[1]
  title <- title
    group_by_2 <- colnames(model_results)[2]   # Grab the name of the first column as a string
  group_by_2_sym <- sym(group_by_2) 
  
  ggplot(model_results, aes(x = !!group_by_1_sym, y = coefficient, color = !!group_by_2_sym)) +
    geom_line() +
    geom_point() +
    geom_errorbar(aes(ymin = coefficient - se, ymax = coefficient + se), 
                  width = 0.2) +
    labs(title = title)
}

#Test

srh_vs_X_per_Y_results_data <- srh_vs_X_per_Y_results(formula = srh ~ heart_dis, 
        # design = svy_nhanes, 
       #  family = stats::gaussian(), 
         group_by_1 = year,
         group_by_2 = age_group)

srh_vs_X_per_Y_plot(srh_vs_X_per_Y_results_data)

### Apply function

## Relationship between SRH and X by year

X <- "heart_dis"
srh_vs_X_per_Y_results_data <- srh_vs_X_per_Y_results(formula = formula(paste0("srh ~ ", X)), group_by_1 = year)
srh_vs_X_per_Y_results_data
srh_vs_X_per_Y_plot(srh_vs_X_per_Y_results_data)

####
# Use this function if you just want the plots:
####

plot_srh_vs_X_per_Y_per_something <- function(X, group_by_1){
      group_by_1_sym <- sym(group_by_1)
      model_results <- srh_vs_X_per_Y_results(formula = formula(paste0("srh ~ ", X)),
                               group_by_1 = !!group_by_1_sym)
      srh_vs_X_per_Y_plot(model_results)
}

# test
plot_srh_vs_X_per_Y_per_something("stroke", "age_group")
```

### Apply functions

Now we have a 

```{r}

# Functions of interest

## Relationship between SRH and X over time

## Relationship between SRH and X for age groups

## Relationship between SRH and X for cohorts


```


## Relationship between SRH and X by age group per wave


