---
title: "Regressions"
author: "Christine Lucille Kuryla"
date: "2024-12-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(Hmisc)
library(dplyr)
library(purrr)
library(broom)
library(tidyr)
library(here)
library(tidyverse)

```

```{r}

data_gss <- read_csv(here("data/cleaned/gss_groups.csv")) %>% 
    mutate(
    health_cat = factor(health, 
                        levels = 1:4,
                        labels = c("Poor", "Fair", "Good", "Excellent")),
    # period_cut_6 = as.factor(cut(data_gss$year, 6)),
    # period_cut_10 = as.factor(cut(data_gss$year, 10)),
    # period_cut_12 = as.factor(cut(data_gss$year, 12)),
    # period_groups = as.factor(cut(data_gss$year, 12)),
    period_7yr = as.factor(
                            cut(
                            year,
                            breaks = c(1973, 1982, 1990, 1998, 2006, 2014, Inf),
                            labels = c("1974-1982", "1983-1990", "1991-1998", 
                                       "1999-2006", "2007-2014", "2015-2022"),
                            right = TRUE
                            )
                          ),
    # period_5yr = as.factor(
    #                         cut(
    #                         year,
    #                         breaks = c(1973, 1978, 1984, 1989, 1994, 1999, 2004, 2009 ,2014, Inf),
    #                         labels = c("1974-1979", "1982-1990", "1990-1998", 
    #                                    "1998-2006", "2006-2014", "2014-2022"),
    #                         right = TRUE
    #                         )
    #                       ),
    period_decade = as.factor(
                            cut(
                            year,
                            breaks = c(1973, 1979, 1989, 1999, 2009, 2019, Inf),
                            labels = c("1974-1979", "1980-1989", "1990-1999",
                                       "2000-2009", "2010-2019", "2020-2024"),
                            right = TRUE
                            )
                          ),
    age_group = as.factor( 
                            cut(
                            age,
                            breaks = c(17, 29, 39, 49, 59, 69, Inf),
                            labels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+"),
                            right = TRUE
                          )),
    age_groups = as.factor( 
                            cut(
                            age,
                            breaks = c(17, 29, 39, 49, 59, 69, Inf),
                            labels = c("18-29", "30-39", "40-49", "50-59", "60-69", "70+"),
                            right = TRUE
                          )),
    age_group_small = as.factor( 
                            cut(
                              age,
                              breaks = c(seq(15, 75, by = 5), Inf),  # Define breaks up to 75 and include Inf for the last group
                              labels = c("16-20", "21-25", "26-30", "31-35", "36-40", "41-45", "46-50", "51-55", "56-60", "61-65", "66-70", "71-75", "76+"),
                              right = FALSE  # Makes intervals left-closed, i.e., [x, y)
                            )
                          )
                          ,
    generation_5total = factor(
      case_when(
        cohort >= 1901 & cohort <= 1927 ~ "Greatest (1901-1927)",
        cohort >= 1928 & cohort <= 1945 ~ "Silent (1928-1945)",
        cohort >= 1946 & cohort <= 1964 ~ "Boomers (1946-1964)",
        cohort >= 1965 & cohort <= 1980 ~ "Gen X (1965-1980)",
        # cohort >= 1981 & cohort <= 1996 ~ "Millennials (1981-1996)",
        # cohort >= 1997 & cohort <= 2012 ~ "Gen Z (1997-2012)",
        cohort >= 1981  ~ "Millennials / Gen Z (1981-2004)",
        TRUE ~ "Other"
      ),
      levels = c(
        "Greatest (1901-1927)",
        "Silent (1928-1945)",
        "Boomers (1946-1964)",
        "Gen X (1965-1980)",
        "Millennials / Gen Z (1981-2004)"
       # "Millennials (1981-1996)",
      #  "Gen Z (1997-2012)"#,
     #   "Other"
      )
    ),
    generation_10total = factor(
      case_when(
        generation == "Greatest (1901-1927)" & cohort <= 1914 ~ "Greatest Early (1901-1914)",
        generation == "Greatest (1901-1927)" & cohort > 1914 ~ "Greatest Late (1915-1927)",
        generation == "Silent (1928-1945)" & cohort <= 1936 ~ "Silent Early (1928-1936)",
        generation == "Silent (1928-1945)" & cohort > 1936 ~ "Silent Late (1937-1945)",
        generation == "Boomers (1946-1964)" & cohort <= 1955 ~ "Boomers Early (1946-1955)",
        generation == "Boomers (1946-1964)" & cohort > 1955 ~ "Boomers Late (1956-1964)",
        generation == "Gen X (1965-1980)" & cohort <= 1972 ~ "Gen X Early (1965-1972)",
        generation == "Gen X (1965-1980)" & cohort > 1972 ~ "Gen X Late (1973-1980)",
        cohort >= 1981 & cohort <= 1988 ~ "Millennials Early (1981-1988)",
        cohort > 1988 ~ "Millennials Late / Gen Z Early (1989-2004)",
        # generation == "Millennials (1981-1996)" & cohort > 1988 ~ "Millennials Late (1989-1996)",
        # generation == "Gen Z (1997-2012)" & cohort <= 2004 ~ "Gen Z Early (1997-2004)",
        # generation == "Gen Z (1997-2012)" & cohort > 2004 ~ "Gen Z Late (2005-2012)",
        TRUE ~ "Other"
      ),
      levels = c(
        "Greatest Early (1901-1914)", "Greatest Late (1915-1927)",
        "Silent Early (1928-1936)", "Silent Late (1937-1945)",
        "Boomers Early (1946-1955)", "Boomers Late (1956-1964)",
        "Gen X Early (1965-1972)", "Gen X Late (1973-1980)",
        "Millennials Early (1981-1988)", "Millennials Late / Gen Z Early (1989-2004)"
      #  "Millennials Early (1981-1988)", "Millennials Late (1989-1996)",
      #  "Gen Z Early (1997-2004)", "Gen Z Late (2005-2012)"#,
     #   "Other"
      )
    ),
    generation_15total = factor(
      case_when(
        cohort >= 1900 & cohort <= 1910 ~ "Greatest Early (1901-1910)",
        cohort >= 1911 & cohort <= 1918 ~ "Greatest Mid (1911-1918)",
        cohort >= 1919 & cohort <= 1927 ~ "Greatest Late (1919-1927)",
        cohort >= 1928 & cohort <= 1934 ~ "Silent Early (1928-1934)",
        cohort >= 1935 & cohort <= 1940 ~ "Silent Mid (1935-1940)",
        cohort >= 1941 & cohort <= 1945 ~ "Silent Late (1941-1945)",
        cohort >= 1945 & cohort <= 1951 ~ "Boomers Early (1946-1951)",
        cohort >= 1952 & cohort <= 1958 ~ "Boomers Mid (1952-1958)",
        cohort >= 1959 & cohort <= 1964 ~ "Boomers Late (1959-1964)",
        cohort >= 1965 & cohort <= 1970 ~ "Gen X Early (1965-1970)",
        cohort >= 1971 & cohort <= 1976 ~ "Gen X Mid (1971-1976)",
        cohort >= 1977 & cohort <= 1980 ~ "Gen X Late (1977-1980)",
        cohort >= 1981 & cohort <= 1986 ~ "Millennials Early (1981-1986)",
        cohort >= 1987 & cohort <= 1992 ~ "Millennials Mid (1987-1992)",
        cohort >= 1993 ~ "Millennials Late / Gen Z (1993-2004)",
    #    generation == "Gen Z (1997-2012)" & cohort <= 2002 ~ "Gen Z Early (1997-2002)",
    #    generation == "Gen Z (1997-2012)" & cohort > 2002 & cohort <= 2008 ~ "Gen Z Mid (2003-2008)",
    #    generation == "Gen Z (1997-2012)" & cohort > 2008 ~ "Gen Z Late (2009-2012)",
        TRUE ~ "Other"
      ),
      levels = c(
        "Greatest Early (1901-1910)", "Greatest Mid (1911-1918)", "Greatest Late (1919-1927)",
        "Silent Early (1928-1934)", "Silent Mid (1935-1940)", "Silent Late (1941-1945)",
        "Boomers Early (1946-1951)", "Boomers Mid (1952-1958)", "Boomers Late (1959-1964)",
        "Gen X Early (1965-1970)", "Gen X Mid (1971-1976)", "Gen X Late (1977-1980)",
        "Millennials Early (1981-1986)", "Millennials Mid (1987-1992)", 
        "Millennials Late / Gen Z (1993-2004)"
        #"Millennials Late (1993-1996)",
      #  "Gen Z Early (1997-2002)", "Gen Z Mid (2003-2008)", "Gen Z Late (2009-2012)" #,
      #  "Other"
      )
    )
  ) %>% 
  mutate(
    # 2-year periods
    period_2yr = as.factor(
      cut(
        year,
        breaks = c(
          1973, 1975, 1977, 1979, 1981, 1983, 1985, 1987, 1989, 1991,
          1993, 1995, 1997, 1999, 2001, 2003, 2005, 2007, 2009, 2011,
          2013, 2015, 2017, 2019, 2021, Inf
        ),
        labels = c(
          "1974-1975", "1976-1977", "1978-1979", "1980-1981", "1982-1983",
          "1984-1985", "1986-1987", "1988-1989", "1990-1991", "1992-1993",
          "1994-1995", "1996-1997", "1998-1999", "2000-2001", "2002-2003",
          "2004-2005", "2006-2007", "2008-2009", "2010-2011", "2012-2013",
          "2014-2015", "2016-2017", "2018-2019", "2020-2021", "2022-2022"
        ),
        right = TRUE
      )
    ),
    
    # 3-year periods
    period_3yr = as.factor(
      cut(
        year,
        breaks = c(
          1973, 1976, 1979, 1982, 1985, 1988, 1991, 1994, 1997, 2000,
          2003, 2006, 2009, 2012, 2015, 2018, 2021, Inf
        ),
        labels = c(
          "1974-1976", "1977-1979", "1980-1982", "1983-1985",
          "1986-1988", "1989-1991", "1992-1994", "1995-1997",
          "1998-2000", "2001-2003", "2004-2006", "2007-2009",
          "2010-2012", "2013-2015", "2016-2018", "2019-2021", "2022-2022"
        ),
        right = TRUE
      )
    ),
    
    # 5-year periods
    period_5yr = as.factor(
      cut(
        year,
        breaks = c(
          1973, 1978, 1983, 1988, 1993, 1998, 2003, 2008, 2013, 2018, Inf
        ),
        labels = c(
          "1974-1978", "1979-1983", "1984-1988", "1989-1993",
          "1994-1998", "1999-2003", "2004-2008", "2009-2013",
          "2014-2018", "2019-2022"
        ),
        right = TRUE
      )
    ),
    
    # 7-year periods
    period_7yr = as.factor(
      cut(
        year,
        breaks = c(
          1973, 1980, 1987, 1994, 2001, 2008, 2015, Inf
        ),
        labels = c(
          "1974-1980", "1981-1987", "1988-1994",
          "1995-2001", "2002-2008", "2009-2015", "2016-2022"
        ),
        right = TRUE
      )
    ),
    
    # 8-year periods
    period_8yr = as.factor(
      cut(
        year,
        breaks = c(
          1973, 1981, 1989, 1997, 2005, 2013, 2021, Inf
        ),
        labels = c(
          "1974-1981", "1982-1989", "1990-1997",
          "1998-2005", "2006-2013", "2014-2021", "2022-2022"
        ),
        right = TRUE
      )
    ),
    
    # 10-year periods
    period_10yr = as.factor(
      cut(
        year,
        breaks = c(
          1973, 1983, 1993, 2003, 2013, Inf
        ),
        labels = c(
          "1974-1983", "1984-1993", "1994-2003",
          "2004-2013", "2014-2022"
        ),
        right = TRUE
      )
    )
  ) %>% 
  mutate(period_7total = period_7yr,
         period_10total = period_5yr,
         period_17total = period_3yr) %>% 
  mutate(
    # -----------------------
    # 6 AGE CATEGORIES
    # -----------------------
    age_6cat = as.factor(
      cut(
        age,
        # 7 breakpoints --> 6 intervals
        breaks = c(17, 30, 40, 50, 60, 70, 89),
        labels = c(
          "18-29",  # (17, 30]
          "30-39",  # (30, 40]
          "40-49",  # (40, 50]
          "50-59",  # (50, 60]
          "60-69",  # (60, 70]
          "70-89"   # (70, 89]
        ),
        right = TRUE
      )
    ),
    
    # -----------------------
    # 10 AGE CATEGORIES
    # -----------------------
    age_10cat = as.factor(
      cut(
        age,
        # 11 breakpoints --> 10 intervals
        breaks = c(17, 24, 31, 38, 45, 52, 59, 66, 73, 80, 89),
        labels = c(
          "18-24",  # (17, 24]
          "25-31",  # (24, 31]
          "32-38",  # (31, 38]
          "39-45",  # (38, 45]
          "46-52",  # (45, 52]
          "53-59",  # (52, 59]
          "60-66",  # (59, 66]
          "67-73",  # (66, 73]
          "74-80",  # (73, 80]
          "81-89"   # (80, 89]
        ),
        right = TRUE
      )
    ),
    
    # -----------------------
    # 16 AGE CATEGORIES
    # -----------------------
    age_16cat = as.factor(
      cut(
        age,
        # 17 breakpoints --> 16 intervals
        breaks = c(17, 22, 27, 32, 37, 42, 47, 52, 57, 62, 67, 72, 77, 82, 85, 87, 89),
        labels = c(
          "18-22",  # (17, 22]
          "23-27",  # (22, 27]
          "28-32",  # (27, 32]
          "33-37",  # (32, 37]
          "38-42",  # (37, 42]
          "43-47",  # (42, 47]
          "48-52",  # (47, 52]
          "53-57",  # (52, 57]
          "58-62",  # (57, 62]
          "63-67",  # (62, 67]
          "68-72",  # (67, 72]
          "73-77",  # (72, 77]
          "78-82",  # (77, 82]
          "83-85",  # (82, 85]
          "86-87",  # (85, 87]
          "88-89"   # (87, 89]
        ),
        right = TRUE
      )
    )
    
  ) %>% 
    mutate(srh_num = health,
         srh_cat = health_cat) 

glimpse(data_gss)
summary(data_gss)

table(data_gss$generation_5total)
table(data_gss$generation_10total)
table(data_gss$generation_15total)

table(data_gss$period_7total)
table(data_gss$period_10total)
table(data_gss$period_17total)

table(data_gss$srh_num)

summary(data_gss$age)
```

Some interpretation guidance: https://chatgpt.com/share/676c4f33-a4b0-8002-adae-8f6368a9cf06 

```{r}


regressions <- data_gss %>%
  mutate(
    period    = period_7yr,
    cohort    = generation_10total,
    age_group = age_group
  ) %>%
  group_by(age_group, period) %>%
  nest() %>%
  mutate(
    model  = map(data, ~ lm(srh_num ~ age, data = .x)),
    tidied = map(model, tidy)
  ) %>%
  unnest(cols = tidied) %>%
  filter(term == "age") %>%
  select(age_group, period, estimate, std.error, p.value)

 regressions
```



```{r}


ggplot(regressions %>% filter(age_group != "16-20"),
       aes(x = period, y = estimate, color = age_group)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbar(
    aes(ymin = estimate - 1.96 * std.error,
        ymax = estimate + 1.96 * std.error),
    width = 0.2,
    position = position_dodge(width = 0.5)
  ) +
  labs(
    title = "Coefficient of Age on srh_num by Age Group and Period",
    x = "Period",
    y = "Estimate (Coefficient for Age)",
    color = "Age Group"
  ) +
  theme_minimal()


```

```{r}

ggplot(regressions, aes(x = period, y = estimate, fill = age_group)) +
  geom_col(position = position_dodge(width = 0.8)) +
  geom_errorbar(
    aes(ymin = estimate - 1.96 * std.error,
        ymax = estimate + 1.96 * std.error),
    width = 0.2,
    position = position_dodge(width = 0.8)
  ) +
  labs(
    title = "Coefficient of Age on srh_num by Age Group and Period",
    x = "Period",
    y = "Estimate (Coefficient for Age)",
    fill = "Age Group"
  ) +
  theme_minimal()


```

```{r}
ggplot(regressions, aes(x = period, y = estimate, color = age_group)) +
  geom_point() +
  geom_errorbar(
    aes(ymin = estimate - 1.96 * std.error,
        ymax = estimate + 1.96 * std.error),
    width = 0.2
  ) +
  facet_wrap(~ age_group) +
  labs(
    title = "Coefficient of Age on srh_num by Age Group and Period",
    x = "Period",
    y = "Estimate (Coefficient for Age)"
  ) +
  theme_minimal() 
```

```{r}
ggplot(regressions %>% filter(age_group != "16-20"),
       aes(x = period, y = estimate, color = age_group, group = age_group)) +
  geom_line() +
  geom_point() +
#  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.2) +
  labs(
    title = "Coefficient of Age on srh_num by Age Group and Period",
    x = "Period",
    y = "Estimate (Coefficient for Age)",
    color = "Age Group"
  ) +
  theme_minimal()
```


```{r}
ggplot(regressions %>% filter(age_group != "16-20"),
       aes(x = period, y = estimate, color = age_group, group = age_group)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.2) +
  facet_wrap(~ age_group) +
  labs(
    title = "Coefficient of Age on srh_num by Age Group and Period",
    x = "Period",
    y = "Estimate (Coefficient for Age)",
    color = "Age Group"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
  
  
  
ggplot(regressions %>% filter(age_group != "16-20"),
       aes(x = age_group, y = estimate, color = period, group = period)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.2) +
  facet_wrap(~ period) +
  labs(
    title = "Coefficient of Age on srh_num by Period and Age Group",
    x = "Period",
    y = "Estimate (Coefficient for Age)",
    color = "Age Group"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 


```




```

```{r}

ggplot(regressions, aes(x = period, y = age_group, fill = estimate)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(
    midpoint = 0, 
    low = "blue", mid = "white", high = "red",
    name = "Est."
  ) +
  labs(
    title = "Heatmap of Age Coefficients by Age Group and Period",
    x = "Period",
    y = "Age Group"
  ) +
  theme_minimal() +
  theme(panel.grid = element_blank())


```

```{r}

ggplot(regressions %>% filter(age_group != "16-20"), aes(x = period, y = estimate)) +
  geom_point() +
  facet_wrap(~ age_group) +
  labs(
    title = "Coefficient of Age on srh_num by Age Group and Period",
    x = "Period",
    y = "Estimate (Coefficient for Age)"
  ) +
  theme_minimal()

```


```{r}

ggplot(regressions %>% filter(age_group != "16-20"),
       aes(x = period, y = estimate, color = age_group, group = age_group)) +
  geom_line() +
  geom_point() +
#  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.2) +
  labs(
    title = "Coefficient of Age on srh_num by Age Group and Period",
    x = "Period",
    y = "Estimate (Coefficient for Age)",
    color = "Age Group"
  ) +
  theme_minimal()


ggplot(regressions %>% filter(age_group != "16-20"),
       aes(x = period, y = estimate, color = age_group, group = age_group)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.2) +
  facet_wrap(~ age_group) +
  labs(
    title = "Coefficient of Age on srh_num by Age Group and Period",
    x = "Period",
    y = "Estimate (Coefficient for Age)",
    color = "Age Group"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 
  
  
  
ggplot(regressions %>% filter(age_group != "16-20"),
       aes(x = age_group, y = estimate, color = age_group, group = period)) +
  geom_line() +
  geom_point() +
  geom_errorbar(aes(ymin = estimate - std.error, ymax = estimate + std.error), width = 0.2) +
  facet_wrap(~ period) +
  labs(
    title = "Coefficient of Age on srh_num by Period and Age Group",
    x = "Period",
    y = "Estimate (Coefficient for Age)",
    color = "Age Group"
  ) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) 


```

# Function

```{r}

library(dplyr)
library(purrr)
library(broom)

# 1. Define your regression function
my_regression_function <- function(data,
                                   period_col,
                                   cohort_col,
                                   age_group_col) {
  result_tbl <- data %>%
    mutate(
      period    = .data[[period_col]],
      cohort    = .data[[cohort_col]],
      age_group = .data[[age_group_col]]
    ) %>%
    group_by(age_group, period) %>%
    nest() %>%
    mutate(
      model  = map(data, ~ lm(srh_num ~ age, data = .x)),
      tidied = map(model, tidy)
    ) %>%
    unnest(cols = tidied) %>%
    filter(term == "age") %>%
    select(age_group, period, estimate, std.error, p.value, model)

  list(
    data_name            = deparse(substitute(data)),
    period_col_name      = period_col,
    cohort_col_name      = cohort_col,
    age_group_col_name   = age_group_col,
    regression_results   = result_tbl
  )
}

# 2. Vectors of column names
age_group_cols <- c("age_6cat", "age_10cat", "age_16cat")
period_cols    <- c("period_7total", "period_10total", "period_17total")
cohort_col     <- "generation_10total"

# 3. All combinations
param_combinations <- expand.grid(
  age_group_col = age_group_cols,
  period_col    = period_cols,
  stringsAsFactors = FALSE
)

# 4. Map the function across each row of param_combinations
results_list <- purrr::pmap(
  param_combinations,
  function(age_group_col, period_col) {
    my_regression_function(
      data          = data_gss,
      period_col    = period_col,
      cohort_col    = cohort_col,
      age_group_col = age_group_col
    )
  }
)

# 5. Store the results in a tibble
final_results <- param_combinations %>%
  mutate(regression_output = results_list)

# now you can inspect the results:
str(final_results$regression_output[[1]])
# or
final_results$regression_output[[2]]$regression_results

final_results %>% filter(age_group_col == "age_6cat") %>% filter(period_col == "period_7total") %>% pull(regression_output)

```

# Final Results

```{r}

final_results$regression_output[[1]]$regression_results
final_results$regression_output[[2]]$regression_results
final_results$regression_output[[3]]$regression_results
final_results$regression_output[[4]]$regression_results
final_results$regression_output[[5]]$regression_results
final_results$regression_output[[6]]$regression_results
final_results$regression_output[[7]]$regression_results
final_results$regression_output[[8]]$regression_results
final_results$regression_output[[9]]$regression_results


```


# Forest Plots

```{r}


# Coeff
model_gen5_period7 %>% 
#model_gen5_period10 %>% 
#model_gen5_period17 %>% 
#model_gen10_period7 %>% 
#model_gen10_period10 %>% 
#model_gen10_period17 %>% 
#model_gen15_period7 %>% 
#model_gen15_period10 %>% 
# model_gen15_period17 %>% 
tidy(conf.int = TRUE) %>%
  filter(grepl("generation_5total|period_7total", term)  ) %>% #,
       #  !grepl(":", term)) %>%  # remove interaction terms
ggplot(aes(x = estimate, y = term)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  labs(
    x = "Coefficient (SRH: 1=Poor, 4=Excellent)",
    y = "",
    title = "Coefficient Plot: Generation & Period (No Interaction Terms Shown)"
  ) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_minimal()


# interactions

model_gen5_period7 %>% 
tidy(conf.int = TRUE) %>%
  filter(#grepl("generation_5total|period_7total", term)  )#,
         grepl(":", term)) %>%  # remove interaction terms %>% 
ggplot(aes(x = estimate, y = term)) +
  geom_point() +
  geom_errorbarh(aes(xmin = conf.low, xmax = conf.high), height = 0.2) +
  labs(
    x = "Coefficient (SRH: 1=Poor, 4=Excellent)",
    y = "",
    title = "Coefficient Plot: Generation & Period (No Interaction Terms Shown)"
  ) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_minimal()


```
