---
title: "Aim 2 Analytic Pipeline - Skeleton - Graphical Analysis"
author: "Gloria Huei-Jong Graf"
date: "2/23/2023"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE)
library(tidyverse)
library(foreign)
library(survey)
library(srvyr)
library(patchwork)
library(kableExtra)
```

```{r loading-data, echo = FALSE, warnings = FALSE}
original_df = read.csv("C:/Users/glory/OneDrive - cumc.columbia.edu/Dissertation/Data/Core_Dataset_Aim2.csv") %>%
  mutate(race = factor(race, levels = c("White", "Black", "Other")),
         gender = factor(gender, levels = c(1, 2), labels = c("Men", "Women")),
         phenoageadv_delta_levinenocrp_scale = scale(phenoageadv_delta_levinenocrp_nosexstrat),
         phenoageadv_delta_elasticnet_scale = scale(phenoageadv_delta_elasticnet_nosexstrat),
         kdmadv_delta_levinenocrp_scale = scale(kdmadv_delta_levinenocrp_nosexstrat),
         kdmadv_delta_elasticnet_scale = scale(kdmadv_delta_elasticnet_nosexstrat),
         hdlog_levinenocrp_scale = scale(hdlog_levinenocrp_nosexstrat),
         hdlog_elasticnet_scale = scale(hdlog_elasticnet_nosexstrat))

# Creating dataset for description of cell-sizes (non-survey object, restricting to NHANES4, Black or White race, age <=80, complete information, n=28991)
descriptive_df =
  original_df %>%
  mutate(age_4yr = as.character(age_4yr),
         cohort_4yr = as.character(cohort_4yr),
         period_4yr = as.character(period_4yr),
         lnWTMEC4YR = ifelse(WTMEC4YR == 0, 0, log(WTMEC4YR))) %>%
  filter(year != 1991 & race %in% c("Black", "White") & age_4yr <= 80 &
          !is.na(phenoageadv_delta_elasticnet_nosexstrat) & !is.na(age) & !is.na(age_squared) & !is.na(race) & !is.na(gender) & !is.na(period_4yr) & !is.na(cohort_4yr)) %>%
  droplevels()

# Creating dataset for weighting models (survey object, n=55081, restricting to NHANES4)
surveymatrix_df =
  original_df %>%
  mutate(lnWTMEC4YR = ifelse(WTMEC4YR == 0, 0, log(WTMEC4YR))) %>%
  filter(year != 1991) %>%
  droplevels()

svydesign_all = 
  surveymatrix_df %>% 
  as_survey_design(ids=SDMVPSU, strata=SDMVSTRA, weights=WTMEC4YR, nest = TRUE)

svydesign_1d =
  subset(svydesign_all, race %in% c("Black", "White") & age_4yr <= 80 &
          !is.na(phenoageadv_delta_elasticnet_nosexstrat) & !is.na(age) & !is.na(age_squared) & !is.na(race) & !is.na(gender) & !is.na(period_4yr) & !is.na(cohort_4yr))

svydesign_2d =
  subset(svydesign_all, race %in% c("Black", "White") & age_4yr <= 80 &
          !is.na(phenoageadv_delta_elasticnet_nosexstrat) & !is.na(age) & !is.na(age_squared) & !is.na(race) & !is.na(gender) & !is.na(period_4yr) & !is.na(cohort_4yr))

svydesign_2d<-update(svydesign_2d, 
                     cohort_4yr = factor(cohort_4yr),
                     period_4yr = factor(period_4yr),
                     age_4yr = factor(age_4yr))
```

This document contains skeleton code needed to conduct initial graphical analysis of APC effects for Aim 2 of Graf's dissertation. APC analyses follow the general approach followed by Yang and Land (specified below). This document covers steps 1 and 2. 

1) Construction of synthetic birth cohorts

2) Graphical assessment of age, period, and cohort effects

3) Comparison of one-, two-, and three-factor models

4) Identification of APC effects using different statistical approaches: the Constrained Coefficients GLIM estimator (CGLIM), Hierarchical APC Estimation with Cross-Classified Random Effects (HAPC-CCREM), and The Median Polish Estimator.

## Participant counts and data setup

```{r echo = FALSE}
table(descriptive_df$period_4yr, descriptive_df$race, descriptive_df$gender) %>%
  knitr::kable(caption = "Number of participants by wave", col.names = c("Wave", "Race", "Sex", "n")) %>%
  kableExtra::kable_classic(html_font = "Times New Roman") %>%
  row_spec(0, bold=TRUE)

table(descriptive_df$cohort_4yr, descriptive_df$race, descriptive_df$gender) %>%
  knitr::kable(caption = "Number of participants by 4-year birth cohorts", col.names = c("Cohort", "Race", "Sex", "n")) %>%
  kableExtra::kable_classic(html_font = "Times New Roman") %>%
  row_spec(0, bold=TRUE)
```

```{r functions-1d, echo = FALSE, warning = FALSE, message = FALSE}
plot1D_all_f = function(var1, bioage_var, ymin, ymax){
    
  var1 = enquo(var1)
  bioage_var = enquo(bioage_var)
  
  df =
    svydesign_1d %>%
    group_by(!!var1) %>%
    summarise(bioage_var = survey_mean(!!bioage_var, na.rm = T)) %>%
    mutate(bioage_var = ifelse(bioage_var == 0, NA, bioage_var))
  
  ggplot(df, aes(x = !!var1, y = bioage_var, group = 1)) +
    geom_line(linewidth = 0.5) +
    geom_point() +
    theme_minimal() + 
    ylab(as_label(bioage_var)) +
    ylim(ymin, ymax)

}

plot1D_strat_f = function(var1, bioage_var, ymin, ymax){
  
  var1 = enquo(var1)
  bioage_var = enquo(bioage_var)
  
  df =
    svydesign_1d %>%
    group_by(!!var1, gender, race) %>%
    summarise(bioage_var = survey_mean(!!bioage_var, na.rm = T)) %>%
    mutate(bioage_var = ifelse(bioage_var == 0, NA, bioage_var))
  
  ggplot(df, aes(x = !!var1, y = bioage_var, group = 1)) +
    geom_line(linewidth = 0.5) +
    geom_point() +
    facet_wrap(c("gender", "race")) +
    theme_minimal() + 
    ylab(as_label(bioage_var)) +
    ylim(ymin, ymax)
  
}
```

## Graphical assessment of one-dimensional age, period, and cohort effects

```{r echo = FALSE,  warnings = FALSE, message = FALSE, fig.width=18, fig.height=14}
a_phenoageadv_delta_elasticnet_scale_all = plot1D_all_f(age_4yr, phenoageadv_delta_elasticnet_scale, -1, 1)
a_kdmadv_delta_elasticnet_scale_all = plot1D_all_f(age_4yr, kdmadv_delta_elasticnet_scale, -2, 2)
a_hdlog_elasticnet_scale_all = plot1D_all_f(age_4yr, hdlog_elasticnet_scale, -2, 2)
```

```{r echo = FALSE,  warnings = FALSE, message = FALSE, fig.width=18, fig.height=14}
p_phenoageadv_delta_elasticnet_scale_all = plot1D_all_f(period_4yr, phenoageadv_delta_elasticnet_scale, -1, 1)
p_kdmadv_delta_elasticnet_scale_all = plot1D_all_f(period_4yr, kdmadv_delta_elasticnet_scale, -2, 2)
p_hdlog_elasticnet_scale_all = plot1D_all_f(period_4yr, hdlog_elasticnet_scale, -2, 2)
```

```{r echo = FALSE,  warnings = FALSE, message = FALSE, fig.width=18, fig.height=14}
c_phenoageadv_delta_elasticnet_scale_all = plot1D_all_f(cohort_4yr, phenoageadv_delta_elasticnet_scale, -1, 1)
c_kdmadv_delta_elasticnet_scale_all = plot1D_all_f(cohort_4yr, kdmadv_delta_elasticnet_scale, -2, 2)
c_hdlog_elasticnet_scale_all = plot1D_all_f(cohort_4yr, hdlog_elasticnet_scale, -2, 2)
```

```{r echo = FALSE,  warnings = FALSE, message = FALSE, fig.width=18, fig.height=14}
(a_phenoageadv_delta_elasticnet_scale_all + p_phenoageadv_delta_elasticnet_scale_all + c_phenoageadv_delta_elasticnet_scale_all) / (a_hdlog_elasticnet_scale_all + p_hdlog_elasticnet_scale_all + c_hdlog_elasticnet_scale_all) + plot_layout(guides = "collect")

rm(a_phenoageadv_delta_elasticnet_scale_all, a_kdmadv_delta_elasticnet_scale_all, a_hdlog_elasticnet_scale_all)
rm(p_phenoageadv_delta_elasticnet_scale_all, p_kdmadv_delta_elasticnet_scale_all, p_hdlog_elasticnet_scale_all)
rm(c_phenoageadv_delta_elasticnet_scale_all, c_kdmadv_delta_elasticnet_scale_all, c_hdlog_elasticnet_scale_all)
```

### Age effects - stratified
http://127.0.0.1:22021/graphics/plot_zoom_png?width=948&height=565
```{r echo = FALSE,  warnings = FALSE, message = FALSE, fig.width=18, fig.height=14}
a_phenoageadv_delta_elasticnet_scale_strat = plot1D_strat_f(age_4yr, phenoageadv_delta_elasticnet_scale, -2, 2)
a_kdmageadv_delta_elasticnet_scale_strat = plot1D_strat_f(age_4yr, kdmadv_delta_elasticnet_scale, -2, 2)
a_hdlog_elasticnet_scale_strat = plot1D_strat_f(age_4yr, hdlog_elasticnet_scale, -2, 2)
```

### Period effects - stratified
```{r echo = FALSE,  warnings = FALSE, message = FALSE, fig.width=18, fig.height=14}
p_phenoageadv_delta_elasticnet_scale_strat = plot1D_strat_f(period_4yr, phenoageadv_delta_elasticnet_scale, -2, 2)
p_kdmadv_delta_elasticnet_scale_strat = plot1D_strat_f(period_4yr, kdmadv_delta_elasticnet_scale, -2, 2)
p_hdlog_elasticnet_scale_strat = plot1D_strat_f(period_4yr, hdlog_elasticnet_scale, -2, 2)
```

### Cohort effects - stratified
```{r echo = FALSE,  warnings = FALSE, message = FALSE, fig.width=18, fig.height=14}
c_phenoageadv_delta_elasticnet_scale_strat = plot1D_strat_f(cohort_4yr, phenoageadv_delta_elasticnet_scale, -2, 2)
c_kdmageadv_delta_elasticnet_scale_strat = plot1D_strat_f(cohort_4yr, kdmadv_delta_elasticnet_scale, -2, 2)
c_hdlog_elasticnet_scale_strat = plot1D_strat_f(cohort_4yr, hdlog_elasticnet_scale, -2, 2)
```

```{r echo = FALSE,  warnings = FALSE, message = FALSE, fig.width=18, fig.height=14}
a_phenoageadv_delta_elasticnet_scale_strat + p_phenoageadv_delta_elasticnet_scale_strat + c_phenoageadv_delta_elasticnet_scale_strat
  
a_hdlog_elasticnet_scale_strat + p_hdlog_elasticnet_scale_strat + c_hdlog_elasticnet_scale_strat

rm(a_phenoageadv_delta_elasticnet_scale_strat, a_kdmageadv_delta_elasticnet_scale_strat, a_hdlog_elasticnet_scale_strat, p_phenoageadv_delta_elasticnet_scale_strat, p_kdmadv_delta_elasticnet_scale_strat, p_hdlog_elasticnet_scale_strat, c_phenoageadv_delta_elasticnet_scale_strat, c_kdmageadv_delta_elasticnet_scale_strat, c_hdlog_elasticnet_scale_strat)
```


## Graphical assessment of two-dimensional age effects

```{r functions-2d, echo = FALSE, warning = FALSE, message = FALSE}
plot2D_all_f = function(var1, var2, bioage_var, ymin, ymax){

  var1 = enquo(var1)
  var2 = enquo(var2)
  bioage_var = enquo(bioage_var)
  
  df =
    svydesign_2d %>%
    group_by(!!var1, !!var2) %>%
    summarise(bioage_var = survey_mean(!!bioage_var, na.rm = T)) %>%
    ungroup() %>%
    mutate(bioage_var = ifelse(bioage_var == 0, NA, bioage_var))
  
  colors <- colorRampPalette(c("blue", "red"))(length(levels(pull(df, !!var2))))
  
  ggplot(df, aes(x = !!var1, y = bioage_var)) +
    geom_line(aes(color = !!var2, group = !!var2), linewidth = 0.5) +
    geom_point(aes(color = !!var2)) +
    theme_minimal() + 
    ylab(as_label(bioage_var)) +
    ylim(ymin, ymax) +
    scale_x_discrete(breaks=c(20, 40, 60, 80)) +
    scale_color_manual(values=colors)

}

plot2D_strat_f = function(var1, var2, bioage_var, ymin, ymax){
    
  var1 = enquo(var1)
  var2 = enquo(var2)
  bioage_var = enquo(bioage_var)
  
  df =
    svydesign_2d %>%
    group_by(!!var1, !!var2, gender, race) %>%
    summarise(bioage_var = survey_mean(!!bioage_var, na.rm = T)) %>%
    mutate(bioage_var = ifelse(bioage_var == 0, NA, bioage_var))
  
  colors <- colorRampPalette(c("blue", "red"))(length(levels(pull(df, !!var2))))

  ggplot(df, aes(x = !!var1, y = bioage_var)) +
    geom_line(aes(color = !!var2, group = !!var2), linewidth = 0.5) +
    geom_point(aes(color = !!var2)) +
    theme_minimal() + 
    ylab(as_label(bioage_var)) +
    facet_wrap(c("gender", "race")) +
    ylim(ymin, ymax) +
    scale_x_discrete(breaks=c(20, 40, 60, 80)) +
    scale_color_manual(values=colors)
  
}
```

### Age effects by period - overall

```{r echo = FALSE, warnings = FALSE, message = FALSE, fig.width=18, fig.height=14}
ap_phenoageadv_delta_elasticnet_scale_all = plot2D_all_f(age_4yr, period_4yr, phenoageadv_delta_elasticnet_scale, -1.5, 1.5)
ap_kdmadv_delta_elasticnet_scale_all = plot2D_all_f(age_4yr, period_4yr, kdmadv_delta_elasticnet_scale, -1.5, 1.5)
ap_hdlog_elasticnet_scale_all = plot2D_all_f(age_4yr, period_4yr, hdlog_elasticnet_scale, -1.5, 1.5)
```

### Age effects by period - stratified

```{r echo = FALSE,  warnings = FALSE, message = FALSE, fig.width=18, fig.height=14}
ap_phenoageadv_delta_elasticnet_scale_strat = plot2D_strat_f(age_4yr, period_4yr, phenoageadv_delta_elasticnet_scale, -1.5, 1.5)
ap_kdmadv_delta_elasticnet_scale_strat = plot2D_strat_f(age_4yr, period_4yr, kdmadv_delta_elasticnet_scale, -1.5, 1.5)
ap_hdlog_elasticnet_scale_strat = plot2D_strat_f(age_4yr, period_4yr, hdlog_elasticnet_scale, -1.5, 1.5)
```

```{r echo = FALSE,  warnings = FALSE, message = FALSE, fig.width=18, fig.height=14}
(ap_phenoageadv_delta_elasticnet_scale_all + ap_phenoageadv_delta_elasticnet_scale_strat) /
(ap_hdlog_elasticnet_scale_all + ap_hdlog_elasticnet_scale_strat) + plot_layout(guides = "collect")

rm(ap_phenoageadv_delta_elasticnet_scale_all, ap_kdmadv_delta_elasticnet_scale_all, ap_hdlog_elasticnet_scale_all, ap_phenoageadv_delta_elasticnet_scale_strat, ap_kdmadv_delta_elasticnet_scale_strat, ap_hdlog_elasticnet_scale_strat)
```


### Age effects by cohort - overall

```{r echo = FALSE,  warnings = FALSE, message = FALSE, fig.width=18, fig.height=14}
ac_phenoageadv_delta_elasticnet_scale_all = plot2D_all_f(age_4yr, cohort_4yr, phenoageadv_delta_elasticnet_scale, -1.5, 1.5)
ac_kdmadv_delta_elasticnet_scale_all = plot2D_all_f(age_4yr, cohort_4yr, kdmadv_delta_elasticnet_scale, -1.5, 1.5)
ac_hdlog_elasticnet_scale_all = plot2D_all_f(age_4yr, cohort_4yr, hdlog_elasticnet_scale, -1.5, 1.5)
```

### Age effects by cohort - stratified

```{r echo = FALSE,  warnings = FALSE, message = FALSE, fig.width=18, fig.height=14}
ac_phenoageadv_delta_elasticnet_scale_strat = plot2D_strat_f(age_4yr, cohort_4yr, phenoageadv_delta_elasticnet_scale, -1.5, 1.5)
ac_kdmadv_delta_elasticnet_scale_strat = plot2D_strat_f(age_4yr, cohort_4yr, kdmadv_delta_elasticnet_scale, -1.5, 1.5)
ac_hdlog_elasticnet_scale_strat = plot2D_strat_f(age_4yr, cohort_4yr, hdlog_elasticnet_scale, -1.5, 1.5)
```

```{r echo = FALSE,  warnings = FALSE, message = FALSE, fig.width=18, fig.height=14}
(ac_phenoageadv_delta_elasticnet_scale_all + ac_phenoageadv_delta_elasticnet_scale_strat) / (ac_hdlog_elasticnet_scale_all + ac_hdlog_elasticnet_scale_strat) +  plot_layout(guides = "collect")

rm(ac_phenoageadv_delta_elasticnet_scale_all, ac_kdmadv_delta_elasticnet_scale_all, ac_hdlog_elasticnet_scale_all, ac_phenoageadv_delta_elasticnet_scale_strat, ac_kdmadv_delta_elasticnet_scale_strat, ac_hdlog_elasticnet_scale_strat)
```


## Graphical assessment of two-dimensional period effects

### Period effects by age - overall

```{r echo = FALSE,  warnings = FALSE, message = FALSE, fig.width=18, fig.height=14}
pa_phenoageadv_delta_elasticnet_scale_all = plot2D_all_f(period_4yr, age_4yr, phenoageadv_delta_elasticnet_scale, -2, 2)
pa_kdmadv_delta_elasticnet_scale_all = plot2D_all_f(period_4yr, age_4yr, kdmadv_delta_elasticnet_scale, -2, 2)
pa_hdlog_elasticnet_scale_all = plot2D_all_f(period_4yr, age_4yr, hdlog_elasticnet_scale, -2, 2)

pa_phenoageadv_delta_elasticnet_scale_all + pa_kdmadv_delta_elasticnet_scale_all + pa_hdlog_elasticnet_scale_all + plot_layout(guides = "collect")

rm(pa_phenoageadv_delta_elasticnet_scale_all, pa_kdmadv_delta_elasticnet_scale_all, pa_hdlog_elasticnet_scale_all)
```

### Period effects by age - stratified

```{r echo = FALSE,  warnings = FALSE, message = FALSE, fig.width=18, fig.height=14}
pa_phenoageadv_delta_elasticnet_scale_strat = plot2D_strat_f(period_4yr, age_4yr, phenoageadv_delta_elasticnet_scale, -2, 2)
pa_kdmadv_delta_elasticnet_scale_strat = plot2D_strat_f(period_4yr, age_4yr, kdmadv_delta_elasticnet_scale, -2, 2)
pa_hdlog_elasticnet_scale_strat = plot2D_strat_f(period_4yr, age_4yr, hdlog_elasticnet_scale, -2, 2)

pa_phenoageadv_delta_elasticnet_scale_strat + pa_kdmadv_delta_elasticnet_scale_strat + pa_hdlog_elasticnet_scale_strat + plot_layout(guides = "collect")

rm(pa_phenoageadv_delta_elasticnet_scale_strat, pa_kdmadv_delta_elasticnet_scale_strat, pa_hdlog_elasticnet_scale_strat)
```

### Period effects by cohort - overall

```{r echo = FALSE,  warnings = FALSE, message = FALSE, fig.width=18, fig.height=14}
pc_phenoageadv_delta_elasticnet_scale_all = plot2D_all_f(period_4yr, cohort_4yr, phenoageadv_delta_elasticnet_scale, -2, 2)
pc_kdmadv_delta_elasticnet_scale_all = plot2D_all_f(period_4yr, cohort_4yr, kdmadv_delta_elasticnet_scale, -2, 2)
pc_hdlog_elasticnet_scale_all = plot2D_all_f(period_4yr, cohort_4yr, hdlog_elasticnet_scale, -2, 2)

pc_phenoageadv_delta_elasticnet_scale_all + pc_kdmadv_delta_elasticnet_scale_all + pc_hdlog_elasticnet_scale_all + plot_layout(guides = "collect")

rm(pc_phenoageadv_delta_elasticnet_scale_all, pc_kdmadv_delta_elasticnet_scale_all, pc_hdlog_elasticnet_scale_all)
```

### Period effects by cohort - stratified

```{r echo = FALSE,  warnings = FALSE, message = FALSE, fig.width=18, fig.height=14}
pc_phenoageadv_delta_elasticnet_scale_strat = plot2D_strat_f(period_4yr, cohort_4yr, phenoageadv_delta_elasticnet_scale, -2, 2)
pc_kdmadv_delta_elasticnet_scale_strat = plot2D_strat_f(period_4yr, cohort_4yr, kdmadv_delta_elasticnet_scale, -2, 2)
pc_hdlog_elasticnet_scale_strat = plot2D_strat_f(period_4yr, cohort_4yr, hdlog_elasticnet_scale, -2, 2)

pc_phenoageadv_delta_elasticnet_scale_strat + pc_kdmadv_delta_elasticnet_scale_strat + pc_hdlog_elasticnet_scale_strat + plot_layout(guides = "collect")

rm(pc_phenoageadv_delta_elasticnet_scale_strat, pc_kdmadv_delta_elasticnet_scale_strat, pc_hdlog_elasticnet_scale_strat)
```

## Graphical assessment of two-dimensional cohort effects

### Cohort effects by age - overall

```{r echo = FALSE,  warnings = FALSE, message = FALSE, fig.width=18, fig.height=14}
ca_phenoageadv_delta_elasticnet_scale_all = plot2D_all_f(cohort_4yr, age_4yr, phenoageadv_delta_elasticnet_scale, -2, 2)
ca_kdmadv_delta_elasticnet_scale_all = plot2D_all_f(cohort_4yr, age_4yr, kdmadv_delta_elasticnet_scale, -2, 2)
ca_hdlog_elasticnet_scale_all = plot2D_all_f(cohort_4yr, age_4yr, hdlog_elasticnet_scale, -2, 2)

ca_phenoageadv_delta_elasticnet_scale_all + ca_kdmadv_delta_elasticnet_scale_all + ca_hdlog_elasticnet_scale_all + plot_layout(guides = "collect")

rm(ca_phenoageadv_delta_elasticnet_scale_all, ca_kdmadv_delta_elasticnet_scale_all, ca_hdlog_elasticnet_scale_all)
```

### Cohort effects by age - stratified

```{r echo = FALSE,  warnings = FALSE, message = FALSE, fig.width=18, fig.height=14}
ca_phenoageadv_delta_elasticnet_scale_strat = plot2D_strat_f(cohort_4yr, age_4yr, phenoageadv_delta_elasticnet_scale, -2, 2)
ca_kdmadv_delta_elasticnet_scale_strat = plot2D_strat_f(cohort_4yr, age_4yr, kdmadv_delta_elasticnet_scale, -2, 2)
ca_hdlog_elasticnet_scale_strat = plot2D_strat_f(cohort_4yr, age_4yr, hdlog_elasticnet_scale, -2, 2)

ca_phenoageadv_delta_elasticnet_scale_strat + ca_kdmadv_delta_elasticnet_scale_strat + ca_hdlog_elasticnet_scale_strat + plot_layout(guides = "collect")

rm(ca_phenoageadv_delta_elasticnet_scale_strat, ca_kdmadv_delta_elasticnet_scale_strat, ca_hdlog_elasticnet_scale_strat)
```

### Cohort effects by period - overall

```{r echo = FALSE,  warnings = FALSE, message = FALSE, fig.width=18, fig.height=14}
cp_phenoageadv_delta_elasticnet_scale_all = plot2D_all_f(cohort_4yr, period_4yr, phenoageadv_delta_elasticnet_scale, -2, 2)
cp_kdmadv_delta_elasticnet_scale_all = plot2D_all_f(cohort_4yr, period_4yr, kdmadv_delta_elasticnet_scale, -2, 2)
cp_hdlog_elasticnet_scale_all = plot2D_all_f(cohort_4yr, period_4yr, hdlog_elasticnet_scale, -2, 2)

cp_phenoageadv_delta_elasticnet_scale_all + cp_kdmadv_delta_elasticnet_scale_all + cp_hdlog_elasticnet_scale_all + plot_layout(guides = "collect")

rm(cp_phenoageadv_delta_elasticnet_scale_all, cp_kdmadv_delta_elasticnet_scale_all, cp_hdlog_elasticnet_scale_all)
```

### Cohort effects by period - stratified

```{r echo = FALSE,  warnings = FALSE, message = FALSE, fig.width=18, fig.height=14}
cp_phenoageadv_delta_elasticnet_scale_strat = plot2D_strat_f(cohort_4yr, period_4yr, phenoageadv_delta_elasticnet_scale, -2, 2)
cp_kdmadv_delta_elasticnet_scale_strat = plot2D_strat_f(cohort_4yr, period_4yr, kdmadv_delta_elasticnet_scale, -2, 2)
cp_hdlog_elasticnet_scale_strat = plot2D_strat_f(cohort_4yr, period_4yr, hdlog_elasticnet_scale, -2, 2)

cp_phenoageadv_delta_elasticnet_scale_strat + cp_kdmadv_delta_elasticnet_scale_strat + cp_hdlog_elasticnet_scale_strat + plot_layout(guides = "collect")

rm(cp_phenoageadv_delta_elasticnet_scale_strat, cp_kdmadv_delta_elasticnet_scale_strat, cp_hdlog_elasticnet_scale_strat)
```
